---
title: "Simple Growth Regime"
---

```{r, include = FALSE}
library(tidyverse)
library(ggpubr)
library(egg)
```

**Purpose:** Derive and visualize the growth regime in two alternative habitat patches *sensu* Armstrong et al. (2021).


Source dependencies
```{r include=FALSE}
source("Habitat.R")
source("Functions.R")
```

Load pre-calculated growth
```{r}
load("data/wt.growth.array.RData")
```


## Simple growth regime

Derive a generic growth regime for a 10 gram fish, sensu Armstrong et al. (2021). *For simplicity, I did not simulate changing growth potential due to increase in fish mass over the course of the year, as in the NCC paper.*

*Notes:*

* In these ~simple scenarios, growth potential for the "optimal mover" is just calculated from the maximum of that across the two habitats. No need to simulate habitat selection/movement.


### Get growth potential

Objects
```{r}
waterTemps <- cbind("pid" = seq(1, length(seq(0.1, 25, 0.1))),"WT" = seq(0.1, 25, 0.1))
rations <- seq(0.02, 0.17, 0.001)
weights <- seq(1, 1500, 1)
```

Get growth potential by temperature for a 10 gram fish, fed at ~max ration
```{r}
# get relative growth rates for a 10 g fish across range of water temps
grate_df <- tibble(temp = waterTemps[,2],
                   grate_ggd = wt.growth[,151,10]) %>%
  mutate(temp = trimws(temp))

# calculate growth potential (g/d/d) and cumulative growth (g) across warm and cold patches and for an "optimal mover"
growthregime_df <- habitat_df %>%
  mutate(temp_warm = trimws(temp_warm),
         temp_cold = trimws(temp_cold)) %>%
  left_join(grate_df, by = c("temp_warm" = "temp")) %>% rename(grpot_warm = grate_ggd) %>%
  left_join(grate_df, by = c("temp_cold" = "temp")) %>% rename(grpot_cold = grate_ggd) %>%
  mutate(grpot_track = pmax(grpot_warm, grpot_cold)) %>%
  mutate(cumul_growth_warm = cumsum(grpot_warm*10),
         cumul_growth_cold = cumsum(grpot_cold*10),
         cumul_growth_track = cumsum(grpot_track*10),
         habitat = case_when(
           grpot_cold > grpot_warm ~ "cold",
           grpot_warm > grpot_cold ~ "warm",
           TRUE ~ "tie"
         ))
  
head(growthregime_df)
```


### Visualize output

Plot growth regimes (growth potential and g/g/d)
```{r}
p1 <- growthregime_df %>% ggplot() +
  geom_line(aes(x = date, y = grpot_warm), col = 2) +
  geom_line(aes(x = date, y = grpot_cold), col = 4) +
  theme_bw() +
  xlab("Date") +
  ylab("Growth potential (g/g/day)") +
  labs(title = "Warm vs. cold habitats")

p2 <- growthregime_df %>% ggplot() +
  geom_line(aes(x = date, y = grpot_track), col = 1, linetype = "dashed") +
  theme_bw() +
  xlab("Date") +
  ylab("Growth potential (g/g/day)") +
  labs(title = "Optimal mover")

ggarrange(p1, p2, ncol = 2)
```

Plot cumulative growth
```{r}
growthregime_df %>% ggplot() +
  geom_line(aes(x = date, y = cumul_growth_warm, color = "Warm resident")) +
  geom_line(aes(x = date, y = cumul_growth_cold, color = "Cold resident")) +
  geom_line(aes(x = date, y = cumul_growth_track, color = "Optimal mover"), linetype = "dashed") +
  scale_color_manual(values = c("Warm resident" = 2, "Cold resident" = 4, "Optimal mover" = 1)) +
  theme_bw() +
  xlab("Date") +
  ylab("Cumulative growth (g)")
```

Plot habitat use, i.e., temporal change in the habitat where growth potential is maximized. 
```{r}
growthregime_df %>% 
  mutate(habitat_num = as.numeric(as.factor(habitat))) %>%
  ggplot() +
  geom_line(aes(x = date, y = habitat_num)) +
  geom_point(aes(x = date, y = habitat_num, color = habitat)) +
  theme_bw() +
  scale_color_manual(values = c("cold" = 4, "warm" = 2)) +
  xlab("Date") +
  ylab("Habitat with max. GP (cold = 1, warm = 2)")
```

Contributions of warm vs. cold patches
```{r}
# time
growthregime_df %>% 
  group_by(habitat) %>%
  summarize(ndays = n()) %>%
  mutate(habitat = factor(habitat)) %>%
  ggplot() +
  geom_bar(aes(x = habitat, y = ndays/dim(growthregime_df)[1], fill = habitat), stat = "identity") +
  theme_bw() + theme(legend.position = "none") +
  scale_fill_manual(values = c("cold" = 4, "warm" = 2)) +
  ylab("Proportion of time spent in cold vs. warm habitats")

# cumulative growth accrued
growthregime_df %>% 
  group_by(habitat) %>%
  summarize(cumul_grpot = sum(grpot_track*10)) %>%
  mutate(habitat = factor(habitat)) %>%
  ggplot() +
  geom_bar(aes(x = habitat, y = cumul_grpot, fill = habitat), stat = "identity") +
  theme_bw() + theme(legend.position = "none") +
  scale_fill_manual(values = c("cold" = 4, "warm" = 2)) +
  ylab("Cumulative growth accrued in cold vs. warm habitats (g)")
```




## Realistic growth regime

Here, I simulate the growth regime where growth potential is an additional function of fish mass to generate more realistic size trajectories.

*Note that growth trajectories are super sensitive to ration size. Simulating growth under the highest ration size (0.17) generates very unrealistic growth outputs. In Armstrong et al. (2021), Fullerton's IBM simulations allowed ration to vary between 0.02 and 0.07 across river networks.*


### Equal ration

Ration size is equal between warm and cold habitat patches (0.10)
```{r}
# set up table
growthregime_df_2 <- habitat_df %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(1))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(2))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(3))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(4))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(5))) %>%
  mutate(grpot_warm = NA, grpot_cold = NA, weight_warm = NA, weight_cold = NA) %>%
  mutate(temp_warm = trimws(temp_warm), temp_cold = trimws(temp_cold)) #%>%
  # filter(date >= date("2024-06-01"))

# set starting weight
startweight <- 10

# habitat specific ration
ration_cold <- 0.10
ration_warm <- 0.10


for (i in 1:dim(growthregime_df_2)[1]) {
  if (i == 1) {
    grpot_warm <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_warm[i]),which(rations == ration_warm),startweight]
    grpot_cold <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_cold[i]),which(rations == ration_cold),startweight]
    weight_warm <- startweight + (startweight * grpot_warm)
    weight_cold <- startweight + (startweight * grpot_cold)
    
    growthregime_df_2$grpot_warm[i] <- grpot_warm
    growthregime_df_2$grpot_cold[i] <- grpot_cold
    growthregime_df_2$weight_warm[i] <- weight_warm
    growthregime_df_2$weight_cold[i] <- weight_cold
  } else {
    grpot_warm <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_warm[i]),which(rations == ration_warm),round(growthregime_df_2$weight_warm[i-1], digits = 0)]
    grpot_cold <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_cold[i]),which(rations == ration_cold),round(growthregime_df_2$weight_cold[i-1], digits = 0)]
    weight_warm <- growthregime_df_2$weight_warm[i-1] + (growthregime_df_2$weight_warm[i-1] * grpot_warm)
    weight_cold <- growthregime_df_2$weight_cold[i-1] + (growthregime_df_2$weight_cold[i-1] * grpot_cold)
    
    growthregime_df_2$grpot_warm[i] <- grpot_warm
    growthregime_df_2$grpot_cold[i] <- grpot_cold
    growthregime_df_2$weight_warm[i] <- weight_warm
    growthregime_df_2$weight_cold[i] <- weight_cold
  }
  # print(i)
}


p1 <- growthregime_df_2 %>% ggplot() +
  geom_line(aes(x = date, y = grpot_warm), col = "red") +
  geom_line(aes(x = date, y = grpot_cold), col = "blue") +
  theme_bw() +
  xlab("Date") +
  ylab("Growth potential (g/g/day)")

p2 <- growthregime_df_2 %>% ggplot() +
  geom_line(aes(x = date, y = weight_warm), col = "red") +
  geom_line(aes(x = date, y = weight_cold), col = "blue") +
  # geom_line(aes(x = date, y = cumul_grpot_track), col = "purple", linetype = "dashed") +
  theme_bw() +
  xlab("Date") +
  ylab("Cumulative growth potential (g)")

ggarrange(p1, p2)
```



### Unequal ration

Ration size is 30% lower in the cold habitat
```{r}
# set up table
growthregime_df_2 <- habitat_df %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(1))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(2))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(3))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(4))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(5))) %>%
  mutate(grpot_warm = NA, grpot_cold = NA, weight_warm = NA, weight_cold = NA) %>%
  mutate(temp_warm = trimws(temp_warm), temp_cold = trimws(temp_cold)) #%>%
  # filter(date >= date("2024-06-01"))

# set starting weight
startweight <- 10

# habitat specific ration
ration_cold <- 0.07
ration_warm <- 0.10


for (i in 1:dim(growthregime_df_2)[1]) {
  if (i == 1) {
    grpot_warm <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_warm[i]),which(rations == ration_warm),startweight]
    grpot_cold <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_cold[i]),which(rations == ration_cold),startweight]
    weight_warm <- startweight + (startweight * grpot_warm)
    weight_cold <- startweight + (startweight * grpot_cold)
    
    growthregime_df_2$grpot_warm[i] <- grpot_warm
    growthregime_df_2$grpot_cold[i] <- grpot_cold
    growthregime_df_2$weight_warm[i] <- weight_warm
    growthregime_df_2$weight_cold[i] <- weight_cold
  } else {
    grpot_warm <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_warm[i]),which(rations == ration_warm),round(growthregime_df_2$weight_warm[i-1], digits = 0)]
    grpot_cold <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_cold[i]),which(rations == ration_cold),round(growthregime_df_2$weight_cold[i-1], digits = 0)]
    weight_warm <- growthregime_df_2$weight_warm[i-1] + (growthregime_df_2$weight_warm[i-1] * grpot_warm)
    weight_cold <- growthregime_df_2$weight_cold[i-1] + (growthregime_df_2$weight_cold[i-1] * grpot_cold)
    
    growthregime_df_2$grpot_warm[i] <- grpot_warm
    growthregime_df_2$grpot_cold[i] <- grpot_cold
    growthregime_df_2$weight_warm[i] <- weight_warm
    growthregime_df_2$weight_cold[i] <- weight_cold
  }
  # print(i)
}


p1 <- growthregime_df_2 %>% ggplot() +
  geom_line(aes(x = date, y = grpot_warm), col = "red") +
  geom_line(aes(x = date, y = grpot_cold), col = "blue") +
  theme_bw() +
  xlab("Date") +
  ylab("Growth potential (g/g/day)")

p2 <- growthregime_df_2 %>% ggplot() +
  geom_line(aes(x = date, y = weight_warm), col = "red") +
  geom_line(aes(x = date, y = weight_cold), col = "blue") +
  # geom_line(aes(x = date, y = cumul_grpot_track), col = "purple", linetype = "dashed") +
  theme_bw() +
  xlab("Date") +
  ylab("Cumulative growth potential (g)")

ggarrange(p1, p2)
```








