---
title: "IBM Simulation"
execute:
  cache: refresh
---

```{r, include = FALSE}
library(tidyverse)
library(ggpubr)
library(egg)
library(RColorBrewer)
library(ggridges)
library(beepr)
```

**Purpose:** construct an individual based model to simulate how fish move to track temperature-based growth opportunities across a habitat network, the resulting effects on growth, and the implications for long-term population dynamics.


Source dependencies
```{r include = FALSE}
#| cache: false
source("Habitat.R")
source("Functions.R")
source("sim_functions.R")
```

Load pre-calculated growth
```{r}
#| cache: false
load("data/wt.growth.array.RData")
```


## Design notes

### Key modules

* **Patch choice (movement)**
  + Optimal movers select habitat patches (move) based on sensed differences in growth potential. Fish compare growth potential in their current habitat to that in the other habitat, where sensed growth potential is either a function of temperature alone or temperature and density (see *Configuration options*). 
  + For all habitat sensing options, growth potential in the alternative habitat (i.e., the growth benefit of moving) is penalized by an energetic cost of movement, which is an allometric function of size (relative growth penalty in g/g/d declines with increasing fish weight).
Finally, the decision to move based on differences in (movement penalized) growth potential follows a probabilistic (softmax) decision rule, where fish switch with a probability that increases with the sensed growth advantage. 

* **Growth**
  + After all fish move, fish densities in each patch are calculated, and the temperature adjusted P_Cmax (physiological capacity for consumption given temperature) is penalized by a density-dependent hyperbolic reduction function. 
  + A Wisconsin bioenergetics model is used to calculate fish growth based on fish weight, temperature, and density (P_Cmax). 
  + All fish sizes are update accordingly.

* **Spawning and reproduction**
  + The decision to spawn (i.e., maturation probability) is function of day of year (potential spawning dates are normally distributed around a mean date), fish size (spawning probability increases with fish weight following a logistic function), and fish condition (probability increases with condition, calculated as current weight / peak weight). 
  + Fish can only spawn once per year
  + When a fish spawns, they experience an energetic cost that is calculated by multiple current body weight by a fixed proportion (e.g., 0.2). In other words, fish lose 20% of their weight if all spawning conditions are met. 
  + Fecundity (number of eggs per spawner) is an increasing function of weight (log-log linear relationship) with log-scale noise. 
  + All eggs are subject to a fixed egg-to-fry survival rate, e.g., 0.1. 
  + Spawning and reproduction occurs instantaneously, in the same time step. I.e., if spawning conditions are met on some day, their offspring are added to the simulation on that same day. 
  + Offspring inherit their parents patch and life history strategy (e.g., optimal mover or resident)

* **Survival**
  + All fish (including new offspring) are subject to weight, temperature, and condition-based mortality sources
    - Weight: daily survival probability increases with fish weight, up to some maximum probability (0.9994) that is tuned to yield an annual survival probability of ~0.8. 
		- High temperature exposure: this mortality source represents the acute breakdown of physiological processes at high temperatures. It does not represent chronic effects of high temperatures on bioenergetics. Daily survival probability decreases above some critical high temperature, following a logistic function.
		- Condition/starvation: fish in poor condition (current weight / peak weight) experience reduced daily survival probability, following a logistic function. 
  + Daily survival probability is calculated as the joint probability of the sources described above, and daily survival is drawn from a binomial distribution. Fish that die are removed from the simulation, but their results are stored. 


### Configuration options

* **Stochasticity in movement** is accounted for by one of three alternative mechanisms:
  + *No stochastic movement/entirely deterministic:* all movers do the same thing...they all can perfectly sense and compare growth opportunities
  + *Probabilistic (softmax) rule:* the softmax function assigns probabilities to each patch based on the growth difference, with higher probabilities for patches offering greater growth benefits.
  + *Individual-level movement threshold:* each fish has a unique threshold for movement, representing individual variation in boldness/dispersal propensity
* **Density-dependent food availaility:** after fish have selected patches, calculate food availability based on patch density and baseline ration levels. Two alternative approaches:
  + *Modified Fullerton approach:* ration size decreases linearly with fish density, up to some (*a priori* defined) threshold above which there is no effect of there is no effect of density of ration size.
  + *Hyperbolic reduction:* describes a non-linear decrease in ration size with increasing fish density, where the rate of decrease slows as density increases.
* **Environment sensing:** how omniscient are fish to temperature and density-dependent growth across the habitat patches?
  + `abiotic_only`: fish can only sense temperature-dependent GP, regardless of which habitat is occupied
  + `density_current`: fish sense density-dependent GP in current habitat, and compare to abiotic (temp-based) GP in the alternative habitat
  + `density_all`: fish can sense density-dependent GP in all habitats (this is how Railsback does it: fish can sense habitat and competition conditions in all cells within a specific radius)


### Senstive parameters

*under construction*

What parameters do we think might have very strong effects on model outcomes?

* Habitat variables (temperature regimes and maximum P_C~max~) 
* Strength of density-dependence and differences between habitats ()
* Egg-to-fry survival rate (could mediate the trade-off between fish size and number)
* Strength of size-based dominance ()

Other parameters:

* General
  + Habitat variables (temperature regimes and food availability in each patch) 
  + Number of fish per strategy/total number of starting fish
* Density-dependence
  + Fullerton upper density threshold
  + Hyperbolic half saturation density
* Movement decision
  + Probabilistic/softmax tau (strength of threshold)
* Movement cost
  + Allometric function b (metabolic exponent) and c (scaling constant)
* Movement stochasticity
  + Probabilistic/softmax tau (sensitivity to growth difference)
  + Individual-level movement threshold, SD in individual boldness
  
  
### Misc. notes

* Currently, we just look-up growth from the precalculated growth array, but this is going to be less accurate for the IBM b/c precalculated growth rates are calculated over 0.25g mass increments, whereas growth in the IBM is continuous. Probably fine for now, but perhaps should borrow code from above/Fullerton IBM to increase accuracy. My guess is that this will dramatically increase run times. 
* If `sense_environment == "density_all"`, the density used for the alternative patch is the current pre-movement density (i.e., what the fish can observe before anyone decides to move that day). This is the most tractable approximation — computing post-movement density would require solving a joint decision problem. 
* If using individual-level movement thresholds (i.e., boldness), residents can be given `move_threshold = Inf` to formally unify all three strategies under one mechanism, rather than branching by strategy label — keeps the code cleaner if we expand to more strategies later.


### To do

* eventually, should probably collate all parameters into a single control script/page...that way we can see where all the sensitive/arbitrary parameters are
* stochasticity in movement: it probably makes sense to combine stochastic movement processes...e.g, probabilistic and individual thresholds, and perhaps even random noise as well
* Size-at-maturity could probably be sampled probabilistically, with noise. Preliminary simulation will just set a specific size threshold. 
* Same for timing of reproduction: rather than all mature fish reproducing on a single day, we could have them spread out over a given period, or make spawn timing size-dependent, with larger fish spawning first. Lots of empirical evidence for this. That would be very interesting because it would give the offspring of the larger/early spawners a competitive advantage over offspring of smaller/later spawners, likely magnifying the effect of size on reproductive success. 


## Set-up

Three movement strategies are simulated: fish that remain year-round in the warm patch (`resident_warm`), fish that remain year-round in the cold patch (`resident_cold`), and fish that choose the better-growing patch each day (`optimal_mover`). Ration sizes match the unequal-ration scenario above.

**Configuration notes:**

* All fish start at same size, no variation
* If relevant, variation in boldness parameter probably needs tuning


Habitat parameters:
```{r}
hab_params <- list(
  # Timeline
  nyears_burnin     = 20,
  nyears_experiment = 20,
  mindate           = as.Date("2001-05-01"),

  # Temperature (warm patch sine wave; cold = warm × cold_multiplier)
  base_temp       = 11,    # mean annual temperature (°C)
  amplitude       = 14,    # seasonal amplitude (°C)
  temp_noise_sd   = 0,     # daily noise SD (0 = deterministic)
  cold_multiplier = 0.6,   # cold patch as fraction of warm

  # Patch burn-in values
  pcmax_warm = 0.5,
  pcmax_cold = 0.5,
  A_warm     = 1.0,
  A_cold     = 1.0,
  K_warm     = 500,
  K_cold     = 500,
  S_max_warm = 0.9994,
  S_max_cold = 0.9994,

  # Experiment targets (NA = no change; ramp_years = 0 for step change)
  pcmax_warm_target = NA,
  pcmax_cold_target = NA,
  A_warm_target     = 2,
  A_cold_target     = 0,
  K_warm_target     = NA,
  K_cold_target     = NA,
  S_max_warm_target = NA,
  S_max_cold_target = NA,
  ramp_years        = 20,

  hab_seed = 123
)

habitat_df <- build_habitat(hab_params)
```


Simulation parameters:
```{r}
sim_params <- list(
  # Behavioural / movement options
  move_stochastic   = "prob",             # "none" | "prob" | "indiv"
  food_densdepen    = "hyperbolic",       # "fullerton" | "hyperbolic"
  sense_environment = "density_current",  # "abiotic_only" | "density_current" | "density_all"

  # Population initialisation
  n_fish_per_strat  = 50,    # fish per strategy
  start_wt          = 0.5,   # mean initial weight (g)
  wt_sd             = 0.05,  # SD of initial weight

  # Movement costs and stochasticity
  movecost_c   = 0.1,    # allometric cost scaling constant
  movecost_b   = 0.3,    # allometric cost decay exponent
  sigma_bold   = 0.002,  # SD of individual movement threshold (g/g/d)
  tau          = 0.001,  # softmax sensitivity to growth difference

  # Survival
  T1_mort           = 30,      # temp (°C) at which thermal daily survival = 0.1
  T9_mort           = 25.8,    # temp (°C) at which thermal daily survival = 0.9
  K9_starv          = 0.55,    # condition at which starvation survival = 0.9
  K1_starv          = 0.45,    # condition at which starvation survival = 0.1
  MaxDensity4Growth = 50,      # Fullerton density cap (fullerton DD only)

  # Competition
  dominance_beta = 1,  # 0 = scramble, 1 = linear dominance, Inf = pure contest

  # Reproduction
  egg_wt       = 0.07,   # weight of a single egg (g)
  repro_cost   = 0.2,    # energetic cost of spawning as fraction of body weight
  sigma_fecund = 0.085,  # log-scale noise in weight-fecundity relationship
  egg_surv     = 0.1,    # egg-to-fry survival probability

  sim_seed = 7843
)
```




## Simulation loop

For each day of year, optimal movers compare the bioenergetics growth rate at both patches — the vectorised equivalent of calling `fncGrowthPossible()` for each fish — and switch only when the movement-cost-penalised gain in the alternative patch exceeds staying (but this may differ among individuals depending on how movement stochasticity is configured). Warm and cold patch residents stay in their pre-defined patch. `fncGrowthFish()` then looks up the actual growth rate for every fish at their current patch, weight, and ration, and weights are updated accordingly.

```{r}
st <- Sys.time()

results <- run_simulation(habitat_df, sim_params, wt.growth)

Sys.time() - st

beep()
```



## Summarize results

Collate results into long tibbles and summarized tibbles
```{r eval=FALSE}
# assign DOY column names
colnames(mass_matrix)     <- as.character(habitat_df$dayofsim)
colnames(patch_matrix)    <- as.character(habitat_df$dayofsim)
colnames(temp_matrix)     <- as.character(habitat_df$dayofsim)
colnames(ggd_matrix)      <- as.character(habitat_df$dayofsim)
colnames(surv_matrix)     <- as.character(habitat_df$dayofsim)
colnames(age_matrix)      <- as.character(habitat_df$dayofsim)
colnames(condition_matrix)<- as.character(habitat_df$dayofsim)

# colnames(pcmax_base_matrix)    <- as.character(habitat_df$dayofsim)
# colnames(pcmax_adj_matrix)     <- as.character(habitat_df$dayofsim)
# colnames(pcmax_adjdd_matrix)   <- as.character(habitat_df$dayofsim)
# colnames(func_temp_matrix)     <- as.character(habitat_df$dayofsim)
# colnames(ration_matrix)        <- as.character(habitat_df$dayofsim)


# Convert weight matrix to long format, join with patch matrix
# Use fish_registry to label rows — covers founders + all offspring born during simulation
mass_long <- as_tibble(mass_matrix) %>%
  bind_cols(fish_registry %>% select(pid, strategy)) %>%
  pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "weight") %>%
  filter(!is.na(weight)) %>%
  mutate(dayofsim = as.integer(dayofsim)) %>%
  left_join(
    as_tibble(patch_matrix) %>%
      bind_cols(fish_registry %>% select(pid)) %>%
      pivot_longer(-pid, names_to = "dayofsim", values_to = "patch") %>%
      mutate(dayofsim = as.integer(dayofsim)),
    by = c("pid", "dayofsim")
  ) %>%
  left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")

# Convert temp matrix to long format
temp_long <- as_tibble(temp_matrix) %>%
  bind_cols(fish_registry %>% select(pid, strategy)) %>%
  pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "temp") %>%
  filter(!is.na(temp)) %>%
  mutate(dayofsim = as.integer(dayofsim)) %>%
  left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")

# convert standardized growth rate matrix to long format
ggd_long <- as_tibble(ggd_matrix) %>%
  bind_cols(fish_registry %>% select(pid, strategy)) %>%
  pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "ggd") %>%
  filter(!is.na(ggd)) %>%
  mutate(dayofsim = as.integer(dayofsim)) %>%
  left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")

# convert survival matrix to long format (1 = survived, 0 = died, NA = already dead)
surv_long <- as_tibble(surv_matrix) %>%
  bind_cols(fish_registry %>% select(pid, strategy)) %>%
  pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "survived") %>%
  filter(!is.na(survived)) %>%
  mutate(dayofsim = as.integer(dayofsim)) %>%
  left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")

# # convert baseline P_Cmax matrix to long format
# pcmax_base_long <- as_tibble(pcmax_base_matrix) %>%
#   bind_cols(fish_registry %>% select(pid, strategy)) %>%
#   pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "pcmax_base") %>%
#   mutate(dayofsim = as.integer(dayofsim)) %>%
#   left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")
# 
# # convert temp-adjusted P_Cmax matrix to long format
# pcmax_adj_long <- as_tibble(pcmax_adj_matrix) %>%
#   bind_cols(fish_registry %>% select(pid, strategy)) %>%
#   pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "pcmax_adj") %>%
#   mutate(dayofsim = as.integer(dayofsim)) %>%
#   left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")
# 
# # convert temp and density adjusted matrix to long format
# pcmax_adjdd_long <- as_tibble(pcmax_adjdd_matrix) %>%
#   bind_cols(fish_registry %>% select(pid, strategy)) %>%
#   pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "pcmax_adjdd") %>%
#   mutate(dayofsim = as.integer(dayofsim)) %>%
#   left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")

# convert temperature function matrix to long format
# func_temp_long <- as_tibble(func_temp_matrix) %>%
#   bind_cols(fish_registry %>% select(pid, strategy)) %>%
#   pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "func_temp") %>%
#   mutate(dayofsim = as.integer(dayofsim)) %>%
#   left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")

# # convert ration matrix to long format
# ration_long <- as_tibble(ration_matrix) %>%
#   bind_cols(fish_registry %>% select(pid, strategy)) %>%
#   pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "ration") %>%
#   mutate(dayofsim = as.integer(dayofsim)) %>%
#   left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")

# convert condition matrix to long
condition_long <- as_tibble(condition_matrix) %>%
  bind_cols(fish_registry %>% select(pid, strategy)) %>%
  pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "condition") %>%
  filter(!is.na(condition)) %>%
  mutate(dayofsim = as.integer(dayofsim)) %>%
  left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")

# convert age matrix to long
age_long <- as_tibble(age_matrix) %>%
  bind_cols(fish_registry %>% select(pid, strategy)) %>%
  pivot_longer(-c(pid, strategy), names_to = "dayofsim", values_to = "age") %>%
  filter(!is.na(age)) %>%
mutate(dayofsim = as.integer(dayofsim)) %>%
  left_join(habitat_df %>% select(dayofsim, date), by = "dayofsim")

# bind all individual data
ibm_long <- mass_long %>%
  left_join(temp_long %>% select(-date), by = c("pid", "strategy", "dayofsim")) %>%
  left_join(ggd_long %>% select(-date),  by = c("pid", "strategy", "dayofsim")) %>%
  left_join(surv_long %>% select(-date), by = c("pid", "strategy", "dayofsim")) %>%
  # left_join(pcmax_base_long %>% select(-date), by = c("pid", "strategy", "dayofsim")) %>%
  # left_join(pcmax_adj_long %>% select(-date), by = c("pid", "strategy", "dayofsim")) %>%
  # left_join(pcmax_adjdd_long %>% select(-date), by = c("pid", "strategy", "dayofsim")) %>%
  # left_join(func_temp_long %>% select(-date), by = c("pid", "strategy", "dayofsim")) %>%
  # left_join(ration_long %>% select(-date), by = c("pid", "strategy", "dayofsim")) %>%
  left_join(condition_long %>% select(-date), by = c("pid", "strategy", "dayofsim")) %>%
  left_join(age_long %>% select(-date), by = c("pid", "strategy", "dayofsim")) %>%
  left_join(fish_registry)

# summarize by strategy
ibm_summary <- ibm_long %>%
  group_by(strategy, dayofsim, date) %>%
  summarise(
    mean_weight      = mean(weight, na.rm = TRUE),
    sd_weight        = sd(weight,   na.rm = TRUE),
    mean_temp        = mean(temp,   na.rm = TRUE),
    sd_temp          = sd(temp,     na.rm = TRUE),
    mean_ggd         = mean(ggd,    na.rm = TRUE),
    sd_ggd           = sd(ggd,      na.rm = TRUE),
    prop_warm        = mean(patch == "warm", na.rm = TRUE),
    n_alive          = sum(survived, na.rm = TRUE),
    # mean_pcmax_bas   = mean(pcmax_base, na.rm = TRUE),
    # sd_pcmax_base    = sd(pcmax_base, na.rm = TRUE),
    # mean_pcmax_adj   = mean(pcmax_adj, na.rm = TRUE),
    # sd_pcmax_adj     = sd(pcmax_adj, na.rm = TRUE),
    # mean_pcmax_adjdd = mean(pcmax_adjdd, na.rm = TRUE),
    # sd_pcmax_adjdd   = sd(pcmax_adjdd, na.rm = TRUE),
    # mean_func_temp   = mean(func_temp, na.rm = TRUE),
    # sd_func_temp     = sd(func_temp, na.rm = TRUE),
    # mean_ration      = mean(ration, na.rm = TRUE),
    # sd_ration        = sd(ration, na.rm = TRUE),
    mean_condition   = mean(condition, na.rm = TRUE),
    sd_condition     = sd(condition, na.rm = TRUE),
    # mean_age         = mean(age, na.rm = TRUE),
    # sd_age           = sd(age, na.rm = TRUE),
    .groups          = "drop"
  )

ibm_summary
```

Collate results into a single long tibble
```{r eval=FALSE}
valid_idx <- which(!is.na(mass_matrix), arr.ind = TRUE)

ibm_long <- tibble(
  row_i     = valid_idx[, 1],
  col_i     = valid_idx[, 2],
  dayofsim  = as.integer(colnames(mass_matrix))[valid_idx[, 2]],
  weight    = mass_matrix[valid_idx],
  temp      = temp_matrix[valid_idx],
  ggd       = ggd_matrix[valid_idx],
  survived  = surv_matrix[valid_idx],
  condition = condition_matrix[valid_idx],
  age       = age_matrix[valid_idx],
  patch     = patch_matrix[valid_idx]
) |>
  mutate(
    pid      = fish_registry$pid[row_i],
    strategy = fish_registry$strategy[row_i]
  ) |>
  select(-row_i, -col_i) |>
  left_join(fish_registry, by = c("pid", "strategy")) |>
  left_join(habitat_df |> select(dayofsim, date), by = "dayofsim")

gc()
```

Compute summary tibble directly from the matrices (not really relevant anymore given that we have multiple cohorts)
```{r eval=FALSE}
strat_idx <- split(seq_len(nrow(mass_matrix)), fish_registry$strategy)

ibm_summary <- lapply(names(strat_idx), function(s) {
  i <- strat_idx[[s]]
  tibble(
    strategy       = s,
    dayofsim       = as.integer(colnames(mass_matrix)),
    mean_weight    = colMeans(mass_matrix[i, , drop = FALSE], na.rm = TRUE),
    sd_weight      = apply(mass_matrix[i, , drop = FALSE], 2, sd, na.rm = TRUE),
    mean_temp      = colMeans(temp_matrix[i, , drop = FALSE], na.rm = TRUE),
    sd_temp        = apply(temp_matrix[i, , drop = FALSE], 2, sd, na.rm = TRUE),
    mean_ggd       = colMeans(ggd_matrix[i, , drop = FALSE], na.rm = TRUE),
    sd_ggd         = apply(ggd_matrix[i, , drop = FALSE], 2, sd, na.rm = TRUE),
    mean_condition = colMeans(condition_matrix[i, , drop = FALSE], na.rm = TRUE),
    sd_condition   = apply(condition_matrix[i, , drop = FALSE], 2, sd, na.rm = TRUE),
    prop_warm      = colMeans(patch_matrix[i, , drop = FALSE] == "warm", na.rm = TRUE),
    n_alive        = colSums(!is.na(mass_matrix[i, , drop = FALSE]))
  )
}) |>
  bind_rows() |>
  left_join(habitat_df |> select(dayofsim, date), by = "dayofsim")
```


Extract results:
```{r}
#| cache: false
ibm_records   <- results$ibm_records
ibm_long      <- results$ibm_long
ibm_summary   <- results$ibm_summary
fish_registry <- results$fish_registry
spawn_log     <- results$spawn_log
switches      <- results$switches
habitat_df    <- results$habitat_df
```


## Mechanistic results

### Bioenergetics

Calculate bioenergetics outputs post-hoc (same as calculated in the loop, but storage limitations necessitate computing externally).
```{r}
bioe_output <- ibm_long %>% 
  filter(survived == 1) %>%
  left_join(habitat_df |> select(dayofsim, pcmax_warm, pcmax_cold, K_warm, K_cold), by = "dayofsim") %>%
  group_by(dayofsim, patch) %>% 
  mutate(density = n()) %>%
  ungroup() %>%
  mutate(func_temp          = fncTempDepend(temp),
         pcmax_baseline     = ifelse(patch == "warm", pcmax_warm, pcmax_cold),
         pcmax_adjusted     = pmin(pcmax_baseline, func_temp),
         cmax_allometric    = fncAllomCmax(weight),
         ration             = cmax_allometric * pcmax_adjusted,
         k                  = if_else(patch == "warm", K_warm, K_cold),
         pcmax_adjusted_dd  = pcmax_adjusted * (k / (k + density - 1)))
```


Ration size by weight: points should generally track allometric scaling of C~max~ (proportionally reduced to account for baseline/treshold P_Cmax, *solid line*). Noise is due to variation in temperature, density, and body size. This looks good: points fall away from baseline during very cold periods when temperature restricts physiological capacity. The distinct "cliffs" show that you get the same behavior during the winter when temps are limited but fish have attained different mean body sizes depending on their life-history strategy
```{r}
# calculate allometric scaling of Cmax
constants <- fncReadConstants() 
xs <- seq(from = min(ibm_long$weight, na.rm = TRUE), to = max(ibm_long$weight, na.rm = TRUE), length.out = 100)
ys <- constants$Consumption$CA * (xs ^ constants$Consumption$CB)

# get ceiling P_cmax (day 1 values, used as reference lines on plot)
pcmax_cold <- habitat_df$pcmax_cold[1]
pcmax_warm <- habitat_df$pcmax_warm[1]

# plot
p1 <- ggplot() +
  geom_point(data = bioe_output %>% filter(patch == "warm"), aes(x = weight, y = ration, color = temp)) +
  geom_line(aes(x = xs, y = ys*pcmax_warm), linetype = "solid", color = "black") +
  theme_bw() + theme(legend.position = "top") + scale_color_gradient(low = "blue", high = "red") +
  ylab("Ration size (g/g/d)") + xlab("Fish mass (g)") + labs(title = "Warm patch")

p2 <- ggplot() +
  geom_point(data = bioe_output %>% filter(patch == "cold"), aes(x = weight, y = ration, color = temp)) +
  geom_line(aes(x = xs, y = ys*pcmax_cold), linetype = "solid", color = "black") +
  theme_bw() + theme(legend.position = "top") + scale_color_gradient(low = "blue", high = "red") +
  ylab("Ration size (g/g/d)") + xlab("Fish mass (g)") + labs(title = "Cold patch")

ggpubr::ggarrange(p2, p1, nrow = 1, common.legend = TRUE, legend = "right")
```

f(T) should perfectly track the temperature dependence function b/c it is only based on temperature and constants that are common to all individuals
```{r}
ggplot() +
  geom_point(data = bioe_output, aes(x = temp, y = func_temp), alpha = 0.3) +
  theme_bw() + 
  ylab("f(Temperature)") + xlab("Temperature (deg. C)")
```


P_Cmax by temp: points should perfect track the temperature dependence function, but max out at the threshold P_Cmax set in each habitat.
```{r}
p1 <- ggplot() +
  geom_point(data = bioe_output %>% filter(patch == "cold"), aes(x = temp, y = pcmax_adjusted)) +
  geom_abline(slope = 0, intercept = pcmax_cold, color = "red", linetype = "dashed") +
  theme_bw() + 
  ylab("Temperature adjusted P_Cmax") + xlab("Temperature (deg. C)") + labs(title = "Cold patch")

p2 <- ggplot() +
  geom_point(data = bioe_output %>% filter(patch == "warm"), aes(x = temp, y = pcmax_adjusted)) +
  geom_abline(slope = 0, intercept = pcmax_warm, color = "red", linetype = "dashed") +
  theme_bw() + 
  ylab("Temperature adjusted P_Cmax") + xlab("Temperature (deg. C)") + labs(title = "Warm patch")

ggpubr::ggarrange(p2, p1, nrow = 1)
```


Now show P_cmax adjusted by both temperature and density: points should generally track temperature-dependence function, capped at baseline/threshold P_Cmax. Noise is due to variation in body size and density. Smaller fish feeding at higher densities typically experience more food limitation...this appears to occur more commonly in the warm habitat during late spring.  
```{r}
p1 <- ggplot() +
  geom_point(data = bioe_output %>% filter(patch == "cold"), aes(x = temp, y = pcmax_adjusted_dd), alpha = 0.01) +
  geom_line(aes(x = waterTemps[,2], y = fncTempDepend(waterTemps[,2])), color = "red") +
  geom_abline(slope = 0, intercept = pcmax_cold, color = "red", linetype = "dashed") +
  theme_bw() + 
  ylab("Temperature and density adjusted P_Cmax") + xlab("Temperature (deg. C)") + labs(title = "Cold patch")

p2 <- ggplot() +
  geom_point(data = bioe_output %>% filter(patch == "warm"), aes(x = temp, y = pcmax_adjusted_dd), alpha = 0.01) +
  geom_line(aes(x = waterTemps[,2], y = fncTempDepend(waterTemps[,2])), color = "red") +
  geom_abline(slope = 0, intercept = pcmax_warm, color = "red", linetype = "dashed") +
  theme_bw() + 
  ylab("Temperature and density adjusted P_Cmax") + xlab("Temperature (deg. C)") + labs(title = "Warm patch")

ggpubr::ggarrange(p1, p2, nrow = 1)
```

Show time series of density-adjustec Pcmax for 10 randomly selected individuals
```{r}
bioe_output %>%
  filter(pid %in% sample(unique(bioe_output$pid), size = 10, replace = FALSE)) %>%
  ggplot() +
  geom_line(aes(x = date, y = pcmax_adjusted_dd, group = pid)) + facet_wrap(~pid, scales = "free_x")
```



### Temperature

Temperature experienced by optimal movers
```{r}
# p1 <- ibm_summary %>%
#   filter(strategy == "optimal_mover") %>%
#   ggplot(aes(x = date, color = strategy, fill = strategy)) +
#   geom_ribbon(aes(ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp),
#               alpha = 0.15, color = NA) +
#   geom_line(aes(y = mean_temp)) +
#   scale_color_manual(values = c(resident_warm = "red", 
#                                 resident_cold = "blue", 
#                                 optimal_mover = "darkgreen")) +
#   scale_fill_manual(values  = c(resident_warm = "red", 
#                                 resident_cold = "blue",
#                                 optimal_mover = "darkgreen")) +
#   theme_bw() + #ylim(0,10000) +
#   xlab("Date") + ylab("Mean experienced temperature (deg. C)")


ibm_long %>%
  filter(strategy == "optimal_mover") %>%
  ggplot(aes(x = date, y = temp, color = strategy, group = pid)) +
  geom_line(alpha = 0.5) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  theme_bw() + #ylim(0,10000) +
  xlab("Date") + ylab("Individual experienced temperature (deg. C)")

ibm_long %>%
  filter(strategy == "optimal_mover", year(date) == 2020) %>%
  ggplot(aes(x = date, y = temp, color = strategy, group = pid)) +
  geom_line(alpha = 0.5) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  theme_bw() + #ylim(0,10000) +
  xlab("Date") + ylab("Individual experienced temperature (deg. C)")

# ggpubr::ggarrange(p1, p2, nrow = 1, common.legend = TRUE)
```


### Growth rate

Experienced standardized growth rates over time, by strategy.
```{r}
# p1 <- ibm_summary %>%
#   # filter(strategy == "optimal_mover") %>%
#   ggplot(aes(x = date, color = strategy, fill = strategy)) +
#   geom_ribbon(aes(ymin = mean_ggd - sd_ggd, ymax = mean_ggd + sd_ggd),
#               alpha = 0.15, color = NA) +
#   geom_line(aes(y = mean_ggd)) +
#   scale_color_manual(values = c(resident_warm = "red", 
#                                 resident_cold = "blue", 
#                                 optimal_mover = "darkgreen")) +
#   scale_fill_manual(values  = c(resident_warm = "red", 
#                                 resident_cold = "blue",
#                                 optimal_mover = "darkgreen")) +
#   theme_bw() + #ylim(0,10000) +
#   geom_abline(slope = 0, intercept = 0, linetype = "dashed") +
#   xlab("Date") + ylab("Mean growth rate (g/g/d)")

ibm_long %>%
  ggplot(aes(x = date, y = ggd, color = strategy, group = pid)) +
  geom_line(alpha = 0.5) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  theme_bw() + #ylim(0,10000) +
  xlab("Date") + ylab("Individual growth rate (g/g/d)")

ibm_long %>%
  filter(year(date) == 2020) %>%
  mutate(age = as.factor(round(age))) %>%
  ggplot(aes(x = date, y = ggd, color = age, group = age)) +
  geom_line(alpha = 0.5) +
  # scale_color_manual(values = c(resident_warm = "red", 
  #                               resident_cold = "blue", 
  #                               optimal_mover = "darkgreen")) +
  theme_bw() + #ylim(0,10000) +
  xlab("Date") + ylab("Individual growth rate (g/g/d)")

# ggpubr::ggarrange(p1, p2, nrow = 1, common.legend = TRUE)
```


### Size trajectories

Weight trajectories by strategy.

*I wonder how realistic overwinter mass losses are...*
```{r}
# p1 <- ibm_summary %>%
#   ggplot() +
#   geom_rect(data = summer_shading,
#             aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
#             fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
#   geom_ribbon(aes(x = date, color = strategy, fill = strategy, ymin = mean_weight - sd_weight, ymax = mean_weight + sd_weight),
#               alpha = 0.15, color = NA) +
#   geom_line(aes(x = date, color = strategy, fill = strategy, y = mean_weight)) +
#   scale_color_manual(values = c(resident_warm = "red", 
#                                 resident_cold = "blue", 
#                                 optimal_mover = "darkgreen")) +
#   scale_fill_manual(values  = c(resident_warm = "red", 
#                                 resident_cold = "blue",
#                                 optimal_mover = "darkgreen")) +
#   theme_bw() + theme(panel.grid = element_blank()) +
#   xlab("Date") + ylab("Mean weight ± SD (g)")

ibm_long %>%
  ggplot() +
  # geom_rect(data = summer_shading,
  #           aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  #           fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
  # geom_ribbon(aes(x = age, color = strategy, fill = strategy, ymin = mean_weight - sd_weight, ymax = mean_weight + sd_weight),
  #             alpha = 0.15, color = NA) +
  geom_line(aes(x = age, y = weight, color = strategy, group = pid), alpha = 0.5) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  scale_fill_manual(values  = c(resident_warm = "red", 
                                resident_cold = "blue",
                                optimal_mover = "darkgreen")) +
  theme_bw() + theme(panel.grid = element_blank()) +
  xlab("Age") + ylab("Individual weight (g)")

ibm_long %>%
  ggplot() +
  geom_rect(data = summer_shading,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
  geom_line(aes(x = date, y = weight, color = strategy, group = pid), alpha = 0.5) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  theme_bw() + theme(panel.grid = element_blank()) +
  xlab("Date") + ylab("Individual weight (g)")

# ggpubr::ggarrange(p1, p2, nrow = 1, common.legend = TRUE)
```

Now zoom into the beginning of the simulation, to look growth trajectories of early recruits
```{r}
# p1 <- ibm_summary %>%
#   filter(date <= min(date) + days(100)) %>%
#   ggplot() +
#   geom_ribbon(aes(x = date, color = strategy, fill = strategy, ymin = mean_weight - sd_weight, ymax = mean_weight + sd_weight),
#               alpha = 0.15, color = NA) +
#   geom_line(aes(x = date, color = strategy, fill = strategy, y = mean_weight)) +
#   scale_color_manual(values = c(resident_warm = "red", 
#                                 resident_cold = "blue", 
#                                 optimal_mover = "darkgreen")) +
#   scale_fill_manual(values  = c(resident_warm = "red", 
#                                 resident_cold = "blue",
#                                 optimal_mover = "darkgreen")) +
#   theme_bw() + theme(panel.grid = element_blank()) +
#   xlab("Date") + ylab("Mean weight ± SD (g)") +
#   geom_abline(slope = 0, intercept = 0, linetype = "dashed")

ibm_long %>%
  filter(date <= min(date) + days(100)) %>%
  ggplot() +
  geom_line(aes(x = date, y = weight, color = strategy, group = pid), alpha = 0.5) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  theme_bw() + theme(panel.grid = element_blank()) +
  xlab("Date") + ylab("Individual weight (g)") +
  geom_abline(slope = 0, intercept = 0, linetype = "dashed")

# ggpubr::ggarrange(p1, p2, nrow = 1, common.legend = TRUE)
```

Now just look at warm residents. Good, these fish are dying as expected, driven by small body sizes and summer starvation.
```{r}
# p1 <- ibm_summary %>%
#   filter(strategy == "resident_warm", !is.na(mean_weight)) %>%
#   ggplot() +
#   # geom_rect(data = summer_shading,
#   #           aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
#   #           fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
#   geom_ribbon(aes(x = date, color = strategy, fill = strategy, ymin = mean_weight - sd_weight, ymax = mean_weight + sd_weight),
#               alpha = 0.15, color = NA) +
#   geom_line(aes(x = date, color = strategy, fill = strategy, y = mean_weight)) +
#   scale_color_manual(values = c(resident_warm = "red", 
#                                 resident_cold = "blue", 
#                                 optimal_mover = "darkgreen")) +
#   scale_fill_manual(values  = c(resident_warm = "red", 
#                                 resident_cold = "blue",
#                                 optimal_mover = "darkgreen")) +
#   theme_bw() + theme(panel.grid = element_blank()) +
#   xlab("Date") + ylab("Mean weight ± SD (g)")

ibm_long %>%
  filter(strategy == "resident_warm", !is.na(weight)) %>%
  ggplot() +
  # geom_rect(data = summer_shading,
  #           aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  #           fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
  geom_line(aes(x = date, y = weight, color = strategy, group = pid), alpha = 0.5) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  theme_bw() + theme(panel.grid = element_blank()) +
  xlab("Date") + ylab("Individual weight (g)")

# ggpubr::ggarrange(p1, p2, nrow = 1, common.legend = TRUE)
```


### Maximum age

Plot histograms of maximum age by strategy. Want to simply make sure that fish are dying at a reasonable time. Most fish won't live much past age-1, but fish shouldn't exceed 10 years or so. 
```{r}
max_age_df <- ibm_long |>
  dplyr::filter(survived == 1) |>
  dplyr::group_by(strategy, pid) |>
  dplyr::summarize(max_age = max(age), .groups = "drop") 

strategy_colors <- c(resident_warm = "red", resident_cold = "blue", optimal_mover = "darkgreen")

ggplot(max_age_df, aes(x = max_age, fill = strategy)) +
  geom_histogram(binwidth = 0.25, color = "white", linewidth = 0.2) +
  scale_fill_manual(values = strategy_colors) +
  facet_wrap(~ strategy, ncol = 1, scales = "free_y") +
  theme_bw() + theme(panel.grid = element_blank(), legend.position = "none") +
  xlab("Maximum age (years)") + ylab("Number of fish") +
  labs(title = "Maximum age by strategy")

max_age_df %>%
  mutate(max_age = as.factor(round(max_age))) %>%
  group_by(max_age) %>% summarize(N = n())
```


### Survival

Look at how the mechanisms of survival manifest over time, for each strategy and cohort.
```{r fig.width=10, fig.height=7}
# Build survival probability components for all fish across all alive days
all_surv_df <- ibm_long |>
  mutate(
    p_temp  = fncSurviveTemp(temp, T1 = T1_mort, T9 = T9_mort),
    p_starv = fncSurviveStarve(condition, K9 = K9_starv, K1 = K1_starv),
    p_sg    = fncSurviveSize(weight, maxprob = ifelse(patch == "warm", S_max_warm, S_max_cold))[[1]]
  ) |>
  # left_join(habitat_df |> select(dayofsim, date), by = "dayofsim") |>
  pivot_longer(c(p_temp, p_starv, p_sg),
               names_to = "source", values_to = "probability") |>
  mutate(
    source   = recode(source,
      p_temp  = "Temperature (p_temp)",
      p_starv = "Starvation (p_starv)",
      p_sg    = "Size (p_sg)"),
    strategy = factor(strategy,
      levels = c("resident_warm", "resident_cold", "optimal_mover"))
  ) |>
  filter(strategy != "resident_warm")

mycols <- brewer.pal(4, "Set2")
surv_source_colors <- c(
  "Temperature (p_temp)" = mycols[1],
  "Starvation (p_starv)" = mycols[2],
  "Size (p_sg)"   = mycols[3]
)

# Individual fish: one line per fish per mortality source, faceted by strategy
all_surv_df |>
  filter(strategy == "optimal_mover") |>
  ggplot(aes(x = date, y = probability, color = source,
             group = interaction(pid, source))) +
  geom_rect(data = summer_shading,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
  geom_line(alpha = 0.4, linewidth = 0.3) +
  scale_color_manual(values = surv_source_colors) +
  #scale_y_continuous(limits = c(0, 1)) +
  facet_wrap(~(cohort), scales = "free_y") +
  theme_bw() + theme(panel.grid = element_blank()) +
  xlab("Date") + ylab("Daily survival probability") +
  labs(title = "OPTIMAL MOVERS: Mortality sources by cohort — individual fish",
       color = "Mortality source")

all_surv_df |>
  filter(strategy == "resident_cold") |>
  ggplot(aes(x = date, y = probability, color = source,
             group = interaction(pid, source))) +
  geom_rect(data = summer_shading,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
  geom_line(alpha = 0.4, linewidth = 0.3) +
  scale_color_manual(values = surv_source_colors) +
  #scale_y_continuous(limits = c(0, 1)) +
  facet_wrap(~(cohort), scales = "free_y") +
  theme_bw() + theme(panel.grid = element_blank()) +
  xlab("Date") + ylab("Daily survival probability") +
  labs(title = "COLD RESIDENTS: Mortality sources by cohort — individual fish",
       color = "Mortality source")

# Group means: daily mean per mortality source, faceted by strategy
all_surv_df |>
  filter(strategy == "optimal_mover") |>
  group_by(strategy, date, cohort, source) |>
  summarise(mean_p = mean(probability), .groups = "drop") |>
  ggplot(aes(x = date, y = mean_p, color = source)) +
  geom_rect(data = summer_shading,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = surv_source_colors) +
  #scale_y_continuous(limits = c(0, 1)) +
  facet_wrap(~cohort, scales = "free_y") +
  theme_bw() + theme(panel.grid = element_blank()) +
  xlab("Date") + ylab("Mean daily survival probability") +
  labs(title = "OPTIMAL MOVERS: Mortality sources by cohort — group means",
       color = "Mortality source")

all_surv_df |>
  filter(strategy == "resident_cold") |>
  group_by(strategy, date, cohort, source) |>
  summarise(mean_p = mean(probability), .groups = "drop") |>
  ggplot(aes(x = date, y = mean_p, color = source)) +
  geom_rect(data = summer_shading,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = surv_source_colors) +
  #scale_y_continuous(limits = c(0, 1)) +
  facet_wrap(~cohort, scales = "free_y") +
  theme_bw() + theme(panel.grid = element_blank()) +
  xlab("Date") + ylab("Mean daily survival probability") +
  labs(title = "COLD RESIDENTS: Mortality sources by cohort — group means",
       color = "Mortality source")
```



Condition/starvation is a one of the primary drivers or survival. Plot relative condition over time, for each strategy. 
```{r fig.width = 8, fig.height=7}
# ibm_summary |>
#   # filter(strategy != "resident_warm") |>   # warm residents extinct early — plot separately
#   ggplot(aes(x = date, color = strategy, fill = strategy)) +
#   geom_rect(data = summer_shading,
#             aes(xmin=xmin, xmax=xmax, ymin=-Inf, ymax=Inf),
#             fill="grey", alpha=0.3, inherit.aes=FALSE) +
#   geom_ribbon(aes(ymin = mean_condition - sd_condition,
#                   ymax = mean_condition + sd_condition),
#               alpha = 0.15, color = NA) +
#   geom_line(aes(y = mean_condition)) +
#   geom_hline(yintercept = c(0.50, 0.30), linetype="dashed", alpha=0.5) +
#   annotate("text", x = as.Date("2001-12-01"), y = 0.53, hjust=0,
#            label="K9 = 0.50", size=3, alpha=0.7) +
#   annotate("text", x = as.Date("2001-12-01"), y = 0.33, hjust=0,
#            label="K1 = 0.30", size=3, alpha=0.7) +
#   scale_color_manual(values=c(resident_warm = "red", resident_cold="blue", optimal_mover="darkgreen")) +
#   scale_fill_manual(values=c(resident_warm = "red", resident_cold="blue",  optimal_mover="darkgreen")) +
#   scale_y_continuous(limits=c(0,1)) +
#   theme_bw() + theme(panel.grid=element_blank()) +
#   xlab("Date") + ylab("Relative condition (W / W_peak)") +
#   labs(title="Mean condition ± SD by strategy over time",
#        subtitle="Dashed lines = starvation anchor points (K9, K1); shaded = June–Aug",
#        color="Strategy", fill="Strategy")

ibm_long |>
  # filter(strategy != "resident_warm") |>   # warm residents extinct early — plot separately
  ggplot(aes(x = date, y = condition, color = strategy, fill = strategy)) +
  # geom_rect(data = summer_shading,
  #           aes(xmin=xmin, xmax=xmax, ymin=-Inf, ymax=Inf),
  #           fill="grey", alpha=0.3, inherit.aes=FALSE) +
  geom_line(alpha = 0.5) +
  geom_hline(yintercept = c(0.50, 0.30), linetype="dashed", alpha=0.5) +
  # annotate("text", x = as.Date("2001-12-01"), y = 0.53, hjust=0,
  #          label="K9 = 0.50", size=3, alpha=0.7) +
  # annotate("text", x = as.Date("2001-12-01"), y = 0.33, hjust=0,
  #          label="K1 = 0.30", size=3, alpha=0.7) +
  scale_color_manual(values=c(resident_warm = "red", resident_cold="blue", optimal_mover="darkgreen")) +
  scale_fill_manual(values=c(resident_warm = "red", resident_cold="blue",  optimal_mover="darkgreen")) +
  scale_y_continuous(limits=c(0,1)) +
  theme_bw() + theme(panel.grid=element_blank()) +
  xlab("Date") + ylab("Relative condition (W / W_peak)") +
  labs(title="Mean condition ± SD by strategy over time",
       subtitle="Dashed lines = starvation anchor points (K9, K1); shaded = June–Aug",
       color="Strategy", fill="Strategy") +
  facet_wrap(~cohort, scales = "free_x")
```
Optimal movers and cold residents maintain relatively high condition, and we only see mass loss during the winter when very cold temperatures limit physiological capacity, and immediately following spawning (although this is not quite as extreme). Neither strategy ever approaches the starvation anchor points, meaning that the starvation mechanism is correctly not firing during ~normal seasonal weight loss. In contrast, condition for warm residents declines steeply during the first summer, when food limitations (threshold P_cmax) are insufficient to meet temperature-driven physiological demands. 




### Movement

#### Population trends

Patch occupancy for optimal movers — proportion in warm patch through the year.

* Stochasticity in movement
  + If `move_stochastic = "prob"`, movers only really start to pile into the warm habitat during the shoulder seasons when benefits of warm habitat far exceed the cold habitat. 
  + If `move_stochastic = "Indiv"`, ...
* Environment sensing
  + If `sense_environment = "density_current"` or `sense_environment = "density_all"`, movers bounce rapidly between the two habitats as they chase the apparently better habitat, but then experience reduced growth due to strong density dependence. This appears to be driven by fact that most movers are doing the same thing.
```{r fig.width=8, fig.height=6}
ibm_long %>%
  filter(strategy == "optimal_mover") %>%
  group_by(cohort, date) %>%
  summarise(prop_warm = mean(patch == "warm")) %>%
  ungroup() %>%
  ggplot(aes(x = date, y = prop_warm)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray50") +
  theme_bw() + facet_wrap(~cohort, scales = "free_x") +
  xlab("Date") + ylab("Proportion of optimal movers in warm patch")
```

Zoom into single cohort
```{r}
ibm_long %>%
  filter(strategy == "optimal_mover") %>%
  group_by(cohort, date) %>%
  summarise(prop_warm = mean(patch == "warm")) %>%
  ungroup() %>%
  filter(cohort == 2012) %>%
  ggplot(aes(x = date, y = prop_warm)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray50") +
  theme_bw() + facet_wrap(~cohort, scales = "free_x") +
  xlab("Date") + ylab("Proportion of optimal movers in warm patch")
```


#### Individual movement patterns

Build a long-form patch history for optimal movers:
```{r eval=FALSE}
patch_long <- as_tibble(patch_matrix) |>
  bind_cols(fish_registry |> select(pid, strategy, cohort)) |>
  pivot_longer(-c(pid, strategy, cohort), names_to = "day", values_to = "patch") |>
  mutate(dayofsim = as.integer(sub("V", "", day))) |>
  left_join(habitat_df |> select(dayofsim, date), by = "dayofsim") |>
  filter(!is.na(patch)) |>
  select(-day)

mover_patch <- patch_long |> filter(strategy == "optimal_mover")
```

**Raster of all optimal movers.** Each row is one fish, each column is a day, fill = patch. Fish are ordered by their overall proportion of time in the warm patch. Synchronised individuals will produce horizontal bands switching at the same time; independent movers will produce a noisier mosaic.

At seasonal and annual timescales, fish show similar movement patterns--tracking the primary gradient in growth potential and it shifts seasonally between warm and cold patches. But we also see individual variation at finer timescales, driven by stochasticity from the softmax patch selection function. Another observation is that the allometric scaling of the cost of movement appears to be working as expected. Early in the simulation, small fish switch infrequently (only when the benefit of switching is very large). But later on, larger fish switch more frequently, navigating smaller magnitude differences in density-dependent growth potential. 
```{r fig.width=9, fig.height=8}
# Order fish by proportion of time spent in warm patch
pid_order <- ibm_long |>
  group_by(pid) |>
  summarise(prop_warm = mean(patch == "warm"), .groups = "drop") |>
  arrange(prop_warm)

# All cohorts
ibm_long |>
  mutate(pid = factor(pid, levels = pid_order$pid)) |>
  ggplot(aes(x = date, y = pid, fill = patch)) +
  # geom_rect(data = summer_shading,
  #           aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
  #           fill = "grey", alpha = 0.4, inherit.aes = FALSE) +
  geom_tile() +
  scale_fill_manual(values = c(warm = "#d95f02", cold = "#1f78b4")) +
  theme_bw() +
  theme(panel.grid  = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  xlab("Date") + ylab("Individual fish (ordered by % time in warm)") +
  labs(title = "Patch occupancy — all optimal movers",
       subtitle = "Shaded = June–Aug; fish ordered bottom-to-top by increasing warm-patch use",
       fill = "Patch") +
  facet_wrap(~cohort, scales = "free")

# selected cohorts: early, mid, late
# mover_patch |>
#   filter(cohort == 2001) %>%
#   mutate(pid = factor(pid, levels = pid_order$pid)) |>
#   ggplot(aes(x = date, y = pid, fill = patch)) +
#   geom_tile() +
#   scale_fill_manual(values = c(warm = "#d95f02", cold = "#1f78b4")) +
#   theme_bw() +
#   theme(panel.grid  = element_blank(),
#         axis.text.y = element_blank(),
#         axis.ticks.y = element_blank()) +
#   xlab("Date") + ylab("Individual fish (ordered by % time in warm)") +
#   labs(title = "Patch occupancy — all optimal movers",
#        subtitle = "Shaded = June–Aug; fish ordered bottom-to-top by increasing warm-patch use",
#        fill = "Patch") +
#   facet_wrap(~cohort, scales = "free")
# 
# 
# mover_patch |>
#   filter(cohort == 2012) %>%
#   mutate(pid = factor(pid, levels = pid_order$pid)) |>
#   ggplot(aes(x = date, y = pid, fill = patch)) +
#   geom_tile() +
#   scale_fill_manual(values = c(warm = "#d95f02", cold = "#1f78b4")) +
#   theme_bw() +
#   theme(panel.grid  = element_blank(),
#         axis.text.y = element_blank(),
#         axis.ticks.y = element_blank()) +
#   xlab("Date") + ylab("Individual fish (ordered by % time in warm)") +
#   labs(title = "Patch occupancy — all optimal movers",
#        subtitle = "Shaded = June–Aug; fish ordered bottom-to-top by increasing warm-patch use",
#        fill = "Patch") +
#   facet_wrap(~cohort, scales = "free")
# 
# 
# mover_patch |>
#   filter(cohort == 2015) %>%
#   mutate(pid = factor(pid, levels = pid_order$pid)) |>
#   ggplot(aes(x = date, y = pid, fill = patch)) +
#   geom_tile() +
#   scale_fill_manual(values = c(warm = "#d95f02", cold = "#1f78b4")) +
#   theme_bw() +
#   theme(panel.grid  = element_blank(),
#         axis.text.y = element_blank(),
#         axis.ticks.y = element_blank()) +
#   xlab("Date") + ylab("Individual fish (ordered by % time in warm)") +
#   labs(title = "Patch occupancy — all optimal movers",
#        subtitle = "Shaded = June–Aug; fish ordered bottom-to-top by increasing warm-patch use",
#        fill = "Patch") +
#   facet_wrap(~cohort, scales = "free")

```

Fish are generally acting similarly at seasonal and annual timescales, but on any given day fish are largely acting independently. To introduce additional variation in movement strategies, consider the following:

- Inrease `tau`, the sensitivity to growth differences specified by the softmax function (steepness). Increasing `tau` would make all fish more randomly noisy in their decisions (similar to if we add simple random variation) but also may have the unintended consequence of making fish more reluctant to move (i.e., fish need a larger growth benefit to initiate a movement). But `tau` also applies identically to every fish...so any reduced synchrony in movement decisions would be entirely stochastic, rather than biologically meaningful individual variation
- To generate structured among-individual differences in movement, we need to give each fish a persistent, individual-specific response to the same environment. `move_stochastic = "indiv"` with `sigma_bold` is the mechanism for this. Each fish draws a fixed movement threshold from a distribution, so some fish are consistently "bold" (move even when the gain is marginal) and others are consistently "reluctant" (only move when the growth advantage is large). That creates genuine individual variation in movement decisions that persists through time — not just day-to-day noise.


**When to fish make their first move?** For each fish, record every day a patch switch occurs, then trim to the first movement and find the fish age and size on this day. Vertical clustering = synchronised switching; diffuse scatter = independent behaviour.
```{r}
switch_events <- ibm_long |>
  arrange(pid, date) |>
  group_by(pid) |>
  mutate(switched = patch != lag(patch)) |>
  filter(switched) |>
  ungroup()

first_switch <- switch_events |>
  group_by(pid, cohort) |>
  slice_min(dayofsim, n = 1, with_ties = FALSE)

# facet by destination patch
ggplot(first_switch, aes(x = weight)) +
  geom_histogram(bins = 50) +
  scale_x_log10() +
  #xlim(0,3) +
  geom_vline(xintercept = start_wt, linetype = "dashed", color = "red") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  facet_wrap(~patch, ncol = 1, labeller = labeller(patch = c(cold = "First move → cold", warm = "First move → warm"))) +
  xlab("Weight at first switch (g, log scale)") +
  ylab("Count") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

# by date
first_switch |>
  mutate(doy = yday(date)) |>
  ggplot(aes(x = doy)) +
  geom_histogram(bins = 52) +
  facet_wrap(~patch, ncol = 1, 
             labeller = labeller(patch = c(cold = "First move → cold", warm = "First move → warm"))) +
  scale_x_continuous(
    breaks = c(1, 60, 121, 182, 244, 305),
    labels = c("Jan", "Mar", "May", "Jul", "Sep", "Nov")
  ) +
  xlab("Day of year") + ylab("Count") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())


# facet by cohort
ggplot(first_switch, aes(x = weight)) +
  geom_histogram(bins = 50) +
  scale_x_log10() +
  #xlim(0,3) +
  geom_vline(xintercept = start_wt, linetype = "dashed", color = "red") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  facet_wrap(~cohort) +
  xlab("Weight at first switch (g, log scale)") +
  ylab("Count") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```
This seems a little more realistic. Fish are now making their first movements at larger body sizes. For fish born in the warm habitat, their first movement (into the cold patch) occurs in mid-June, just after fish have exceeded the size threshold above which movement becomes less costly but also at a time when growth potentialy in the warm habitat is very low as temperatures exceed metabolic optima. For fish born in the cold habitat, their first movement (into the warm patch) occurs in October, when fish are larger and cold-temps in the cold habitat are limiting to growth.


**How frequently do fish move and how does this change over their life span?** Calculate the number of movements per year for each individual, then plot cohort means over time. 
```{r}
# Join switch_events with ibm_long to get age at each switch
# then count switches per fish per year-of-life
switches_by_age <- switch_events |>
  mutate(age_year = floor(age))

# Count switches per fish per age year
annual_switches <- switches_by_age |>
  count(pid, cohort, age_year, name = "n_switches")

# Quick look at coverage and range
# annual_switches |>
#   summarise(
#     n_fish      = n_distinct(pid),
#     age_yr_max  = max(age_year),
#     median_sw   = median(n_switches),
#     max_sw      = max(n_switches)
#   )

# Summarise: mean + 1 SD per (cohort × age year)
switch_summary <- annual_switches |>
  group_by(cohort, age_year) |>
  summarise(
    mean_sw  = mean(n_switches),
    sd_sw    = sd(n_switches),
    median_sw = median(n_switches),
    q25      = quantile(n_switches, 0.25),
    q75      = quantile(n_switches, 0.75),
    n        = n(),
    .groups  = "drop"
  )

# Plot: mean ± SD ribbon per cohort-year, faceted by cohort
# Keep only age years with ≥ 3 fish to avoid noisy tails
switch_summary |>
  filter(n >= 3) |>
  ggplot(aes(x = age_year, y = mean_sw)) +
  geom_ribbon(aes(ymin = pmax(0, mean_sw - sd_sw),
                  ymax = mean_sw + sd_sw),
              alpha = 0.25) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.2) +
  facet_wrap(~cohort, ncol = 6) +
  scale_x_continuous(breaks = seq(0, 16, by = 4)) +
  labs(
    x        = "Age (years)",
    y        = "Switches per year (mean ± SD)",
    title    = "Annual patch switching frequency over the lifespan — optimal movers",
    subtitle = "Each panel = cohort year; shaded band = ±1 SD; points shown for age classes with ≥ 3 fish"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        strip.text       = element_text(size = 8))

# Overlaid view: all cohorts on one panel, colored by cohort year
switch_summary |>
  filter(n >= 3) |>
  ggplot(aes(x = age_year, y = mean_sw,
             color = factor(cohort), group = cohort)) +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(size = 1.5, alpha = 0.8) +
  scale_color_viridis_d(option = "turbo", name = "Cohort") +
  scale_x_continuous(breaks = seq(0, 16, by = 2)) +
  labs(
    x        = "Age (years)",
    y        = "Mean switches per year",
    title    = "Annual patch switching frequency by cohort — optimal movers",
    subtitle = "Each line = one cohort; points shown for age classes with ≥ 3 fish"
  ) +
  theme_bw() +
  theme(panel.grid.minor  = element_blank(),
        legend.position   = "right",
        legend.key.height = unit(0.5, "cm"))
```
This confirms a couple things. First, small fish move infrequently b/c the relative energetic cost of movement is quite high. As fish age and grow larger, they move more frequently b/c the cost of movement is considerably lower (on a g/g/d basis). 

(Switch rate almost always drops off during the final year b/c fish are only observed for a partial year during the year of death)

What is striking about these results is that older fish are making very frequent movement during the winter. This is likely driven by two things. First, temperatures are very similar between the two habitat during winter, so the benefit of being in the warm habitat is not that strong. Second, 
large fish experience a very low cost of movement on a g/g/d basis, so they are more free to move back and forth to negotiate these relatively small differences in growth potential.


### Reproduction

Build spawner traits database
```{r}
spawn_days <- unique(spawn_log$dayofsim)

# Spawner traits: join spawn_log → fish_registry (strategy) → ibm_long (age only)
# weight and condition come from spawn_log directly (pre-spawn values logged at spawn time)
spawner_traits <- spawn_log |>
  left_join(fish_registry |> select(pid, strategy), by = c("parent_pid" = "pid")) |>
  left_join(
    ibm_long |> select(pid, dayofsim, age),
    by = c("parent_pid" = "pid", "dayofsim")
  ) |>
  transmute(
    pid        = parent_pid,
    dayofsim,
    weight, condition, age, strategy,
    spawned    = "Spawner"
  )

# Non-spawner traits: all fish alive on those same days that didn't spawn
non_spawner_traits <- ibm_long |>
  filter(dayofsim %in% spawn_days, !is.na(weight), !is.na(condition)) |>
  anti_join(spawn_log, by = c("pid" = "parent_pid", "dayofsim")) |>
  transmute(pid, dayofsim, weight, condition, age, strategy, spawned = "Non-spawner")

spawner_df <- bind_rows(spawner_traits, non_spawner_traits)

cat("Spawner records:", nrow(spawner_traits), "\n")
cat("Non-spawner records:", nrow(non_spawner_traits), "\n")
cat("NAs in spawner traits:\n")
spawner_traits |> summarise(across(everything(), ~sum(is.na(.))))
```


#### Spawner traits

Compare spawner vs. non-spawner traits
```{r}
# Add sample size labels for annotation
n_labels <- spawner_df |>
  group_by(strategy, spawned) |>
  summarise(n = n(), .groups = "drop") |>
  mutate(label = paste0("n=", n))

# Filter to strategies with spawners
plot_df_spawn <- spawner_df |>
  filter(strategy != "resident_warm") |>
  mutate(strategy = factor(strategy, levels = c("optimal_mover", "resident_cold"),
                           labels = c("Optimal mover", "Resident (cold)")))

n_labels_plot <- n_labels |>
  filter(strategy != "resident_warm") |>
  mutate(strategy = factor(strategy, levels = c("optimal_mover", "resident_cold"),
                           labels = c("Optimal mover", "Resident (cold)")))

make_trait_plot <- function(data, x_var, x_label) {
  ggplot(data, aes(x = .data[[x_var]], fill = spawned, color = spawned)) +
    geom_density(aes(y = after_stat(scaled)), alpha = 0.35, linewidth = 0.6) +
    geom_rug(
      data = filter(data, spawned == "Spawner"),
      sides = "b", linewidth = 0.5, alpha = 0.8, color = "#d62728"
    ) +
    facet_wrap(~strategy, ncol = 1) +
    scale_fill_manual(values = c("Non-spawner" = "#7bafd4", "Spawner" = "#d62728")) +
    scale_color_manual(values = c("Non-spawner" = "#7bafd4", "Spawner" = "#d62728")) +
    labs(x = x_label, y = "Scaled density", fill = NULL, color = NULL) +
    theme_minimal(base_size = 11) +
    theme(
      legend.position = "bottom",
      strip.text = element_text(face = "bold"),
      panel.spacing = unit(0.8, "lines")
    )
}

p_wt_dist2  <- make_trait_plot(plot_df_spawn, "weight",    "Weight (g)")
p_cond_dist2 <- make_trait_plot(plot_df_spawn, "condition", "Condition (K)")

(p_wt_dist2 + p_cond_dist2 + plot_layout(guides = "collect")) +
  plot_annotation(
    title    = "Spawner vs. non-spawner traits on spawning days",
    subtitle = "Resident (warm) excluded — no spawning events observed",
    theme    = theme(plot.title = element_text(face = "bold", size = 13))
  ) &
  theme(legend.position = "bottom")
```
A few things stand out from these plots:

**Weight** — Spawners (red) skew strongly toward larger body sizes compared to the broader non-spawner pool, which is expected since the non-spawner group includes many small/young fish. 

**Condition** — The condition plot is more striking. Spawners cluster at intermediate condition (~0.75–0.95), while a large portion of non-spawners pile up at K ≈ 1.0. That spike at the upper end of non-spawners likely reflects the maximum condition cap in the IBM. This suggests spawning is triggered before fish reach peak condition — consistent with a threshold trigger — or that fish with very high condition are young/small and not yet mature.


#### Spawn timing
```{r fig.width=6, fig.height=9}
spawn_plot_df <- spawn_log |>
  left_join(fish_registry |> select(pid, strategy), by = c("parent_pid" = "pid")) |>
  left_join(
    ibm_long |> select(pid, dayofsim, date),
    by = c("parent_pid" = "pid", "dayofsim")
  ) |>
  mutate(
    year       = year(date),
    doy        = yday(date),
    strategy_lbl = recode(strategy,
      optimal_mover = "Optimal mover",
      resident_cold = "Resident (cold)"
    ),
    year_f = factor(year)
  )

# Top panel: pooled density by strategy (overall DOY distribution)
p_density <- ggplot(spawn_plot_df, aes(x = doy, y = strategy_lbl, fill = strategy_lbl)) +
  geom_density_ridges(
    alpha = 0.55, scale = 0.85, bandwidth = 3,
    jittered_points = TRUE,
    point_size = 2.5, point_alpha = 0.8,
    position = position_points_jitter(height = 0.05, seed = 812)
  ) +
  scale_x_continuous(
    breaks = seq(114, 136, by = 4),
    labels = function(x) format(as.Date(x - 1, origin = "2001-01-01"), "%b %d")
  ) +
  scale_fill_manual(values = c("Optimal mover" = "darkgreen", "Resident (cold)" = "blue"),
                    guide = "none") +
  labs(
    title = "Distribution of spawn timing (pooled across years)",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(axis.text.y = element_text(face = "bold"))

# Bottom panel: per-year dot strip, size = n_offspring
p_annual <- ggplot(spawn_plot_df,
                   aes(x = doy, y = fct_rev(year_f), color = strategy_lbl, size = weight)) +
  geom_point(alpha = 0.75) +
  scale_x_continuous(
    breaks = seq(114, 136, by = 4),
    labels = function(x) format(as.Date(x - 1, origin = "2001-01-01"), "%b %d")
  ) +
  scale_color_manual(values = c("Optimal mover" = "darkgreen", "Resident (cold)" = "blue")) +
  scale_size_continuous(range = c(2, 9), name = "Weight (g)") +
  labs(
    title    = "Spawning events by year",
    subtitle = "Point size proportional to weight at spawning",
    x = "Day of year", y = "Year",
    color = "Strategy"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "right")

# combine plots
ggpubr::ggarrange(p_density, p_annual, nrow = 2, heights = c(1, 2.5))
```

Does mean spawn timing change over time?
```{r}
spawn_plot_df %>%
  ggplot(aes(x = jitter(year), y = doy, color = strategy, group = strategy)) +
  geom_point() +
  theme_bw() +
  geom_smooth() + #facet_wrap(~strategy, scales = "free_y") +
  scale_color_manual(values = c("optimal_mover" = "darkgreen", "resident_cold" = "blue")) +
  xlab("Year") + ylab("Spawning day of year")
```

Is spawning date related to weight or condition?
```{r fig.width=6, fig.height=6}
p1 <- spawn_plot_df %>%
  ggplot(aes(x = weight, y = doy)) +
  geom_point() + 
  theme_bw() +
  geom_smooth(method = "lm") + facet_wrap(~strategy, scales = "free") +
  xlab("Weight (g)") + ylab("Spawning day of year")

p2 <- spawn_plot_df %>%
  ggplot(aes(x = condition, y = doy)) +
  geom_point() + 
  theme_bw() +
  geom_smooth(method = "lm") + facet_wrap(~strategy, scales = "free") +
  xlab("Condition") + ylab("Spawning day of year")

egg::ggarrange(p1, p2, nrow = 2)
```


Do weight and condition interact to drive spawn timing?
```{r}
mod_spawn <- lm(doy ~ weight * condition, data = spawn_plot_df)
summary(mod_spawn)
```

```{r}
# Prediction grid over the full observed range (pooled model)
pred_grid <- expand.grid(
  weight    = seq(min(spawn_plot_df$weight),    max(spawn_plot_df$weight),    length.out = 120),
  condition = seq(min(spawn_plot_df$condition), max(spawn_plot_df$condition), length.out = 120)
)
pred_grid$doy_pred <- predict(mod_spawn, newdata = pred_grid)

ggplot(pred_grid, aes(x = weight, y = condition, fill = doy_pred)) +
  geom_raster(interpolate = TRUE) +
  geom_point(
    data = spawn_plot_df |>
      mutate(strategy_lbl = recode(strategy,
        optimal_mover = "Optimal mover",
        resident_cold = "Resident (cold)"
      )),
    aes(x = weight, y = condition, shape = strategy_lbl),
    color = "white", size = 1.5, alpha = 0.7,
    inherit.aes = FALSE
  ) +
  scale_fill_viridis_c(name = "Predicted\nspawn DOY", option = "plasma") +
  scale_shape_manual(values = c("Optimal mover" = 16, "Resident (cold)" = 17),
                     name = "Strategy") +
  labs(
    x        = "Weight (g)",
    y        = "Condition (K)",
    title    = "Interactive effects of weight and condition on spawn timing",
    subtitle = "Pooled model: DOY ~ weight × condition; points = observed spawners"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "right")
```

Fit separate models per strategy to test whether the interaction differs:
```{r}
library(broom)

strat_models <- spawn_plot_df |>
  mutate(strategy_lbl = recode(strategy,
    optimal_mover = "Optimal mover",
    resident_cold = "Resident (cold)"
  )) |>
  nest_by(strategy_lbl) |>
  mutate(mod = list(lm(doy ~ weight * condition, data = data)))

strat_models |>
  reframe(tidy(mod, conf.int = TRUE)) |>
  filter(term != "(Intercept)") |>
  select(strategy_lbl, term, estimate, std.error, p.value, conf.low, conf.high)
```

```{r}
# Strategy-specific prediction grids
pred_grid_strat <- purrr::map2_dfr(
  strat_models$strategy_lbl,
  strat_models$mod,
  function(strat, mod) {
    obs <- spawn_plot_df |>
      mutate(strategy_lbl = recode(strategy,
        optimal_mover = "Optimal mover",
        resident_cold = "Resident (cold)"
      )) |>
      filter(strategy_lbl == strat)
    grid <- expand.grid(
      weight    = seq(min(obs$weight),    max(obs$weight),    length.out = 100),
      condition = seq(min(obs$condition), max(obs$condition), length.out = 100)
    )
    grid$doy_pred     <- predict(mod, newdata = grid)
    grid$strategy_lbl <- strat
    grid
  }
)

obs_points <- spawn_plot_df |>
  mutate(strategy_lbl = recode(strategy,
    optimal_mover = "Optimal mover",
    resident_cold = "Resident (cold)"
  ))

ggplot(pred_grid_strat, aes(x = weight, y = condition, fill = doy_pred)) +
  geom_tile() +
  geom_point(
    data = obs_points,
    aes(x = weight, y = condition),
    color = "white", size = 1.5, alpha = 0.6,
    inherit.aes = FALSE
  ) +
  facet_wrap(~strategy_lbl, scales = "free") +
  scale_fill_viridis_c(name = "Predicted\nspawn DOY", option = "plasma") +
  labs(
    x        = "Weight (g)",
    y        = "Condition (K)",
    title    = "Interactive effects of weight and condition on spawn timing",
    subtitle = "Separate models per strategy; DOY ~ weight × condition"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "right",
    strip.text      = element_text(face = "bold"),
    panel.spacing   = unit(1, "lines")
  )
```
The coefficient table tells the key story: for optimal movers, weight, condition, and their interaction are all highly significant (p < 0.0001) — heavier fish with low condition spawn earlier, but this weight effect diminishes as condition increases. For resident (cold), none of the terms reach significance, likely reflecting the small sample (n=63) rather than a genuine absence of the pattern.

The sign and magnitude of the interaction term is similar between strategies (−0.17 vs. −0.62), so the shape of the relationship is probably the same — the cold resident model just lacks the power to detect it.


#### Spawning by age and cohort

Age at first spawning
```{r}
# First spawning event per fish
age_first_spawn <- spawn_log |>
  group_by(parent_pid) |>
  slice_min(dayofsim, n = 1, with_ties = FALSE) |>
  ungroup() |>
  left_join(fish_registry |> select(pid, strategy, cohort), by = c("parent_pid" = "pid")) |>
  left_join(
    ibm_long |> select(pid, dayofsim, age, date),
    by = c("parent_pid" = "pid", "dayofsim")
  ) |>
  mutate(
    spawn_year    = year(date),
    age_first_yrs = age,
    strategy_lbl  = recode(strategy,
      optimal_mover = "Optimal mover",
      resident_cold = "Resident (cold)"
    )
  )

cat("n fish that ever spawned:", nrow(age_first_spawn), "\n\n")
age_first_spawn |>
  group_by(strategy_lbl) |>
  summarise(
    n      = n(),
    mean   = mean(age_first_yrs),
    median = median(age_first_yrs),
    min    = min(age_first_yrs),
    max    = max(age_first_yrs),
    .groups = "drop"
  )

ggplot(age_first_spawn, aes(x = strategy_lbl, y = age_first_yrs)) +
  # IQR bar + mean for optimal_mover
  geom_boxplot(aes(fill = strategy_lbl), alpha = 0.2) +
  # geom_linerange(
  #   data = summary_afs,
  #   aes(x = strategy_lbl, ymin = lo, ymax = hi, color = strategy_lbl),
  #   linewidth = 4, alpha = 0.25, inherit.aes = FALSE
  # ) +
  # geom_point(
  #   data = summary_afs,
  #   aes(x = strategy_lbl, y = median, color = strategy_lbl),
  #   shape = 18, size = 5, inherit.aes = FALSE
  # ) +
  # Individual fish
  geom_jitter(aes(color = strategy_lbl), width = 0.12, size = 2.5, alpha = 0.65) +
  # n label
  # geom_text(
  #   data = summary_afs,
  #   aes(x = strategy_lbl, y = -0.3, label = paste0("n = ", n), color = strategy_lbl),
  #   size = 3.5, fontface = "italic", inherit.aes = FALSE
  # ) +
  scale_color_manual(
    values = c("Optimal mover" = "darkgreen", "Resident (cold)" = "blue"),
    guide = "none"
  ) +
  scale_fill_manual(
    values = c("Optimal mover" = "darkgreen", "Resident (cold)" = "blue"),
    guide = "none"
  ) +
  # scale_y_continuous(breaks = seq(0, 12, by = 2), limits = c(-0.7, 13)) +
  labs(
    title    = "Age at first spawning by strategy",
    subtitle = "Diamond = median  |  Shaded bar = IQR  |  One artifact excluded (age < 30 days)",
    x = NULL, y = "Age at first spawn (years)"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.x = element_blank())
```


Compute and plot number and proportion of fish spawning in each year, by cohort and strategy
```{r fig.width=6, fig.height=9}
# ── Spawner counts per cohort × year ──────────────────────────────────────────
spawners_by_cohort_yr <- spawn_log |>
  left_join(fish_registry |> select(pid, strategy, cohort), by = c("parent_pid" = "pid")) |>
  left_join(
    ibm_long |> select(pid, dayofsim, date) |> distinct(),
    by = c("parent_pid" = "pid", "dayofsim")
  ) |>
  mutate(spawn_year = year(date)) |>
  group_by(cohort, spawn_year, strategy) |>
  summarise(n_spawners = n_distinct(parent_pid), .groups = "drop")

# ── Fish alive during spawning season (May) per cohort × year ─────────────────
alive_spawning_season <- ibm_long |>
  filter(month(date) == 5) |>
  mutate(spawn_year = year(date)) |>
  group_by(cohort, spawn_year, strategy) |>
  summarise(n_alive = n_distinct(pid), .groups = "drop")

# ── Join and compute proportion ────────────────────────────────────────────────
cohort_heatmap_df <- alive_spawning_season |>
  left_join(spawners_by_cohort_yr, by = c("cohort", "spawn_year", "strategy")) |>
  mutate(
    n_spawners   = replace_na(n_spawners, 0),
    prop_spawned = n_spawners / n_alive
  ) |>
  # Only show years where cohort could plausibly be alive (spawn_year >= cohort)
  filter(spawn_year > cohort)

# cat("Preview:\n")
# cohort_heatmap_df |> filter(n_spawners > 0)


### PLOT 

strategy_labels <- c(
  optimal_mover = "Optimal mover",
  resident_cold = "Resident (cold)",
  resident_warm = "Resident (warm)"
)

# Age at spawning (years since cohort birth) for axis annotation
plot_df_heat <- cohort_heatmap_df |>
  mutate(
    age_at_spawn  = spawn_year - cohort,
    strategy_lbl  = recode(strategy, !!!strategy_labels),
    # Treat zero-spawning cells as NA so they're visually distinct from
    # years where the cohort was simply absent
    prop_display  = if_else(prop_spawned == 0, NA_real_, prop_spawned)
  )

ggplot(plot_df_heat, aes(x = factor(spawn_year), y = factor(cohort), fill = prop_display)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(
    data = filter(plot_df_heat, n_spawners > 0),
    aes(label = n_spawners),
    size = 2.8, color = "white", fontface = "bold"
  ) +
  scale_fill_viridis_c(
    option = "magma", direction = -1,
    na.value = "grey92",
    name = "Proportion\nspawning",
    labels = scales::label_percent(accuracy = 1)
  ) +
  facet_wrap(~strategy_lbl, ncol = 1, scales = "free_y") +
  labs(
    title    = "Proportion of each cohort spawning per year",
    subtitle = "Cell labels show number of spawners  |  Grey = cohort alive but no spawning events",
    x = "Spawning year", y = "Cohort (birth year)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x    = element_text(angle = 45, hjust = 1),
    strip.text     = element_text(face = "bold"),
    panel.grid     = element_blank(),
    legend.key.height = unit(1.2, "cm")
  )
```


A few things stand out here:

**Optimal mover** — The 2001 cohort shows a strong burst of 6 spawners in 2003 (~32% of survivors), which is the highest single-year proportion. But spawning is generally sparse and episodic — most cohort × year cells with survivors show no spawning event. The dark cell (2004 cohort, 2014) hits ~100% but that's a very small surviving cohort (1 of 1 fish).

**Resident (cold)** — Only the 2001 cohort produced any spawners (4 total across three years), and no subsequent cohort spawned at all. This is consistent with the earlier finding that resident_cold almost never reproduces.

These results may be driven by the fact that fish are not growing large enough to spawn consistently. 


#### Reproductive success

Reproductive success by fish size. 
```{r}
fecundity_df <- spawn_log |>
  left_join(fish_registry |> select(pid, strategy), by = c("parent_pid" = "pid")) |>
  left_join(
    ibm_long |> select(pid, dayofsim, age, date),
    by = c("parent_pid" = "pid", "dayofsim")
  ) |>
  mutate(
    strategy_lbl = recode(strategy,
      optimal_mover = "Optimal mover",
      resident_cold = "Resident (cold)"
    ),
    spawn_year = year(date)
  ) |>
  filter(age >= 30 / 365)  # remove newborn artifact


pal <- c("Optimal mover" = "darkgreen", "Resident (cold)" = "blue")

# Pearson correlation for annotation
r_om <- with(filter(fecundity_df, strategy_lbl == "Optimal mover"),
             cor(weight, n_offspring))
r_rc <- with(filter(fecundity_df, strategy_lbl == "Resident (cold)"),
             cor(weight, n_offspring))

ann_df <- tibble(
  strategy_lbl = c("Optimal mover", "Resident (cold)"),
  label        = c(sprintf("r = %.2f  (n = %d)", r_om,
                           sum(fecundity_df$strategy_lbl == "Optimal mover")),
                   sprintf("r = %.2f  (n = %d)", r_rc,
                           sum(fecundity_df$strategy_lbl == "Resident (cold)")))
)

ggplot(fecundity_df, aes(x = weight, y = n_offspring, color = strategy_lbl)) +
  geom_point(aes(size = age), alpha = 0.75) +
  geom_smooth(
    #data  = filter(fecundity_df, strategy_lbl == "Optimal mover"),
    method = "lm", se = TRUE, linewidth = 0.8, color = "black"#, alpha = 0.15
  ) +
  geom_text(
    data    = ann_df,
    aes(label = label),
    x = -Inf, y = Inf, hjust = -0.1, vjust = 1.5,
    size = 3.5, fontface = "italic", inherit.aes = FALSE
  ) +
  scale_color_manual(values = pal, guide = "none") +
  scale_size_continuous(range = c(2, 8), name = "Age at\nspawn (yrs)") +
  facet_wrap(~strategy_lbl, scales = "free_x") +
  labs(
    title    = "Fecundity–weight relationship at spawning",
    subtitle = "Point size = age at spawning  |  Shaded band = 95% CI on linear fit",
    x        = "Weight at spawning (g)",
    y        = "Number of offspring"
  ) +
  theme_bw(base_size = 12) +
  theme(
    panel.grid.minor = element_blank()
  )
```


View total recruitment (offspring production) and per-capita reproductive success over time, by strategy
```{r fig.width=6, fig.height=9}
# Build annual offspring summary
offspring_annual <- spawn_log |>
  left_join(fish_registry |> select(pid, strategy), by = c("parent_pid" = "pid")) |>
  left_join(
    ibm_long |> select(pid, dayofsim, date) |> distinct(),
    by = c("parent_pid" = "pid", "dayofsim")
  ) |>
  mutate(
    spawn_year   = year(date),
    strategy_lbl = recode(strategy,
      optimal_mover = "Optimal mover",
      resident_cold = "Resident (cold)"
    )
  ) |>
  group_by(spawn_year, strategy_lbl) |>
  summarise(
    total_offspring  = sum(n_offspring),
    n_spawners       = n_distinct(parent_pid),
    offspring_per_spawner = total_offspring / n_spawners,
    .groups = "drop"
  )

# offspring_annual


# Fill in years with no spawning as explicit zeros for total offspring
all_years <- tibble(spawn_year = 2002:2020)
strategies <- tibble(strategy_lbl = c("Optimal mover", "Resident (cold)"))

offspring_full <- crossing(all_years, strategies) |>
  left_join(offspring_annual, by = c("spawn_year", "strategy_lbl")) |>
  mutate(
    total_offspring       = replace_na(total_offspring, 0),
    n_spawners            = replace_na(n_spawners, 0)
    # leave offspring_per_spawner as NA when no spawning
  )

pal <- c("Optimal mover" = "darkgreen", "Resident (cold)" = "blue")

# ── Panel 1: Total offspring per year ─────────────────────────────────────────
p_total <- ggplot(offspring_full,
                  aes(x = spawn_year, y = total_offspring, color = strategy_lbl)) +
  geom_line(linewidth = 0.8, alpha = 0.8) +
  geom_point(aes(size = n_spawners), alpha = 0.8) +
  scale_color_manual(values = pal, guide = "none") +
  scale_size_continuous(range = c(1.5, 7), name = "# spawners",
                        breaks = c(1, 3, 5, 7)) +
  scale_x_continuous(breaks = 2002:2020) +
  labs(
    title = "Total offspring produced per year",
    x = NULL, y = "Total offspring"
  ) +
  facet_wrap(~strategy_lbl, ncol = 1, scales = "free_y") +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x      = element_text(angle = 45, hjust = 1),
    strip.text       = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )

# Colour strip backgrounds per strategy
p_total <- p_total +
  theme(strip.text = element_text(face = "bold"))

# ── Panel 2: Per-spawner offspring (only years with spawning) ─────────────────
p_per_spawner <- ggplot(
    offspring_full |> filter(!is.na(offspring_per_spawner)),
    aes(x = spawn_year, y = offspring_per_spawner, color = strategy_lbl)
  ) +
  geom_hline(
    data = offspring_annual |>
      group_by(strategy_lbl) |>
      summarise(grand_mean = mean(offspring_per_spawner), .groups = "drop"),
    aes(yintercept = grand_mean, color = strategy_lbl),
    linetype = "dashed", linewidth = 0.6, alpha = 0.6
  ) +
  geom_line(linewidth = 0.8, alpha = 0.7) +
  geom_point(size = 3, alpha = 0.85) +
  scale_color_manual(values = pal, name = "Strategy") +
  scale_x_continuous(breaks = 2002:2020) +
  labs(
    title    = "Offspring per spawner (per capita reproductive success)",
    subtitle = "Dashed line = grand mean per strategy",
    x = "Year", y = "Offspring per spawner"
  ) +
  facet_wrap(~strategy_lbl, ncol = 1, scales = "free_y") +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x      = element_text(angle = 45, hjust = 1),
    strip.text       = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    legend.position  = "none"
  )

p_total / p_per_spawner +
  plot_annotation(
    title = "Reproductive output over time",
    theme = theme(plot.title = element_text(face = "bold", size = 13))
  )
```


### Boldness

Compile end of simulation results
```{r eval=FALSE}
fish_results <- fish_pop %>%
  mutate(
    final_weight   = mass_matrix[, n_days],
    n_switches     = switches
    # boldness       = move_threshold   # positive = shy, negative = bold
  )

fish_results %>%
  group_by(strategy) %>%
  summarise(mean_wt = round(mean(final_weight), 1),
            mean_switches = round(mean(n_switches), 1))
```

Plot only if `move_stochastic = "indiv"`
```{r fig.width=5, fig.height=7}
if (move_stochastic == "indiv") {
  
movers <- fish_results %>% filter(strategy == "optimal_mover")

# Panel A: boldness vs final weight
p_wt <- ggplot(movers, aes(x = boldness, y = final_weight)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "steelblue", linewidth = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  annotate("text", x =  0.006, y = min(movers$final_weight) + 5,
           label = "shy →",  hjust = 0, size = 3.5, color = "grey40") +
  annotate("text", x = -0.006, y = min(movers$final_weight) + 5,
           label = "← bold", hjust = 1, size = 3.5, color = "grey40") +
  labs(x = "Boldness (move_threshold)", y = "Final weight (g)",
       title = "A: Boldness vs. end-of-season weight")

# Panel B: boldness vs number of patch switches
p_sw <- ggplot(movers, aes(x = boldness, y = n_switches)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "tomato", linewidth = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  labs(x = "Boldness (move_threshold)", y = "No. patch switches",
       title = "B: Boldness vs. switching frequency")

# Panel C: switching frequency vs final weight  
p_sw_wt <- ggplot(movers, aes(x = n_switches, y = final_weight)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "mediumpurple", linewidth = 0.8) +
  labs(x = "No. patch switches", y = "Final weight (g)",
       title = "C: Switching frequency vs. end-of-season weight")

cowplot::plot_grid(p_wt, p_sw, p_sw_wt, ncol = 1)

}

```




## Emergent population outcomes

### Shared setup

Shared palette, labels, and data prep used across all population-level plots.

```{r pop-setup}
library(ggridges)

strategy_pal <- c(
  resident_warm = "tomato",
  resident_cold = "steelblue",
  optimal_mover = "darkgreen"
)

strategy_labs <- c(
  resident_warm = "Resident (warm)",
  resident_cold = "Resident (cold)",
  optimal_mover = "Optimal mover"
)

# Integer age class (0 = age-0, 1 = age-1, ...)
ibm_long <- ibm_long |>
  mutate(age_class = floor(age))

# Annual October 1 census: alive fish only
census <- ibm_long |>
  filter(month(date) == 10, day(date) == 1, survived == 1) |>
  mutate(year = year(date))

# Annual totals from census
census_summary <- census |>
  group_by(year, strategy) |>
  summarise(
    n_total   = n(),
    biomass_g = sum(weight),
    mean_wt   = mean(weight),
    .groups   = "drop"
  ) |>
  mutate(strategy_lbl = recode(strategy, !!!strategy_labs))

# Age-structured counts from census (cap display at Age-5+)
census_age <- census |>
  mutate(
    age_class_f = factor(
      paste0("Age-", pmin(age_class, 5L)),
      levels = paste0("Age-", 0:5)
    ),
    strategy_lbl = recode(strategy, !!!strategy_labs)
  ) |>
  group_by(year, strategy, strategy_lbl, age_class_f) |>
  summarise(n = n(), .groups = "drop")

# Year-over-year finite rate of increase (lambda) from census totals
lambda_df <- census_summary |>
  arrange(strategy, year) |>
  group_by(strategy, strategy_lbl) |>
  mutate(lambda = n_total / lag(n_total)) |>
  filter(!is.na(lambda))

# Apparent annual survival by age class, tracked by cohort through the census
cohort_census <- census |>
  group_by(strategy, cohort, age_class) |>
  summarise(n = n(), .groups = "drop") |>
  arrange(strategy, cohort, age_class)

apparent_survival <- cohort_census |>
  group_by(strategy, cohort) |>
  mutate(
    n_next    = lead(n),
    surv_rate = n_next / n
  ) |>
  filter(!is.na(surv_rate), age_class <= 8) |>
  mutate(strategy_lbl = recode(strategy, !!!strategy_labs))

# Cohort survivorship (lx): fraction of each cohort surviving to each age class
cohort_max_age <- ibm_long |>
  group_by(strategy, cohort, pid) |>
  summarise(max_age = max(age), .groups = "drop") |>
  group_by(strategy, cohort) |>
  mutate(cohort_n0 = n()) |>
  ungroup()

lx_df <- cohort_max_age |>
  crossing(age_class = 0:12) |>
  group_by(strategy, cohort, age_class) |>
  summarise(
    lx = sum(max_age > age_class) / first(cohort_n0),
    .groups = "drop"
  ) |>
  filter(lx > 0) |>
  mutate(strategy_lbl = recode(strategy, !!!strategy_labs))

# Daily summary by strategy (alive fish only)
ibm_summary <- ibm_long |>
  filter(survived == 1) |>
  group_by(strategy, dayofsim, date) |>
  summarise(
    n_alive        = n(),
    mean_weight    = mean(weight, na.rm = TRUE),
    sd_weight      = sd(weight,   na.rm = TRUE),
    mean_condition = mean(condition, na.rm = TRUE),
    prop_warm      = mean(patch == "warm", na.rm = TRUE),
    .groups        = "drop"
  )
```


### Daily time series

#### Abundance

Daily abundance by strategy. Recruitment pulses from new cohorts are visible as sudden jumps.

```{r fig.width=9, fig.height=4}
ibm_summary |>
  mutate(strategy_lbl = recode(strategy, !!!strategy_labs)) |>
  ggplot(aes(x = date, y = n_alive, color = strategy_lbl)) +
  geom_rect(
    data = summer_shading,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = "grey80", alpha = 0.4, inherit.aes = FALSE
  ) +
  geom_line(linewidth = 0.6) +
  geom_vline(xintercept = exp_start_date, linetype = "dashed") +
  scale_color_manual(values = setNames(strategy_pal, strategy_labs), name = NULL) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(x = NULL, y = "Number alive", title = "Daily abundance by strategy") +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "top")
```

#### Biomass

Daily biomass (total weight of all living individuals), a combined signal of abundance and mean body size.

```{r fig.width=9, fig.height=4}
ibm_summary |>
  mutate(
    strategy_lbl = recode(strategy, !!!strategy_labs),
    biomass_kg   = (n_alive * mean_weight) / 1000
  ) |>
  ggplot(aes(x = date, y = biomass_kg, color = strategy_lbl)) +
  geom_rect(
    data = summer_shading,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = "grey80", alpha = 0.4, inherit.aes = FALSE
  ) +
  geom_line(linewidth = 0.6) +
  geom_vline(xintercept = exp_start_date, linetype = "dashed") +
  scale_color_manual(values = setNames(strategy_pal, strategy_labs), name = NULL) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(x = NULL, y = "Biomass (kg)", title = "Daily biomass by strategy") +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "top")
```

#### Daily age-class abundance

Stacked area chart of daily abundance broken down by integer age class. This reveals the seasonal pulse structure of each cohort as it ages through the population.

```{r fig.width=9, fig.height=8}
# Compute daily alive counts per strategy × age class (down-sampled to weekly)
daily_age <- ibm_long |>
  filter(survived == 1, dayofsim %% 7 == 1) |>
  mutate(
    age_class_f  = factor(paste0("Age-", pmin(age_class, 5L)), levels = paste0("Age-", 0:5)),
    strategy_lbl = recode(strategy, !!!strategy_labs)
  ) |>
  count(date, strategy_lbl, age_class_f)

age_pal <- c(
  "Age-0" = "#d4e6f1", "Age-1" = "#5dade2", "Age-2" = "#1a5276",
  "Age-3" = "#a9dfbf", "Age-4" = "#1e8449", "Age-5" = "#784212"
)

ggplot(daily_age, aes(x = date, y = n, fill = age_class_f)) +
  geom_area(position = "stack", alpha = 0.85) +
  geom_vline(xintercept = exp_start_date, linetype = "dashed") +
  scale_fill_manual(values = age_pal, name = "Age class") +
  facet_wrap(~strategy_lbl, ncol = 1, scales = "free_y") +
  labs(x = NULL, y = "Number alive", title = "Daily abundance by age class") +
  theme_bw() +
  theme(
    panel.grid   = element_blank(),
    strip.text   = element_text(face = "bold"),
    legend.position = "right"
  )
```


### Annual census (October 1)

#### Abundance and biomass

```{r fig.width=9, fig.height=6}
p_ann_abund <- ggplot(
    census_summary |> filter(strategy != "resident_warm"),
    aes(x = year, y = n_total, color = strategy_lbl, group = strategy_lbl)
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed") +
  scale_color_manual(values = setNames(strategy_pal, strategy_labs), name = NULL) +
  # scale_x_continuous(breaks = seq(2001, 2020, 2)) +
  labs(x = NULL, y = "Count", title = "Annual abundance (October 1 census)") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), legend.position = "top")

p_ann_bio <- ggplot(
    census_summary |> filter(strategy != "resident_warm"),
    aes(x = year, y = biomass_g / 1000, color = strategy_lbl, group = strategy_lbl)
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed") +
  scale_color_manual(values = setNames(strategy_pal, strategy_labs), name = NULL) +
  # scale_x_continuous(breaks = seq(2001, 2020, 2)) +
  labs(x = "Year", y = "Biomass (kg)", title = "Annual biomass (October 1 census)") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), legend.position = "none")

p_ann_abund / p_ann_bio
```

#### Age-structured abundance

Stacked bar chart of the October 1 census broken down by integer age class, showing how population age structure varies across years and strategies.

```{r fig.width=9, fig.height=7}
age_pal <- c(
  "Age-0" = "#d4e6f1", "Age-1" = "#5dade2", "Age-2" = "#1a5276",
  "Age-3" = "#a9dfbf", "Age-4" = "#1e8449", "Age-5" = "#784212"
)

census_age |>
  filter(strategy != "resident_warm") |>
  ggplot(aes(x = year, y = n, fill = age_class_f)) +
  geom_col(width = 0.85) +
  scale_fill_manual(values = age_pal, name = "Age class") +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed") +
  # scale_x_continuous(breaks = seq(2001, 2020, 2)) +
  facet_wrap(~strategy_lbl, ncol = 1, scales = "free_y") +
  labs(
    x = "Year", y = "Count",
    title = "Annual age-structured abundance (October 1)"
  ) +
  theme_bw() +
  theme(panel.grid = element_blank(), strip.text = element_text(face = "bold"))
```


### Weight-at-age curves

Von Bertalanffy-style mean weight-at-age by cohort and strategy. Each line traces one cohort from birth to its last observed age class in the census. Individual variation around the mean (±1 SD) is shown as ribbons.

```{r fig.width=10, fig.height=5}
# Monthly snapshots: mean and SD of weight within cohort × age class
waa <- ibm_long |>
  filter(day(date) == 1) |>
  group_by(strategy, cohort, age_class) |>
  summarise(
    mean_wt = mean(weight),
    sd_wt   = sd(weight, na.rm = TRUE),
    age_mid = mean(age),
    .groups = "drop"
  ) |>
  mutate(
    strategy_lbl = recode(strategy, !!!strategy_labs),
    cohort_f     = factor(cohort)
  )

ggplot(waa |> filter(strategy != "resident_warm"),
       aes(x = age_mid, y = mean_wt, group = cohort_f, color = cohort_f)) +
  geom_ribbon(
    aes(ymin = mean_wt - sd_wt, ymax = mean_wt + sd_wt, fill = cohort_f),
    alpha = 0.08, color = NA
  ) +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  scale_color_viridis_d(option = "turbo", name = "Cohort") +
  scale_fill_viridis_d(option = "turbo", guide = "none") +
  facet_wrap(~strategy_lbl) +
  labs(
    x     = "Age (years)",
    y     = "Mean weight (g)",
    title = "Weight-at-age by cohort and strategy",
    subtitle = "Ribbons = ±1 SD; each line = one cohort"
  ) +
  theme_bw() +
  theme(panel.grid = element_blank())
```


### Mean weight at age over time

Plotting mean weight within each age class across census years (not by cohort) reveals whether growth rates have changed over time — in particular, whether increasing population size has suppressed individual growth (density dependence). The decline from early to later years is especially striking in young age classes.

```{r fig.width=10, fig.height=6}
waa_annual <- census |>
  filter(strategy != "resident_warm") |>
  mutate(strategy_lbl = recode(strategy, !!!strategy_labs)) |>
  group_by(year, strategy, strategy_lbl, age_class) |>
  summarise(
    mean_wt = mean(weight),
    sd_wt   = sd(weight),
    n       = n(),
    .groups = "drop"
  ) |>
  filter(n >= 3, age_class <= 7) |>
  left_join(
    census_summary |> select(year, strategy, n_total),
    by = c("year", "strategy")
  )

waa_annual |>
  mutate(age_class_f = paste0("Age-", age_class)) |>
  ggplot(aes(x = year, y = mean_wt, color = strategy_lbl, group = strategy_lbl)) +
  geom_ribbon(
    aes(ymin = mean_wt - sd_wt, ymax = mean_wt + sd_wt, fill = strategy_lbl),
    alpha = 0.15, color = NA
  ) +
  geom_line(linewidth = 0.7) +
  geom_point(aes(size = n), alpha = 0.8) +
  scale_color_manual(values = setNames(strategy_pal, strategy_labs), name = NULL) +
  scale_fill_manual(values  = setNames(strategy_pal, strategy_labs), guide = "none") +
  scale_size_continuous(range = c(1, 4), name = "n fish", guide = "none") +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed") +
  # scale_x_continuous(breaks = seq(2001, 2020, 3)) +
  facet_wrap(~age_class_f, scales = "free_y", ncol = 4) +
  labs(
    x        = "Census year",
    y        = "Mean weight (g)",
    title    = "Mean weight at age over time (October 1 census)",
    subtitle = "Ribbon = ±1 SD  |  Point size = sample size"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    strip.text       = element_text(face = "bold"),
    legend.position  = "top"
  )
```

The sharp drop in mean weight-at-age between 2001–2003 coincides with the first recruitment pulse expanding the optimal mover population from ~25 to ~600 fish. To assess whether this reflects density-dependent growth suppression rather than cohort effects, we can plot mean weight at age directly against total abundance:

```{r fig.width=8, fig.height=8}
waa_annual |>
  filter(strategy == "optimal_mover", age_class <= 4) |>
  mutate(age_class_f = paste0("Age-", age_class)) |>
  ggplot(aes(x = n_total, y = mean_wt)) +
  geom_smooth(method = "lm", se = TRUE, color = "grey40", linewidth = 0.7, alpha = 0.2) +
  geom_point(aes(color = year), size = 3, alpha = 0.9) +
  scale_color_viridis_c(option = "plasma", name = "Year") +
  facet_wrap(~age_class_f, scales = "free", ncol = 2) +
  labs(
    x        = "Total population size at October 1 census",
    y        = "Mean weight (g)",
    title    = "Density–growth relationship by age class (Optimal mover)",
    subtitle = "Point colour = year  |  Line = OLS fit"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```


### Weight frequency distributions

Ridgeline plots of body-size distributions at each October 1 census. Shifts in modal size and the broadening of distributions over time reflect multi-cohort structure.

```{r fig.width=9, fig.height=9}
census |>
  filter(strategy != "resident_warm") |>
  mutate(
    strategy_lbl = recode(strategy, !!!strategy_labs),
    year_f       = factor(year, levels = rev(sort(unique(year))))
  ) |>
  ggplot(aes(x = weight, y = year_f, fill = strategy_lbl)) +
  geom_density_ridges(
    alpha = 0.6, scale = 1.5,
    rel_min_height = 0.01,
    bandwidth = 5
  ) +
  scale_fill_manual(values = setNames(strategy_pal, strategy_labs), name = NULL) +
  facet_wrap(~strategy_lbl, scales = "free_x") +
  labs(
    x     = "Weight (g)",
    y     = "Year",
    title = "Weight frequency distributions at October 1 census"
  ) +
  theme_bw() +
  theme(
    panel.grid  = element_blank(),
    strip.text  = element_text(face = "bold"),
    legend.position = "none"
  )
```


### Annual survival

#### Apparent survival between age classes

Annual apparent survival rates from the October 1 census, calculated as the ratio of fish in age class *x+1* in year *t+1* to age class *x* in year *t* within the same cohort. Rates above 1.0 can indicate new recruits entering an age class or census error.

```{r fig.width=9, fig.height=6}
apparent_survival |>
  filter(strategy != "resident_warm") |>
  ggplot(aes(
    x    = factor(age_class),
    y    = surv_rate,
    fill = strategy_lbl
  )) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey50") +
  geom_boxplot(alpha = 0.6, outlier.shape = 21, outlier.size = 1.5, width = 0.5) +
  scale_fill_manual(values = setNames(strategy_pal, strategy_labs), name = NULL) +
  scale_y_log10(labels = scales::label_number(accuracy = 0.01)) +
  facet_wrap(~strategy_lbl) +
  labs(
    x        = "Age class (from)",
    y        = "Apparent annual survival (log scale)",
    title    = "Apparent survival between consecutive age classes",
    subtitle = "Dashed line = survival = 1.0  |  each box aggregates all cohorts"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"),
        legend.position = "none")
```


### Population growth rate: lambda and intrinsic rate (r)

#### Realised annual lambda

Annual finite rate of increase (λ = N~t+1~ / N~t~) from the October 1 census. Values above 1 indicate population growth; below 1, decline.

```{r fig.width=9, fig.height=5}
ggplot(
    lambda_df |> filter(strategy != "resident_warm"),
    aes(x = year, y = lambda, color = strategy_lbl, group = strategy_lbl)
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey40") +
  geom_line(linewidth = 0.8, alpha = 0.8) +
  geom_point(size = 2, alpha = 0.9) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed") +
  scale_color_manual(values = setNames(strategy_pal, strategy_labs), name = NULL) +
  scale_y_log10(labels = scales::label_number(accuracy = 0.1)) +
  # scale_x_continuous(breaks = seq(2002, 2020, 2)) +
  labs(
    x        = "Year",
    y        = "λ (log scale)",
    title    = "Annual finite rate of increase (October 1 census)",
    subtitle = "Dashed line = λ = 1 (stable population)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), legend.position = "top")
```


#### Intrinsic rate of increase (r) via Euler–Lotka equation

The **intrinsic rate of natural increase** (r, or equivalently λ = e^r^) is the theoretical per-capita growth rate a population would achieve if its current age-specific vital rates (lx, mx) remained constant with no density dependence — i.e., at zero density. This is the life-history analogue of the α parameter in a Ricker stock-recruitment model: the slope of recruits per spawner as spawner density approaches zero.

The standard approach is the **Euler–Lotka equation**, which finds the r that satisfies:

$$\sum_{x=0}^{\infty} e^{-rx} \cdot l_x \cdot m_x = 1$$

where lx is survivorship to age class x, and mx is mean per-capita offspring production at age x (averaging spawning probability × fecundity across all fish of that age, including non-spawners).

A key estimation challenge here is that lx and mx are not density-independent in this IBM — fish that grow at high density are smaller, less likely to spawn, and less likely to survive. Using vital rates averaged over all simulation years (including high-density years) will produce a **downward-biased** estimate of r, while using only the founding 2001 cohort (which experienced near-zero density) gives the closest approximation to r_max.

```{r}
# Annotate each spawn event with parent age class and year
spawn_annotated <- spawn_log |>
  left_join(fish_registry |> select(pid, strategy), by = c("parent_pid" = "pid")) |>
  left_join(ibm_long |> select(pid, dayofsim, age, date),
            by = c("parent_pid" = "pid", "dayofsim")) |>
  mutate(spawn_year = year(date), age_class = floor(age)) |>
  filter(strategy != "resident_warm")

# mx: per-capita annual offspring production at each age class
# Numerator: offspring from fish of age x in year t
# Denominator: fish of age x in October 1 census of year t
offspring_by_age <- spawn_annotated |>
  group_by(strategy, spawn_year, age_class) |>
  summarise(total_offspring = sum(n_offspring), .groups = "drop")

census_age_n <- census |>
  filter(strategy != "resident_warm") |>
  count(strategy, year, age_class)

mx_data <- offspring_by_age |>
  left_join(census_age_n, by = c("strategy", "spawn_year" = "year", "age_class")) |>
  filter(!is.na(n)) |>
  mutate(mx_annual = total_offspring / n) |>
  group_by(strategy, age_class) |>
  summarise(mx = mean(mx_annual), .groups = "drop")

# Solve Euler–Lotka equation numerically
euler_lotka_r <- function(lt) {
  R0    <- sum(lt$lx * lt$mx)
  T_gen <- sum(lt$age_class * lt$lx * lt$mx) / R0
  if (R0 <= 1) return(tibble(r = NA_real_, lambda_r = NA_real_, R0 = R0, T_gen = T_gen))
  fn    <- function(r) sum(exp(-r * lt$age_class) * lt$lx * lt$mx) - 1
  r_hat <- uniroot(fn, c(-3, 10), tol = 1e-8)$root
  tibble(r = r_hat, lambda_r = exp(r_hat), R0 = R0, T_gen = T_gen)
}

# Version 1: average lx across all cohorts (includes density-suppressed years)
lx_avg <- lx_df |>
  filter(strategy != "resident_warm") |>
  group_by(strategy, strategy_lbl, age_class) |>
  summarise(lx = mean(lx), .groups = "drop")

life_table_avg <- lx_avg |>
  left_join(mx_data |> select(strategy, age_class, mx), by = c("strategy", "age_class")) |>
  replace_na(list(mx = 0)) |>
  filter(lx > 0)

# Version 2: 2001 founder cohort lx only (near-zero-density approximation of r_max)
lx_2001 <- lx_df |>
  filter(strategy != "resident_warm", cohort == 2001) |>
  select(strategy, strategy_lbl, age_class, lx)

life_table_2001 <- lx_2001 |>
  left_join(mx_data |> select(strategy, age_class, mx), by = c("strategy", "age_class")) |>
  replace_na(list(mx = 0)) |>
  filter(lx > 0)

r_avg <- life_table_avg |>
  nest(.by = c(strategy, strategy_lbl)) |>
  mutate(result = map(data, euler_lotka_r), lx_source = "All cohorts (avg lx)") |>
  unnest(result) |> select(-data)

r_2001 <- life_table_2001 |>
  nest(.by = c(strategy, strategy_lbl)) |>
  mutate(result = map(data, euler_lotka_r), lx_source = "2001 cohort lx (r_max approx)") |>
  unnest(result) |> select(-data)

r_estimates <- bind_rows(r_avg, r_2001) |>
  mutate(across(c(R0, T_gen, r, lambda_r), \(x) round(x, 2)))

r_estimates |> select(lx_source, strategy_lbl, R0, T_gen, r, lambda_r) |>
  arrange(lx_source, strategy_lbl)
```

**Life-table reproductive output (lx × mx) by age class:**

```{r fig.width=9, fig.height=4}
life_table_avg |>
  mutate(lx_mx = lx * mx) |>
  group_by(strategy) |>
  mutate(lx_scaled = lx * max(lx_mx) / max(lx)) |>
  ungroup() |>
  ggplot(aes(x = age_class)) +
  geom_col(aes(y = lx_mx, fill = strategy_lbl), show.legend = FALSE) +
  geom_line(aes(y = lx_scaled), linetype = "dashed", color = "grey30", linewidth = 0.8) +
  scale_fill_manual(values = setNames(strategy_pal, strategy_labs)) +
  scale_x_continuous(breaks = 0:15) +
  facet_wrap(~strategy_lbl, scales = "free_y") +
  labs(
    x        = "Age class",
    y        = "lx × mx  (expected offspring per individual)",
    title    = "Life-table reproductive output",
    subtitle = "Bars = lx·mx  |  Dashed = lx survivorship (right-scaled)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```


#### Spawner–recruit relationship and Ricker α

A complementary approach to estimating the intrinsic growth rate is to fit a **stock-recruitment model** to annual spawner counts and subsequent age-0 recruits. The slope of the spawner–recruit curve at the origin (α) is directly analogous to the per-spawner reproductive rate at zero density, and is conceptually equivalent to the Ricker α parameter.

This approach requires sufficient spawner–recruit pairs and variation in spawner density, so it is only feasible for the optimal mover strategy.

```{r}
# Build spawner–recruit pairs: spawners in year t → age-0 recruits in year t+1
spawners_annual <- spawn_annotated |>
  group_by(strategy, spawn_year) |>
  summarise(n_spawners = n_distinct(parent_pid), total_offspring = sum(n_offspring),
            .groups = "drop")

recruits_annual <- census |>
  filter(age_class == 0) |>
  count(strategy, year, name = "n_recruits")

sr_data <- spawners_annual |>
  mutate(recruit_year = spawn_year + 1) |>
  left_join(recruits_annual, by = c("strategy", "recruit_year" = "year")) |>
  filter(!is.na(n_recruits), strategy %in% c("optimal_mover", "resident_cold"))

om_sr <- sr_data |> filter(strategy == "optimal_mover")

# Fit Beverton-Holt and Ricker
bh_fit     <- nls(n_recruits ~ (alpha * n_spawners) / (1 + n_spawners / K),
                  data = om_sr, start = list(alpha = 5, K = 500))
ricker_fit <- nls(n_recruits ~ alpha * n_spawners * exp(-beta * n_spawners),
                  data = om_sr, start = list(alpha = 5, beta = 0.001))
```

```{r fig.width=6, fig.height=5}
s_seq   <- seq(0, max(om_sr$n_spawners) * 1.1, length.out = 300)
pred_df <- bind_rows(
  tibble(n_spawners = s_seq,
         n_recruits = predict(bh_fit,     newdata = list(n_spawners = s_seq)),
         model = sprintf("Beverton-Holt  (α = %.0f)", coef(bh_fit)["alpha"])),
  tibble(n_spawners = s_seq,
         n_recruits = predict(ricker_fit, newdata = list(n_spawners = s_seq)),
         model = sprintf("Ricker  (α = %.0f)", coef(ricker_fit)["alpha"]))
)

ggplot(om_sr, aes(x = n_spawners, y = n_recruits)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "grey60") +
  geom_line(data = pred_df, aes(color = model), linewidth = 1) +
  geom_point(aes(fill = factor(spawn_year)), size = 3, shape = 21, stroke = 0.3) +
  scale_color_manual(values = c("steelblue", "tomato"), name = "Model") +
  scale_fill_viridis_d(option = "plasma", name = "Year") +
  scale_x_continuous(limits = c(0, NA)) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(
    x        = "Spawners in year t",
    y        = "Age-0 recruits in year t+1",
    title    = "Spawner–recruit relationship (Optimal mover)",
    subtitle = "Dotted line = 1:1  |  α = recruits per spawner at zero density"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), legend.position = "right")
```


### Cohort survivorship curves

Fraction of each cohort still alive at each integer age class (lx curves). These visualise variation in cohort longevity across years and strategies.

```{r fig.width=9, fig.height=5}
lx_df |>
  filter(strategy != "resident_warm") |>
  mutate(cohort_f = factor(cohort)) |>
  ggplot(aes(
    x     = age_class,
    y     = lx,
    group = cohort_f,
    color = cohort_f
  )) +
  geom_line(linewidth = 0.6, alpha = 0.8) +
  scale_color_viridis_d(option = "turbo", name = "Cohort") +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  scale_x_continuous(breaks = 0:12) +
  facet_wrap(~strategy_lbl) +
  labs(
    x        = "Age class",
    y        = "Fraction of cohort surviving",
    title    = "Cohort survivorship curves (lx)",
    subtitle = "Each line = one birth cohort"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```



## Habitat value

```{r}
#| label: hab-value-prep

# Calendar year used throughout this section.
# For seasonal plots, January and February are assigned to the preceding
# December's year (meteorological winter convention), so that
# Winter YYYY = Dec YYYY + Jan/Feb YYYY+1 forms a contiguous season.
hab_value <- ibm_long |>
  filter(!is.na(ggd)) |>
  mutate(
    cal_year       = year(date),
    daily_growth_g = weight * ggd   # specific rate × mass = g/d
  )

# Patch colour palette
patch_pal <- c(warm = "tomato", cold = "steelblue")

# Annual mean habitat area per calendar year
annual_area <- habitat_df |>
  group_by(cal_year = year(date)) |>
  summarise(A_warm = mean(A_warm), A_cold = mean(A_cold), .groups = "drop")

# Seasonal mean habitat area — meteorological winter convention
seasonal_area <- habitat_df |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, year(date) - 1L, year(date))
  ) |>
  group_by(season_year, season) |>
  summarise(A_warm = mean(A_warm), A_cold = mean(A_cold), .groups = "drop")
```

All plots show results for `optimal_mover` individuals only, as resident strategies cannot use alternative habitats and therefore do not illustrate habitat-switching value. The vertical dashed line marks the start of the climate-change scenario (`r year(exp_start_date)`), after which cold habitat area ramps down and warm habitat area ramps up. Equivalence curve x-axes show cold habitat as a proportion of total habitat (50% = equal areas; 0% = no cold habitat remaining). Seasonal plots use the meteorological winter convention: Winter YYYY = December YYYY + January/February YYYY+1.


### Occupancy

Proportion of days spent in warm habitat, averaged across individuals. Because occupancy is a single proportion (not a per-patch quantity), the equivalence curves use a different format: x = proportion of total habitat that is *warm*, y = proportion of *time* in warm habitat, with the diagonal (y = x) marking proportional/unbiased use.

#### Annual

```{r}
#| label: p-occupancy-annual
#| fig.width: 8
#| fig.height: 4

hab_value |>
  filter(strategy == "optimal_mover") |>
  group_by(pid, cal_year) |>
  summarise(prop_warm = mean(patch == "warm"), .groups = "drop") |>
  group_by(cal_year) |>
  summarise(
    mean_prop = mean(prop_warm),
    sd_prop   = sd(prop_warm),
    .groups   = "drop"
  ) |>
  ggplot(aes(x = cal_year, y = mean_prop)) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed", color = "grey50") +
  geom_ribbon(aes(ymin = pmax(mean_prop - sd_prop, 0),
                  ymax = pmin(mean_prop + sd_prop, 1)),
              alpha = 0.2, fill = "tomato") +
  geom_line(linewidth = 0.8, color = "tomato") +
  geom_point(size = 2, color = "tomato") +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1), limits = c(0, 1)) +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 8)) +
  labs(
    x       = "Year",
    y       = "Proportion of year in warm habitat",
    title   = "Annual warm habitat occupancy (optimal movers)",
    caption = "Band = ±1 SD (among-individual variation)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

#### Seasonal

```{r}
#| label: p-occupancy-seasonal
#| fig.width: 8
#| fig.height: 7

hab_value |>
  filter(strategy == "optimal_mover") |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, cal_year - 1L, cal_year)
  ) |>
  group_by(pid, season_year, season) |>
  summarise(prop_warm = mean(patch == "warm"), .groups = "drop") |>
  group_by(season_year, season) |>
  summarise(
    mean_prop = mean(prop_warm),
    sd_prop   = sd(prop_warm),
    .groups   = "drop"
  ) |>
  ggplot(aes(x = season_year, y = mean_prop)) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed", color = "grey50") +
  geom_ribbon(aes(ymin = pmax(mean_prop - sd_prop, 0),
                  ymax = pmin(mean_prop + sd_prop, 1)),
              alpha = 0.2, fill = "tomato") +
  geom_line(linewidth = 0.8, color = "tomato") +
  geom_point(size = 2, color = "tomato") +
  facet_wrap(~season, ncol = 2) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1), limits = c(0, 1)) +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 5)) +
  labs(
    x       = "Year",
    y       = "Proportion of season in warm habitat",
    title   = "Seasonal warm habitat occupancy (optimal movers)",
    caption = "Band = ±1 SD (among-individual variation)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```

#### Annual equivalence curve

```{r}
#| label: p-occupancy-equiv
#| fig.width: 8
#| fig.height: 5

hab_value |>
  filter(strategy == "optimal_mover") |>
  group_by(pid, cal_year) |>
  summarise(prop_warm = mean(patch == "warm"), .groups = "drop") |>
  group_by(cal_year) |>
  summarise(mean_prop_warm = mean(prop_warm), .groups = "drop") |>
  left_join(annual_area |> mutate(prop_warm_area = A_warm / (A_warm + A_cold)),
            by = "cal_year") |>
  ggplot(aes(x = prop_warm_area, y = mean_prop_warm)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  geom_path(aes(color = cal_year), linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = cal_year), size = 2.5) +
  annotate("text", x = 0.52, y = 0.45, label = "Cold preference\n(below line)",
           hjust = 0, size = 3, color = "steelblue4") +
  annotate("text", x = 0.52, y = 0.92, label = "Warm preference\n(above line)",
           hjust = 0, size = 3, color = "tomato4") +
  scale_color_viridis_c(name = "Year", option = "plasma") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0.5, 1, by = 0.1), limits = c(0.5, 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1), limits = c(0, 1)) +
  labs(
    x        = "Warm habitat as proportion of total habitat",
    y        = "Mean proportion of time in warm habitat",
    title    = "Occupancy equivalence curve (optimal movers)",
    subtitle = "Dashed line = proportional use (no preference)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

#### Seasonal equivalence curve

```{r}
#| label: p-occupancy-equiv-seasonal
#| fig.width: 9
#| fig.height: 7

hab_value |>
  filter(strategy == "optimal_mover") |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, cal_year - 1L, cal_year)
  ) |>
  group_by(pid, season_year, season) |>
  summarise(prop_warm = mean(patch == "warm"), .groups = "drop") |>
  group_by(season_year, season) |>
  summarise(mean_prop_warm = mean(prop_warm), .groups = "drop") |>
  left_join(
    seasonal_area |> mutate(prop_warm_area = A_warm / (A_warm + A_cold)),
    by = c("season_year", "season")
  ) |>
  ggplot(aes(x = prop_warm_area, y = mean_prop_warm)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  geom_path(aes(color = season_year), linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = season_year), size = 2) +
  facet_wrap(~season, ncol = 2) +
  scale_color_viridis_c(name = "Year", option = "plasma") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0.5, 1, by = 0.1), limits = c(0.5, 1)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1), limits = c(0, 1)) +
  labs(
    x        = "Warm habitat as proportion of total habitat",
    y        = "Mean proportion of season in warm habitat",
    title    = "Occupancy equivalence curve by season (optimal movers)",
    subtitle = "Dashed line = proportional use (no preference)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```


### Total biomass

Total daily biomass production (`weight × ggd`) summed across all individuals, split by habitat patch. Equivalence curves show (cold_kg − warm_kg); positive values indicate cold habitat produces more total biomass; y = 0 is equivalence. Free y-axes on the seasonal version preserve the absolute magnitude differences across seasons.

#### Annual

```{r}
#| label: p-warm-biomass
#| fig.width: 8
#| fig.height: 4

hab_value |>
  filter(strategy == "optimal_mover") |>
  group_by(patch, cal_year) |>
  summarise(total_kg = sum(daily_growth_g, na.rm = TRUE) / 1000, .groups = "drop") |>
  ggplot(aes(x = cal_year, y = total_kg, color = patch)) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed", color = "grey50") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_color_manual(values = patch_pal, name = "Patch") +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 8)) +
  labs(
    x     = "Year",
    y     = "Total biomass accrued (kg)",
    title = "Annual biomass accrued by habitat patch (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

#### Seasonal

```{r}
#| label: p-warm-biomass-seasonal
#| fig.width: 8
#| fig.height: 7

hab_value |>
  filter(strategy == "optimal_mover") |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, cal_year - 1L, cal_year)
  ) |>
  group_by(patch, season_year, season) |>
  summarise(total_kg = sum(daily_growth_g, na.rm = TRUE) / 1000, .groups = "drop") |>
  ggplot(aes(x = season_year, y = total_kg, color = patch)) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed", color = "grey50") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~season, ncol = 2, scales = "free_y") +
  scale_color_manual(values = patch_pal, name = "Patch") +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 5)) +
  labs(
    x     = "Year",
    y     = "Total biomass accrued (kg)",
    title = "Seasonal biomass accrued by habitat patch (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```

#### Annual equivalence curve

```{r}
#| label: p-biomass-ratio
#| fig.width: 8
#| fig.height: 5

hab_value |>
  filter(strategy == "optimal_mover") |>
  group_by(patch, cal_year) |>
  summarise(total_kg = sum(daily_growth_g, na.rm = TRUE) / 1000, .groups = "drop") |>
  pivot_wider(names_from = patch, values_from = total_kg, values_fill = 0) |>
  mutate(biomass_diff = cold - warm) |>
  left_join(annual_area |> mutate(prop_cold = A_cold / (A_cold + A_warm)),
            by = "cal_year") |>
  ggplot(aes(x = prop_cold, y = biomass_diff)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  0, ymax =  Inf,
           fill = "steelblue", alpha = 0.05) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "tomato", alpha = 0.05) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  geom_path(aes(color = cal_year), linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = cal_year), size = 2.5) +
  annotate("text", x = 0.01, y =  12, label = "Cold better",
           hjust = 0, size = 3, color = "steelblue4") +
  annotate("text", x = 0.01, y = -12, label = "Warm better",
           hjust = 0, size = 3, color = "tomato4") +
  scale_color_viridis_c(name = "Year", option = "plasma") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0, 0.5, by = 0.1), limits = c(0, 0.5)) +
  labs(
    x     = "Cold habitat as proportion of total habitat",
    y     = "Biomass accrued (cold) \u2212 biomass accrued (warm)  (kg)",
    title = "Total biomass equivalence curve (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

#### Seasonal equivalence curve

```{r}
#| label: p-biomass-ratio-seasonal
#| fig.width: 8
#| fig.height: 7

hab_value |>
  filter(strategy == "optimal_mover") |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, cal_year - 1L, cal_year)
  ) |>
  group_by(patch, season_year, season) |>
  summarise(total_kg = sum(daily_growth_g, na.rm = TRUE) / 1000, .groups = "drop") |>
  pivot_wider(names_from = patch, values_from = total_kg, values_fill = 0) |>
  mutate(biomass_diff = cold - warm) |>
  left_join(
    seasonal_area |> mutate(prop_cold = A_cold / (A_cold + A_warm)),
    by = c("season_year", "season")
  ) |>
  ggplot(aes(x = prop_cold, y = biomass_diff)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  0, ymax =  Inf,
           fill = "steelblue", alpha = 0.05) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "tomato", alpha = 0.05) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  geom_path(aes(color = season_year), linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = season_year), size = 2) +
  facet_wrap(~season, ncol = 2, scales = "free_y") +
  scale_color_viridis_c(name = "Year", option = "plasma") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0, 0.5, by = 0.1), limits = c(0, 0.5)) +
  labs(
    x     = "Cold habitat as proportion of total habitat",
    y     = "Biomass accrued (cold) \u2212 biomass accrued (warm)  (kg)",
    title = "Total biomass equivalence curve by season (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```


### Per-area biomass

Total biomass accrued per unit habitat area, revealing how per-unit productivity of each patch changes as habitat availability shifts. Equivalence curve y-axis = (cold kg/area) − (warm kg/area); positive = cold more productive per unit area.

#### Annual

```{r}
#| label: p-per-area-productivity
#| fig.width: 9
#| fig.height: 5

hab_value |>
  filter(strategy == "optimal_mover") |>
  group_by(patch, cal_year) |>
  summarise(total_kg = sum(daily_growth_g, na.rm = TRUE) / 1000, .groups = "drop") |>
  pivot_wider(names_from = patch, values_from = total_kg) |>
  left_join(annual_area, by = "cal_year") |>
  mutate(warm = warm / A_warm, cold = cold / A_cold) |>
  select(cal_year, warm, cold) |>
  pivot_longer(c(warm, cold), names_to = "patch", values_to = "kg_per_area") |>
  ggplot(aes(x = cal_year, y = kg_per_area, color = patch)) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, color = "grey80") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_color_manual(values = patch_pal, name = "Patch") +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 8)) +
  labs(
    x     = "Year",
    y     = "Biomass accrued per unit area (kg)",
    title = "Annual per-area biomass productivity: warm vs. cold habitat (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

#### Seasonal

```{r}
#| label: p-per-area-productivity-seasonal
#| fig.width: 9
#| fig.height: 7

hab_value |>
  filter(strategy == "optimal_mover") |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, cal_year - 1L, cal_year)
  ) |>
  group_by(patch, season_year, season) |>
  summarise(total_kg = sum(daily_growth_g, na.rm = TRUE) / 1000, .groups = "drop") |>
  pivot_wider(names_from = patch, values_from = total_kg, values_fill = 0) |>
  left_join(seasonal_area, by = c("season_year", "season")) |>
  mutate(warm = warm / A_warm, cold = cold / A_cold) |>
  select(season_year, season, warm, cold) |>
  pivot_longer(c(warm, cold), names_to = "patch", values_to = "kg_per_area") |>
  ggplot(aes(x = season_year, y = kg_per_area, color = patch)) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, color = "grey80") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~season, ncol = 2, scales = "free_y") +
  scale_color_manual(values = patch_pal, name = "Patch") +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 5)) +
  labs(
    x     = "Year",
    y     = "Biomass accrued per unit area (kg)",
    title = "Seasonal per-area biomass productivity: warm vs. cold habitat (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```

#### Annual equivalence curve

```{r}
#| label: p-per-area-equiv
#| fig.width: 8
#| fig.height: 5

hab_value |>
  filter(strategy == "optimal_mover") |>
  group_by(patch, cal_year) |>
  summarise(total_kg = sum(daily_growth_g, na.rm = TRUE) / 1000, .groups = "drop") |>
  pivot_wider(names_from = patch, values_from = total_kg) |>
  left_join(annual_area, by = "cal_year") |>
  mutate(
    diff_per_area = cold / A_cold - warm / A_warm,
    prop_cold     = A_cold / (A_cold + A_warm)
  ) |>
  ggplot(aes(x = prop_cold, y = diff_per_area)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  0, ymax =  Inf,
           fill = "steelblue", alpha = 0.05) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "tomato", alpha = 0.05) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  geom_path(aes(color = cal_year), linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = cal_year), size = 2.5) +
  annotate("text", x = 0.01, y =  14, label = "Cold better",
           hjust = 0, size = 3, color = "steelblue4") +
  annotate("text", x = 0.01, y = -35, label = "Warm better",
           hjust = 0, size = 3, color = "tomato4") +
  scale_color_viridis_c(name = "Year", option = "plasma") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0, 0.5, by = 0.1), limits = c(0, 0.5)) +
  labs(
    x     = "Cold habitat as proportion of total habitat",
    y     = "Biomass/area (cold) \u2212 biomass/area (warm)  (kg)",
    title = "Per-area biomass equivalence curve (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

#### Seasonal equivalence curve

```{r}
#| label: p-per-area-equiv-seasonal
#| fig.width: 9
#| fig.height: 7

hab_value |>
  filter(strategy == "optimal_mover") |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, cal_year - 1L, cal_year)
  ) |>
  group_by(patch, season_year, season) |>
  summarise(total_kg = sum(daily_growth_g, na.rm = TRUE) / 1000, .groups = "drop") |>
  pivot_wider(names_from = patch, values_from = total_kg, values_fill = 0) |>
  left_join(
    seasonal_area |> mutate(prop_cold = A_cold / (A_cold + A_warm)),
    by = c("season_year", "season")
  ) |>
  mutate(diff_per_area = cold / A_cold - warm / A_warm) |>
  ggplot(aes(x = prop_cold, y = diff_per_area)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  0, ymax =  Inf,
           fill = "steelblue", alpha = 0.05) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "tomato", alpha = 0.05) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  geom_path(aes(color = season_year), linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = season_year), size = 2) +
  facet_wrap(~season, ncol = 2, scales = "free_y") +
  scale_color_viridis_c(name = "Year", option = "plasma") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0, 0.5, by = 0.1), limits = c(0, 0.5)) +
  labs(
    x     = "Cold habitat as proportion of total habitat",
    y     = "Biomass/area (cold) \u2212 biomass/area (warm)  (kg)",
    title = "Per-area biomass equivalence curve by season (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```


### Growth

Mean specific growth rate (g/g/d) experienced by individuals while occupying each patch. Equivalence curve y-axis = mean ggd (cold) − mean ggd (warm); positive = cold provides higher per-capita growth. Seasonal plots include ±1 SD bands.

#### Annual

```{r}
#| label: p-ggd-annual
#| fig.width: 8
#| fig.height: 4

hab_value |>
  filter(strategy == "optimal_mover") |>
  group_by(patch, cal_year) |>
  summarise(mean_ggd = mean(ggd, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = cal_year, y = mean_ggd, color = patch)) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0, color = "grey80") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_color_manual(values = patch_pal, name = "Patch") +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 8)) +
  labs(
    x     = "Year",
    y     = "Mean specific growth rate (g/g/d)",
    title = "Annual realized growth rate by habitat patch (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

#### Seasonal

```{r}
#| label: p-ggd-seasonal
#| fig.width: 8
#| fig.height: 7

hab_value |>
  filter(strategy == "optimal_mover") |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, cal_year - 1L, cal_year)
  ) |>
  group_by(patch, season_year, season) |>
  summarise(
    mean_ggd = mean(ggd, na.rm = TRUE),
    sd_ggd   = sd(ggd,   na.rm = TRUE),
    .groups  = "drop"
  ) |>
  ggplot(aes(x = season_year, y = mean_ggd, color = patch, fill = patch)) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed", color = "grey50") +
  geom_ribbon(aes(ymin = mean_ggd - sd_ggd, ymax = mean_ggd + sd_ggd),
              alpha = 0.2, color = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~season, ncol = 2) +
  scale_color_manual(values = patch_pal, name = "Patch") +
  scale_fill_manual(values = patch_pal, name = "Patch") +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 5)) +
  labs(
    x       = "Year",
    y       = "Mean specific growth rate (g/g/d)",
    title   = "Seasonal realized growth rate by habitat patch (optimal movers)",
    caption = "Band = ±1 SD (among individual-days)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```

#### Annual equivalence curve

```{r}
#| label: p-ggd-equiv
#| fig.width: 8
#| fig.height: 5

hab_value |>
  filter(strategy == "optimal_mover") |>
  group_by(cal_year, patch) |>
  summarise(mean_ggd = mean(ggd, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = patch, values_from = mean_ggd) |>
  mutate(ggd_diff = cold - warm) |>
  left_join(annual_area |> mutate(prop_cold = A_cold / (A_cold + A_warm)),
            by = "cal_year") |>
  ggplot(aes(x = prop_cold, y = ggd_diff)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  0, ymax =  Inf,
           fill = "steelblue", alpha = 0.05) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "tomato", alpha = 0.05) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  geom_path(aes(color = cal_year), linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = cal_year), size = 2.5) +
  annotate("text", x = 0.01, y =  0.013, label = "Cold better",
           hjust = 0, size = 3, color = "steelblue4") +
  annotate("text", x = 0.01, y = -0.029, label = "Warm better",
           hjust = 0, size = 3, color = "tomato4") +
  scale_color_viridis_c(name = "Year", option = "plasma") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0, 0.5, by = 0.1), limits = c(0, 0.5)) +
  labs(
    x     = "Cold habitat as proportion of total habitat",
    y     = "Mean ggd (cold) \u2212 mean ggd (warm)  (g/g/d)",
    title = "Growth equivalence curve (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

#### Seasonal equivalence curve

```{r}
#| label: p-ggd-equiv-seasonal
#| fig.width: 9
#| fig.height: 7

hab_value |>
  filter(strategy == "optimal_mover") |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, cal_year - 1L, cal_year)
  ) |>
  group_by(season_year, patch, season) |>
  summarise(mean_ggd = mean(ggd, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = patch, values_from = mean_ggd) |>
  mutate(ggd_diff = cold - warm) |>
  left_join(
    seasonal_area |> mutate(prop_cold = A_cold / (A_cold + A_warm)),
    by = c("season_year", "season")
  ) |>
  ggplot(aes(x = prop_cold, y = ggd_diff)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  0, ymax =  Inf,
           fill = "steelblue", alpha = 0.05) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "tomato", alpha = 0.05) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  geom_path(aes(color = season_year), linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = season_year), size = 2) +
  facet_wrap(~season, ncol = 2, scales = "free_y") +
  scale_color_viridis_c(name = "Year", option = "plasma") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0, 0.5, by = 0.1), limits = c(0, 0.5)) +
  labs(
    x     = "Cold habitat as proportion of total habitat",
    y     = "Mean ggd (cold) \u2212 mean ggd (warm)  (g/g/d)",
    title = "Growth equivalence curve by season (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```


### Survival

Mean daily survival probability (`p_survive`) experienced while occupying each patch. Equivalence curve y-axis = mean p_survive (cold) − mean p_survive (warm); positive = cold provides higher daily survival.

#### Annual

```{r}
#| label: p-survival-annual
#| fig.width: 9
#| fig.height: 5

hab_value |>
  filter(strategy == "optimal_mover") |>
  group_by(patch, cal_year) |>
  summarise(mean_psurv = mean(p_survive, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = cal_year, y = mean_psurv, color = patch)) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed", color = "grey50") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_color_manual(values = patch_pal, name = "Patch") +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.01)) +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 8)) +
  labs(
    x     = "Year",
    y     = "Mean daily survival probability",
    title = "Annual mean daily survival probability by habitat patch (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

#### Seasonal

```{r}
#| label: p-survival-seasonal
#| fig.width: 9
#| fig.height: 7

hab_value |>
  filter(strategy == "optimal_mover") |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, cal_year - 1L, cal_year)
  ) |>
  group_by(patch, season_year, season) |>
  summarise(mean_psurv = mean(p_survive, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = season_year, y = mean_psurv, color = patch)) +
  geom_vline(xintercept = year(exp_start_date), linetype = "dashed", color = "grey50") +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.5) +
  facet_wrap(~season, ncol = 2, scales = "free_y") +
  scale_color_manual(values = patch_pal, name = "Patch") +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.1)) +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 5)) +
  labs(
    x     = "Year",
    y     = "Mean daily survival probability",
    title = "Seasonal mean daily survival probability by habitat patch (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```

#### Annual equivalence curve

```{r}
#| label: p-survival-equiv
#| fig.width: 8
#| fig.height: 5

hab_value |>
  filter(strategy == "optimal_mover") |>
  group_by(cal_year, patch) |>
  summarise(mean_psurv = mean(p_survive, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = patch, values_from = mean_psurv) |>
  mutate(psurv_diff = cold - warm) |>
  left_join(annual_area |> mutate(prop_cold = A_cold / (A_cold + A_warm)),
            by = "cal_year") |>
  ggplot(aes(x = prop_cold, y = psurv_diff)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  0, ymax =  Inf,
           fill = "steelblue", alpha = 0.05) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "tomato", alpha = 0.05) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  geom_path(aes(color = cal_year), linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = cal_year), size = 2.5) +
  annotate("text", x = 0.01, y =  0.005, label = "Cold better",
           hjust = 0, size = 3, color = "steelblue4") +
  annotate("text", x = 0.01, y = -0.040, label = "Warm better",
           hjust = 0, size = 3, color = "tomato4") +
  scale_color_viridis_c(name = "Year", option = "plasma") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0, 0.5, by = 0.1), limits = c(0, 0.5)) +
  labs(
    x     = "Cold habitat as proportion of total habitat",
    y     = "Mean p_survive (cold) \u2212 mean p_survive (warm)",
    title = "Survival equivalence curve (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

#### Seasonal equivalence curve

```{r}
#| label: p-survival-equiv-seasonal
#| fig.width: 9
#| fig.height: 7

hab_value |>
  filter(strategy == "optimal_mover") |>
  mutate(
    season = factor(
      case_when(
        month(date) %in% 3:5  ~ "Spring (Mar–May)",
        month(date) %in% 6:8  ~ "Summer (Jun–Aug)",
        month(date) %in% 9:11 ~ "Autumn (Sep–Nov)",
        TRUE                  ~ "Winter (Dec–Feb)"
      ),
      levels = c("Spring (Mar–May)", "Summer (Jun–Aug)",
                 "Autumn (Sep–Nov)", "Winter (Dec–Feb)")
    ),
    season_year = if_else(month(date) %in% 1:2, cal_year - 1L, cal_year)
  ) |>
  group_by(season_year, patch, season) |>
  summarise(mean_psurv = mean(p_survive, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = patch, values_from = mean_psurv) |>
  mutate(psurv_diff = cold - warm) |>
  left_join(
    seasonal_area |> mutate(prop_cold = A_cold / (A_cold + A_warm)),
    by = c("season_year", "season")
  ) |>
  ggplot(aes(x = prop_cold, y = psurv_diff)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  0, ymax =  Inf,
           fill = "steelblue", alpha = 0.05) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0,
           fill = "tomato", alpha = 0.05) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  geom_path(aes(color = season_year), linewidth = 0.6, alpha = 0.7) +
  geom_point(aes(color = season_year), size = 2) +
  facet_wrap(~season, ncol = 2, scales = "free_y") +
  scale_color_viridis_c(name = "Year", option = "plasma") +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0, 0.5, by = 0.1), limits = c(0, 0.5)) +
  labs(
    x     = "Cold habitat as proportion of total habitat",
    y     = "Mean p_survive (cold) \u2212 mean p_survive (warm)",
    title = "Survival equivalence curve by season (optimal movers)"
  ) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), strip.text = element_text(face = "bold"))
```

## OLD



When are fish dying and how large are they/how fast are they growing relative to the survivor mean?
```{r eval=FALSE}
n_days <- ncol(mass_matrix)

# Per-fish fate summary (used for panel C)
fish_fate <- fish_registry %>%
  mutate(
    survived   = if_else(!is.na(mass_matrix[pid, n_days]), "Survived", "Died"),
    days_alive = rowSums(!is.na(mass_matrix))
  )

# fish_fate %>% count(strategy, survived)
```

```{r  eval=FALSE, fig.height=7, fig.width=5}
strategy_order <- c("resident_warm", "resident_cold", "optimal_mover")

# Panels A & B: compare dying fish to surviving fish on the same day
# -----------------------------------------------------------------------
# Dying fish: weight and growth rate on their day of death (survived == 0)
death_records <- ibm_long %>%
  filter(survived == 0) %>%
  select(pid, strategy, date, dayofsim, weight, ggd) %>%
  mutate(fate = "Died")

# Death days across all strategies
death_days <- death_records %>% distinct(dayofsim)

# Surviving fish (those alive at end): weight and growth on those same death days
survivor_pids <- fish_fate %>% filter(survived == "Survived") %>% pull(pid)

survivor_records <- ibm_long %>%
  filter(pid %in% survivor_pids) %>%
  inner_join(death_days, by = "dayofsim") %>%
  select(pid, strategy, dayofsim, weight, ggd) %>%
  mutate(fate = "Survived")

# Mean of ALL survivors (across strategies) on each death day
survivor_means <- survivor_records %>%
  group_by(dayofsim) %>%
  summarise(mean_wt_surv  = mean(weight, na.rm = TRUE),
            mean_ggd_surv = mean(ggd,    na.rm = TRUE),
            .groups = "drop")

# Difference: dying fish value minus all-survivor mean on same day
df_diff <- death_records %>%
  left_join(survivor_means, by = "dayofsim") %>%
  mutate(
    wt_diff  = weight - mean_wt_surv,
    ggd_diff = ggd    - mean_ggd_surv,
    strategy = factor(strategy, levels = strategy_order)
  )

# -----------------------------------------------------------------------
p_wt <- ggplot(df_diff, aes(x = strategy, y = wt_diff)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_boxplot(width = 0.4, outlier.shape = NA, fill = "grey") +
  geom_jitter(width = 0.15, size = 2, alpha = 0.7) +
  labs(x = NULL, y = "Weight difference (g)",
       title = "Dying fish wt. minus mean survivor wt. (same day)") +
  theme_bw() + theme(panel.grid = element_blank())

p_ggd <- ggplot(df_diff, aes(x = strategy, y = ggd_diff)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_boxplot(width = 0.4, outlier.shape = NA, fill = "grey") +
  geom_jitter(width = 0.15, size = 2, alpha = 0.7) +
  labs(x = NULL, y = "Growth rate difference (g/g/d)",
       title = "Dying fish growth minus mean survivor growth (same day)") +
  theme_bw() + theme(panel.grid = element_blank())

p_days <- ggplot() +
  geom_rect(data = summer_shading,
            aes(ymin = xmin, ymax = xmax, xmin = ymin, xmax = ymax),
            fill = "grey", alpha = 0.3, inherit.aes = FALSE) +
  geom_jitter(data = death_records,
              aes(x = strategy, y = date, color = strategy), 
              width = 0.15, size = 2.5, alpha = 0.8) +
  scale_x_discrete(limits = strategy_order) +
  labs(x = NULL, y = "Day of death (day of year)",
       title = "When did non-survivors die?") +
  theme_bw() + theme(legend.position = "none", panel.grid = element_blank()) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen"))

egg::ggarrange(p_wt, p_ggd, p_days, ncol = 1)
```
Yes, overall, the survival module is behaving correctly: it's largely selecting against small, slow-growing fish in a biologically interpretable way.

`resident_warm` fish that die in summer are smaller and actively losing mass relative to the population-wide survivor mean on those same days. The tight clustering near zero for `resident_cold` and `optimal_mover` reflects that deaths in those strategies tend to occur early in the year when fish are small and the all-strategy survivor mean hasn't yet diverged much.

Timing of deaths (panel C) also makes biological sense:
  + `resident_cold` deaths cluster early when fish are small and baseline survival is lowest, and fish are also experiencing negative growth due to low temps
  + `optimal_mover` deaths also cluster early for the same reasons
  + `resident_warm` deaths cluster mid-summer as the warm patch becomes suboptimal. Likely not as much early season mortality b/c growth potential is just a bit higher

Now, let's investigate the context of `optimal_mover` mortalities. Where and when to they occur and what are the thermal and growth conditions that they experienced?
```{r fig.width=5, fig.height=8, eval=FALSE}
mover_deaths <- ibm_long %>%
  filter(strategy == "optimal_mover", survived == 0) %>%
  select(pid, date, dayofsim, patch, weight, ggd)

dayofsim_max <- max(mover_deaths$dayofsim) + 10

temps <- habitat_df %>%
  filter(dayofsim <= dayofsim_max) %>%
  select(dayofsim, date, temp_warm, temp_cold) %>%
  pivot_longer(c(temp_warm, temp_cold), names_to = "patch", values_to = "temp") %>%
  mutate(patch = if_else(patch == "temp_warm", "warm", "cold"))

p_temp <- ggplot(temps, aes(x = date, y = temp, color = patch)) +
  geom_line(linewidth = 0.8) +
  geom_rug(
    data = mover_deaths, aes(x = date, color = patch),
    sides = "b", linewidth = 1.2, length = unit(0.06, "npc"), inherit.aes = FALSE
  ) +
  scale_color_manual(values = c(warm = "tomato", cold = "steelblue")) +
  labs(x = NULL, y = "Temperature (°C)", color = "Patch",
       title = "Optimal mover mortality: patch context and timing") +
  theme_bw() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

p_death <- ggplot(mover_deaths, aes(x = date, y = ggd, color = patch, size = weight)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.85) +
  scale_color_manual(values = c(warm = "tomato", cold = "steelblue")) +
  scale_size_continuous(range = c(3, 8), name = "Weight (g)") +
  labs(x = "Date", y = "Growth rate at death (g/g/d)", color = "Patch") +
  theme_bw()

egg::ggarrange(p_temp, p_death, ncol = 1, heights = c(1, 1.2))
```
The early cluster of deaths is dominated by warm-patch fish with negative growth rates and small body sizes — these fish are actively losing mass and dying. Temperatures in both patches are essentially identical and near-zero during this period, so the warm patch offers no thermal advantage, but movers are nonetheless ending up there and dying. The one cold-patch early death follows the same pattern.

The two late deaths (days ~123–133) are different in character: both have positive growth rates and are much larger, suggesting they're stochastic rather than condition-driven mortality.

The key diagnostic question this raises: why are movers preferentially occupying the warm patch in early winter when temperatures are identical and growth is negative in both? That points to investigating the patch-choice logic at low temperatures — the growth potential lookup may not be differentiating between patches when temperatures are near-flat, leading to essentially arbitrary placement. This makes sense, especially when combined with the softmax/probabilistic stochasticity in patch choice...fish won't switch patches if the growth advantage is tiny...so they stay in the warm habitat where densities are ~high and 

