---
title: "Null model: cold only"
---

```{r, include = FALSE, cache = FALSE}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(egg)
library(RColorBrewer)
library(ggridges)
library(patchwork)
library(scales)

source("Functions.R")
source("PlotResults.R")
```

**Purpose:** This page provides the results for the `base` experiment/scenario:

* Thermal regimes are ~multiplicative of each other: maximum thermal heterogeneity during summer, but heterogeneity collapses to ~0 during winter (everything is cold). 
* Habitat area is equal during burn-in; cold habitat is entirely replaced by warm habitat over experiment

```{r}
#| cache: false
res           <- readRDS("results/null_cold.rds")
ibm_long      <- res$ibm_long
habitat_df    <- res$habitat_df
fish_registry <- res$fish_registry
spawn_log     <- res$spawn_log
sim_params    <- res$params

exp_start_date <- sim_params$mindate + years(sim_params$nyears_burnin)
```


## Show habitat 

```{r}
plot_habitat(habitat_df, exp_start_date)
```


## Mechanistic plots

Individual-level bioenergetics, temperature and growth experienced, size trajectories, age structure, survival decomposition, movement, and reproduction.

```{r eval=FALSE}
#| fig-width: 10
#| fig-height: 6
plot_mechanistic(
  ibm_long      = ibm_long,
  habitat_df    = habitat_df,
  sim_params    = sim_params,
  spawn_log     = spawn_log,
  fish_registry = fish_registry,
  exp_start_date = exp_start_date
)
```


## Population plots

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_long,
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


## Habitat value plots

Patch occupancy, total biomass, per-area biomass, growth rate, and daily survival for optimal movers — shown as annual time series, seasonal time series, and equivalence curves.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_patchvalue(
  ibm_long       = ibm_long,
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```
