---
title: "Parameter Sensitivity Analysis"
subtitle: "Effects of key model parameters on population dynamics"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 11
    fig-height: 7
execute:
  warning: false
  message: false
  cache: true
---

```{r include=FALSE}
#| label: setup
#| cache: false

library(tidyverse)
library(lubridate)
library(patchwork)

source("Habitat.R")
source("Functions.R")
source("SimulationLoop.R")
source("ScenarioDefs.R")
```

```{r}
#| label: base parameters
#| cache: false

# ── Baseline parameter set ────────────────────────────────────────────────────
# All sensitivity runs keep every parameter at these values except the focal one.
baseline <- scenarios[["null_cold"]]

exp_start_date <- as.Date(baseline$mindate) + lubridate::years(baseline$nyears_burnin)
burn_start     <- year(as.Date(baseline$mindate))
exp_start_yr   <- year(exp_start_date)
exp_end_yr     <- exp_start_yr + baseline$nyears_experiment

cat("Baseline parameters:\n")
cat("  K_cold / K_warm         :", baseline$K_cold, "/", baseline$K_warm, "\n")
cat("  egg_surv                :", baseline$egg_surv, "\n")
cat("  dominance_beta          :", baseline$dominance_beta, "\n")
cat("  age_structured_competition:", baseline$age_structured_competition, "\n")
cat("  crit_pcmax_lo / hi      :", baseline$crit_pcmax_lo, "/", baseline$crit_pcmax_hi, "\n")
cat("  crit_period_days        :", baseline$crit_period_days, "\n")
cat("Burn-in:", burn_start, "–", exp_start_yr - 1,
    " | Experiment:", exp_start_yr, "–", exp_end_yr, "\n")
```


```{r}
#| label: helpers
#| cache: false

# ── Helper: run one sensitivity simulation ─────────────────────────────────────
run_sens <- function(param_name, param_val, base = baseline) {
  p   <- modifyList(base, setNames(list(param_val), param_name))
  # K must be set in both habitat and params
  if (param_name %in% c("K_cold", "K_warm")) {
    p <- modifyList(p, list(K_cold = param_val, K_warm = param_val))
  }
  hdf <- build_habitat(p)
  run_simulation(habitat_df = hdf, params = p, wt_growth = wt.growth)
}

# ── Helper: Oct 1 annual census (N and total biomass) ─────────────────────────
oct1_census <- function(sim, label) {
  sim$ibm_long |>
    filter(!is.na(date), survived == 1,
           month(date) == 10, day(date) == 1) |>
    mutate(year  = year(date),
           phase = if_else(year < exp_start_yr, "burn-in", "experiment")) |>
    group_by(year, phase) |>
    summarise(N        = n_distinct(pid),
              biomass_kg = sum(weight) / 1000,
              .groups  = "drop") |>
    mutate(label = label)
}

# ── Helper: population map (burn-in N[t+1] ~ N[t]) ───────────────────────────
pop_map <- function(census_df) {
  census_df |>
    filter(phase == "burn-in") |>
    group_by(label) |>
    arrange(year) |>
    mutate(N_next = lead(N)) |>
    ungroup() |>
    filter(!is.na(N_next))
}

# ── Helper: stock-recruitment (60-day survivors vs spawner biomass) ────────────
# Uses fry surviving the critical period (60 days post-hatch) as the recruitment
# metric rather than fry born, because fry born is the *input* to the
# density-dependent critical-period function. 60-day survivors represent the
# post-compensation output and should display a Beverton-Holt plateau against
# spawner biomass if density-dependent thinning is operating correctly.
stock_recruit <- function(sim, label) {
  ibm <- sim$ibm_long
  hdf <- sim$habitat_df

  # Fry surviving at least 60 days post-hatch (past the critical period)
  recruits_60d <- ibm |>
    filter(!is.na(parent_pid)) |>
    mutate(birth_yr = year(hdf$date[birth_dayofsim])) |>
    filter(birth_yr < exp_start_yr) |>
    group_by(pid, birth_yr) |>
    summarise(days_alive = n(), .groups = "drop") |>
    filter(days_alive >= 60) |>
    count(birth_yr, name = "n_surv60")

  # Spawner biomass: adults (≥ 100 g) present during spawning window (Apr–Jun)
  spawner_bio <- ibm |>
    filter(!is.na(date), survived == 1) |>
    mutate(year = year(date), doy = yday(date)) |>
    filter(doy >= 90, doy <= 150, weight >= 100, year < exp_start_yr) |>
    group_by(year) |>
    summarise(spawner_kg = sum(weight) / 1000, .groups = "drop")

  left_join(recruits_60d, spawner_bio, by = c("birth_yr" = "year")) |>
    mutate(label = label, spawner_kg = replace_na(spawner_kg, 0))
}

# ── Helper: critical-period thinning rate vs cohort size ──────────────────────
thinning_rate <- function(sim, label) {
  ibm <- sim$ibm_long
  hdf <- sim$habitat_df

  born <- ibm |>
    filter(!is.na(parent_pid)) |>
    mutate(birth_yr = year(hdf$date[birth_dayofsim])) |>
    filter(birth_yr < exp_start_yr) |>
    distinct(pid, birth_yr) |>
    count(birth_yr, name = "n_born")

  surv60 <- ibm |>
    filter(!is.na(parent_pid)) |>
    mutate(birth_yr = year(hdf$date[birth_dayofsim])) |>
    filter(birth_yr < exp_start_yr) |>
    group_by(pid, birth_yr) |>
    summarise(days_alive = n(), .groups = "drop") |>
    filter(days_alive >= 60) |>
    count(birth_yr, name = "n_surv60")

  left_join(born, surv60, by = "birth_yr") |>
    mutate(surv_rate = replace_na(n_surv60 / n_born, 0),
           label = label)
}

# ── Helper: biomass by age class (Oct 1, burn-in) ────────────────────────────
biomass_age <- function(sim, label) {
  sim$ibm_long |>
    filter(!is.na(date), survived == 1,
           month(date) == 10, day(date) == 1) |>
    mutate(year      = year(date),
           age_class = if_else(cohort == year, "Age-0 fry", "Age-1+"),
           phase     = if_else(year < exp_start_yr, "burn-in", "experiment")) |>
    filter(phase == "burn-in") |>
    group_by(year, age_class) |>
    summarise(biomass_kg = sum(weight) / 1000,
              n_fish     = n_distinct(pid), .groups = "drop") |>
    group_by(age_class) |>
    summarise(mean_bio_kg = mean(biomass_kg),
              mean_n      = mean(n_fish),
              mean_wt_g   = mean(biomass_kg * 1000 / pmax(n_fish, 1)),
              .groups = "drop") |>
    mutate(label = label)
}

# ── Helper: weight-at-age (Oct 1, burn-in mean, ages 0-4) ────────────────────
wt_at_age <- function(sim, label) {
  sim$ibm_long |>
    filter(!is.na(date), survived == 1,
           month(date) == 10, day(date) == 1) |>
    mutate(year     = year(date),
           age_yrs  = floor(age),
           phase    = if_else(year < exp_start_yr, "burn-in", "experiment")) |>
    filter(phase == "burn-in", age_yrs <= 4) |>
    group_by(year, age_yrs) |>
    summarise(mean_wt = mean(weight), n = n_distinct(pid), .groups = "drop") |>
    filter(n >= 2) |>
    group_by(age_yrs) |>
    summarise(mean_wt = mean(mean_wt), .groups = "drop") |>
    mutate(label = label)
}

# ── Helper: age-0 daily weight (doy mean across burn-in cohorts) ──────────────
age0_daily_wt <- function(sim, label) {
  sim$ibm_long |>
    filter(!is.na(date), survived == 1, !is.na(ggd)) |>
    mutate(year = year(date),
           doy  = yday(date)) |>
    filter(cohort == year, year < exp_start_yr) |>
    group_by(doy) |>
    summarise(mean_wt = mean(weight), n = n(), .groups = "drop") |>
    filter(n >= 5) |>
    mutate(label = label)
}

# ── Stability summary ──────────────────────────────────────────────────────────
stability_stats <- function(census_df) {
  census_df |>
    group_by(label, phase) |>
    summarise(mean_N     = round(mean(N)),
              cv         = round(sd(N) / mean(N), 2),
              mean_bio   = round(mean(biomass_kg), 1),
              sim_yrs    = n(),
              .groups    = "drop")
}

# ── Total years simulated (burn-in + experiment combined) ─────────────────────
# Used in the cross-parameter summary to identify premature collapse.
# max possible = nyears_burnin + nyears_experiment = 100.
total_sim_yrs <- function(census_df) {
  census_df |>
    group_by(label) |>
    summarise(total_yrs = n_distinct(year), .groups = "drop")
}

# ── Helper: 4-panel stability bar chart (mean N and CV × phase) ───────────────
# stable_df: output of stability_stats(); cols: length-3 colour vector.
# Layout: rows = metric (mean N, CV), columns = phase (burn-in, experiment).
plot_stability_fig <- function(stable_df, cols) {
  stable_long <- stable_df |>
    select(label, phase, mean_N, cv) |>
    pivot_longer(c(mean_N, cv),
                 names_to  = "metric",
                 values_to = "value") |>
    mutate(
      metric = factor(metric,
                      levels = c("mean_N", "cv"),
                      labels = c("Mean N (Oct 1)", "CV (abundance)")),
      phase  = factor(phase,
                      levels = c("burn-in", "experiment"),
                      labels = c("Burn-in", "Experiment"))
    )

  ggplot(stable_long, aes(x = label, y = value, fill = label)) +
    geom_col(width = 0.65, show.legend = FALSE) +
    facet_grid(metric ~ phase, scales = "free_y") +
    scale_fill_manual(values = cols) +
    labs(x = NULL, y = NULL) +
    theme_sens() +
    theme(axis.text.x  = element_text(size = 7.5, lineheight = 0.85),
          strip.text   = element_text(size = 9),
          panel.spacing = unit(0.8, "lines"))
}

# ── Shared colour palette (Okabe-Ito: blue / teal / vermillion) ──────────────
# Colorblind-friendly; all three are clearly distinguishable.
# Mapping: low value = blue, mid/baseline value = teal, high value = vermillion.
sens_cols <- c("#0072B2", "#009E73", "#D55E00")

# ── Shared plot theme ─────────────────────────────────────────────────────────
theme_sens <- function() {
  theme_bw() +
    theme(legend.position  = "top",
          strip.background = element_rect(fill = "grey88"),
          panel.grid.minor = element_blank(),
          plot.title       = element_text(size = 10, face = "bold"),
          plot.subtitle    = element_text(size = 8, color = "grey30"))
}
```

---

## `K_cold`: strength of density dependence

The half-saturation constant K controls the density at which consumption suppression
becomes binding. Smaller K creates stronger suppression at lower densities; larger K
allows the population to grow more before density effects manifest. All other parameters
held at baseline (`egg_surv = 0.1`, `dominance_beta = 1`).

```{r}
#| label: sim-K
#| cache: true

K_vals   <- c(100, 200, 500)
K_labels <- paste0("K = ", K_vals)
K_cols   <- sens_cols

cat("Running K sensitivity simulations...\n")
K_sims <- lapply(seq_along(K_vals), function(i) {
  cat(" K =", K_vals[i], "... ")
  p   <- modifyList(baseline, list(K_cold = K_vals[i], K_warm = K_vals[i]))
  hdf <- build_habitat(p)
  res <- run_simulation(hdf, p, wt_growth = wt.growth)
  cat("ends", format(max(res$ibm_long$date, na.rm = TRUE)), "\n")
  res
})
names(K_sims) <- K_labels
```

```{r}
#| label: K-outcomes

K_census   <- bind_rows(Map(oct1_census,   K_sims, K_labels)) |>
  mutate(label = factor(label, levels = K_labels))
K_sr       <- bind_rows(Map(stock_recruit, K_sims, K_labels)) |>
  mutate(label = factor(label, levels = K_labels))
K_thin     <- bind_rows(Map(thinning_rate, K_sims, K_labels)) |>
  mutate(label = factor(label, levels = K_labels))

K_waa      <- bind_rows(Map(wt_at_age,     K_sims, K_labels)) |>
  mutate(label = factor(label, levels = K_labels))
K_age0_dly <- bind_rows(Map(age0_daily_wt, K_sims, K_labels)) |>
  mutate(label = factor(label, levels = K_labels))
K_stable   <- stability_stats(K_census)
```

### Abundance and biomass

```{r}
#| label: fig-K-ts
#| fig-cap: "Oct 1 annual abundance and total biomass for each K value. Grey band = burn-in."
#| fig-height: 5

p_N <- ggplot(K_census, aes(x = year, y = N, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = K_cols) +
  labs(x = NULL, y = "N (Oct 1)", color = NULL, title = "Annual abundance") +
  theme_sens() + theme(legend.position = "none")

p_B <- ggplot(K_census, aes(x = year, y = biomass_kg, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = K_cols) +
  labs(x = NULL, y = "Total biomass (kg)", color = NULL, title = "Total biomass") +
  theme_sens()

p_N | p_B
```

### Compensatory dynamics

```{r}
#| label: fig-K-comp
#| fig-cap: "Left: burn-in population map. Center: stock-recruitment relationship. Right: critical-period survival rate vs cohort size."
#| fig-height: 5

K_map <- pop_map(K_census)

p_map <- ggplot(K_map, aes(x = N, y = N_next, color = label)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.75, linewidth = 0.9) +
  scale_color_manual(values = K_cols) +
  labs(x = "N (year t, Oct 1)", y = "N (year t+1, Oct 1)", color = NULL,
       title = "Population map (burn-in)",
       subtitle = "Curve crossing below dashed line = compensatory") +
  theme_sens() + theme(legend.position = "none")

p_sr <- ggplot(K_sr |> filter(spawner_kg > 0),
               aes(x = spawner_kg, y = n_surv60, color = label)) +
  geom_point(alpha = 0.5, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_color_manual(values = K_cols) +
  labs(x = "Spawner biomass (kg, Apr–Jun)", y = "Fry surviving 60 days",
       color = NULL, title = "Stock-recruitment",
       subtitle = "60-day survivors vs. spawner biomass \n— should show B-H plateau") +
  theme_sens()

p_thin <- ggplot(K_thin, aes(x = n_born, y = surv_rate, color = label)) +
  geom_point(alpha = 0.45, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = K_cols) +
  labs(x = "Fry born", y = "% surviving 60-day critical period",
       color = NULL, title = "Critical-period thinning",
       subtitle = "Declining rate = density-dependent compensation") +
  theme_sens() + theme(legend.position = "none")

p_map | p_sr | p_thin
```

### Age structure

```{r}
#| label: fig-K-thin
#| fig-cap: "Left: mean weight at age (burn-in). Right: Mean age-0 fry weight by day of year, averaged across burn-in cohorts."
#| fig-height: 5

p_waa <- ggplot(K_waa, aes(x = age_yrs, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_color_manual(values = K_cols) +
  scale_x_continuous(breaks = 0:4) +
  labs(x = "Age (years)", y = "Mean weight (g, Oct 1)",
       color = NULL, title = "Weight at age (burn-in mean)",
       subtitle = "Shows how K affects individual growth trajectories") +
  theme_sens()

month_doys   <- c(121, 152, 182, 213, 244, 274, 305, 335)
month_labels <- c("May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")

p_aget <- ggplot(K_age0_dly |> filter(doy >= 118),
       aes(x = doy, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = K_cols) +
  scale_x_continuous(breaks = month_doys, labels = month_labels) +
  labs(x = NULL, y = "Mean weight (g)", color = NULL,
       title = "Age-0 daily weight — burn-in cohort mean",
       subtitle = "Smaller K = more density suppression = slower early growth") +
  theme_sens()

p_waa | p_aget
```


### Stability summary {.unnumbered}

```{r}
#| label: fig-K-stable
#| fig-cap: "Mean Oct 1 abundance and CV by K value, split by simulation phase."
#| fig-height: 5

plot_stability_fig(K_stable, K_cols)
```

```{r}
#| label: tbl-K-stable
#| tbl-cap: "Burn-in and experiment stability statistics by K value."

K_stable |>
  knitr::kable(col.names = c("K", "Phase", "Mean N",
                              "CV (abundance)", "Mean biomass (kg)", "Years"))
```

---

## `egg_surv`: proportional egg survival

Egg-to-fry survival scales the number of recruits entering the simulation each spring.
It represents spawning habitat quality, redd conditions, and egg-incubation success.
All other parameters held at baseline (`K = 200`, `dominance_beta = 1`).

```{r}
#| label: sim-egg
#| cache: true

egg_vals   <- c(0.01, 0.1, 0.2)
egg_labels <- paste0("egg_surv = ", egg_vals)
egg_cols   <- sens_cols

cat("Running egg_surv sensitivity simulations...\n")
egg_sims <- lapply(seq_along(egg_vals), function(i) {
  cat(" egg_surv =", egg_vals[i], "... ")
  p   <- modifyList(baseline, list(egg_surv = egg_vals[i]))
  hdf <- build_habitat(p)
  res <- run_simulation(hdf, p, wt_growth = wt.growth)
  cat("ends", format(max(res$ibm_long$date, na.rm = TRUE)), "\n")
  res
})
names(egg_sims) <- egg_labels
```

```{r}
#| label: egg-outcomes

egg_census   <- bind_rows(Map(oct1_census,   egg_sims, egg_labels)) |>
  mutate(label = factor(label, levels = egg_labels))
egg_sr       <- bind_rows(Map(stock_recruit, egg_sims, egg_labels)) |>
  mutate(label = factor(label, levels = egg_labels))
egg_thin     <- bind_rows(Map(thinning_rate, egg_sims, egg_labels)) |>
  mutate(label = factor(label, levels = egg_labels))

egg_waa      <- bind_rows(Map(wt_at_age,     egg_sims, egg_labels)) |>
  mutate(label = factor(label, levels = egg_labels))
egg_age0_dly <- bind_rows(Map(age0_daily_wt, egg_sims, egg_labels)) |>
  mutate(label = factor(label, levels = egg_labels))
egg_stable   <- stability_stats(egg_census)
```

### Abundance and biomass

```{r}
#| label: fig-egg-ts
#| fig-cap: "Oct 1 annual abundance and total biomass for each egg_surv value."
#| fig-height: 5

p_N <- ggplot(egg_census, aes(x = year, y = N, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = egg_cols) +
  labs(x = NULL, y = "N (Oct 1)", color = NULL, title = "Annual abundance") +
  theme_sens() + theme(legend.position = "none")

p_B <- ggplot(egg_census, aes(x = year, y = biomass_kg, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = egg_cols) +
  labs(x = NULL, y = "Total biomass (kg)", color = NULL, title = "Total biomass") +
  theme_sens()

p_N | p_B
```

### Compensatory dynamics

```{r}
#| label: fig-egg-comp
#| fig-cap: "Left: burn-in population map. Center: stock-recruitment. Right: critical-period survival rate vs cohort size."
#| fig-height: 5

egg_map <- pop_map(egg_census)

p_map <- ggplot(egg_map, aes(x = N, y = N_next, color = label)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.75, linewidth = 0.9) +
  scale_color_manual(values = egg_cols) +
  labs(x = "N (year t, Oct 1)", y = "N (year t+1, Oct 1)", color = NULL,
       title = "Population map (burn-in)") +
  theme_sens() + theme(legend.position = "none")

p_sr <- ggplot(egg_sr |> filter(spawner_kg > 0),
               aes(x = spawner_kg, y = n_surv60, color = label)) +
  geom_point(alpha = 0.5, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_color_manual(values = egg_cols) +
  labs(x = "Spawner biomass (kg)", y = "Fry surviving 60 days",
       color = NULL, title = "Stock-recruitment") +
  theme_sens()

p_thin <- ggplot(egg_thin, aes(x = n_born, y = surv_rate, color = label)) +
  geom_point(alpha = 0.45, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = egg_cols) +
  labs(x = "Fry born", y = "% surviving 60-day critical period",
       color = NULL, title = "Critical-period thinning") +
  theme_sens() + theme(legend.position = "none")

p_map | p_sr | p_thin
```

### Age structure

```{r}
#| label: fig-egg-age0
#| fig-cap: "Left: mean weight at age (burn-in). Right: age-0 daily weight by day of year, averaged across burn-in cohorts."
#| fig-height: 5

p_waa <- ggplot(egg_waa, aes(x = age_yrs, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_color_manual(values = egg_cols) +
  scale_x_continuous(breaks = 0:4) +
  labs(x = "Age (years)", y = "Mean weight (g, Oct 1)",
       color = NULL, title = "Weight at age (burn-in mean)",
       subtitle = "Higher egg_surv: smaller age-0, larger age-2+ (size-selective thinning)") +
  theme_sens()

p_aget <- ggplot(egg_age0_dly |> filter(doy >= 118),
                 aes(x = doy, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = egg_cols) +
  scale_x_continuous(breaks = month_doys, labels = month_labels) +
  labs(x = NULL, y = "Mean weight (g)", color = NULL,
       title = "Age-0 daily weight — burn-in cohort mean") +
  theme_sens()

p_waa | p_aget
```

### Stability summary {.unnumbered}

```{r}
#| label: fig-egg-stable
#| fig-cap: "Mean Oct 1 abundance and CV by egg_surv, split by simulation phase."
#| fig-height: 5

plot_stability_fig(egg_stable, egg_cols)
```

```{r}
#| label: tbl-egg-stable
#| tbl-cap: "Stability statistics by egg_surv."

egg_stable |>
  knitr::kable(col.names = c("egg_surv", "Phase", "Mean N",
                              "CV (abundance)", "Mean biomass (kg)", "Years"))
```

---

## `dominance_beta`: size-based competitive dominance

`dominance_beta` controls the strength of size-based competitive dominance *within*
each age class. β = 0 is pure scramble (all fish experience the same effective density
regardless of size); β = 1 means a fish twice the size of its competitor imposes a full
competitive unit of pressure, while a fish half its size imposes only half a unit.
All other parameters held at baseline (`K = 200`, `egg_surv = 0.1`).

```{r}
#| label: sim-beta
#| cache: true

beta_vals   <- c(0, 0.5, 1)
beta_labels <- paste0("β = ", beta_vals)
beta_cols   <- sens_cols

cat("Running dominance_beta sensitivity simulations...\n")
beta_sims <- lapply(seq_along(beta_vals), function(i) {
  cat(" beta =", beta_vals[i], "... ")
  p   <- modifyList(baseline, list(dominance_beta = beta_vals[i]))
  hdf <- build_habitat(p)
  res <- run_simulation(hdf, p, wt_growth = wt.growth)
  cat("ends", format(max(res$ibm_long$date, na.rm = TRUE)), "\n")
  res
})
names(beta_sims) <- beta_labels
```

```{r}
#| label: beta-outcomes

beta_census   <- bind_rows(Map(oct1_census,   beta_sims, beta_labels)) |>
  mutate(label = factor(label, levels = beta_labels))
beta_sr       <- bind_rows(Map(stock_recruit, beta_sims, beta_labels)) |>
  mutate(label = factor(label, levels = beta_labels))
beta_thin     <- bind_rows(Map(thinning_rate, beta_sims, beta_labels)) |>
  mutate(label = factor(label, levels = beta_labels))
beta_waa      <- bind_rows(Map(wt_at_age,     beta_sims, beta_labels)) |>
  mutate(label = factor(label, levels = beta_labels))
beta_age0_dly <- bind_rows(Map(age0_daily_wt, beta_sims, beta_labels)) |>
  mutate(label = factor(label, levels = beta_labels))
beta_stable   <- stability_stats(beta_census)
```

### Abundance and biomass

```{r}
#| label: fig-beta-ts
#| fig-cap: "Oct 1 annual abundance and total biomass for each dominance_beta value."
#| fig-height: 5

p_N <- ggplot(beta_census, aes(x = year, y = N, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = beta_cols) +
  labs(x = NULL, y = "N (Oct 1)", color = NULL, title = "Annual abundance") +
  theme_sens() + theme(legend.position = "none")

p_B <- ggplot(beta_census, aes(x = year, y = biomass_kg, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = beta_cols) +
  labs(x = NULL, y = "Total biomass (kg)", color = NULL, title = "Total biomass") +
  theme_sens()

p_N | p_B
```

### Compensatory dynamics

```{r}
#| label: fig-beta-comp
#| fig-cap: "Left: burn-in population map. Center: stock-recruitment. Right: critical-period survival rate vs cohort size."
#| fig-height: 5

beta_map <- pop_map(beta_census)

p_map <- ggplot(beta_map, aes(x = N, y = N_next, color = label)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.75, linewidth = 0.9) +
  scale_color_manual(values = beta_cols) +
  labs(x = "N (year t, Oct 1)", y = "N (year t+1, Oct 1)", color = NULL,
       title = "Population map (burn-in)") +
  theme_sens() + theme(legend.position = "none")

p_sr <- ggplot(beta_sr |> filter(spawner_kg > 0),
               aes(x = spawner_kg, y = n_surv60, color = label)) +
  geom_point(alpha = 0.5, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_color_manual(values = beta_cols) +
  labs(x = "Spawner biomass (kg)", y = "Fry surviving 60 days",
       color = NULL, title = "Stock-recruitment") +
  theme_sens()

p_thin <- ggplot(beta_thin, aes(x = n_born, y = surv_rate, color = label)) +
  geom_point(alpha = 0.45, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = beta_cols) +
  labs(x = "Fry born", y = "% surviving 60-day critical period",
       color = NULL, title = "Critical-period thinning") +
  theme_sens() + theme(legend.position = "none")

p_map | p_sr | p_thin
```

### Age structure

```{r}
#| label: fig-beta-age0
#| fig-cap: "Left: mean weight at age (burn-in). Right: age-0 daily weight by day of year, averaged across burn-in cohorts."
#| fig-height: 5

p_waa <- ggplot(beta_waa, aes(x = age_yrs, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_color_manual(values = beta_cols) +
  scale_x_continuous(breaks = 0:4) +
  labs(x = "Age (years)", y = "Mean weight (g, Oct 1)",
       color = NULL, title = "Weight at age (burn-in mean)",
       subtitle = "Higher β: larger dominant adults, more skewed size distribution") +
  theme_sens()

p_aget <- ggplot(beta_age0_dly |> filter(doy >= 118),
                 aes(x = doy, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = beta_cols) +
  scale_x_continuous(breaks = month_doys, labels = month_labels) +
  labs(x = NULL, y = "Mean weight (g)", color = NULL,
       title = "Age-0 daily weight — burn-in cohort mean") +
  theme_sens()

p_waa | p_aget
```

### Stability summary {.unnumbered}

```{r}
#| label: fig-beta-stable
#| fig-cap: "Mean Oct 1 abundance and CV by dominance_beta, split by simulation phase."
#| fig-height: 5

plot_stability_fig(beta_stable, beta_cols)
```

```{r}
#| label: tbl-beta-stable
#| tbl-cap: "Stability statistics by dominance_beta."

beta_stable |>
  knitr::kable(col.names = c("β", "Phase", "Mean N",
                              "CV (abundance)", "Mean biomass (kg)", "Years"))
```

---

## `crit_pcmax_lo` / `crit_pcmax_hi`: critical-period threshold 

`fncSurviveConsumption` is parameterised by two thresholds that define the lower and
upper ends of the logistic sigmoid. At `crit_pcmax_lo`, the 60-day period survival
approaches 1%; at `crit_pcmax_hi`, it approaches 99%. These values are phenomenological
— they were calibrated to the K = 200 consumption distributions observed during the
diagnostic session, not derived from first principles. This section tests how sensitive
the compensatory regulation is to shifting the sigmoid window up or down.

| Scenario | `crit_pcmax_lo` | `crit_pcmax_hi` | Midpoint | Interpretation |
|---|---|---|---|---|
| Low window | 0.25 | 0.35 | 0.30 | Lenient — mortality only kicks in below pcmax_dd ≈ 0.35; most boom-year fry (at ~0.38–0.40) escape |
| **Baseline** | **0.30** | **0.40** | **0.35** | **Current default** |
| High window | 0.35 | 0.45 | 0.40 | Strict — boom-year fry at ~0.38–0.40 fall within the penalty zone; stronger thinning |

All other parameters held at baseline (`K = 200`, `egg_surv = 0.1`, `dominance_beta = 1`).

```{r}
#| label: sim-crit-thresh
#| cache: true

thresh_vals <- list(
  c(lo = 0.25, hi = 0.35),
  c(lo = 0.30, hi = 0.40),
  c(lo = 0.35, hi = 0.45)
)
thresh_labels <- c("lo=0.25 / hi=0.35\n(lenient)", 
                   "lo=0.30 / hi=0.40\n(baseline)",
                   "lo=0.35 / hi=0.45\n(strict)")
thresh_cols <- sens_cols

cat("Running crit_pcmax threshold sensitivity simulations...\n")
thresh_sims <- lapply(seq_along(thresh_vals), function(i) {
  tv <- thresh_vals[[i]]
  cat(" lo=", tv["lo"], "hi=", tv["hi"], "... ")
  p   <- modifyList(baseline, list(crit_pcmax_lo = tv["lo"],
                                   crit_pcmax_hi = tv["hi"]))
  hdf <- build_habitat(p)
  res <- run_simulation(hdf, p, wt_growth = wt.growth)
  cat("ends", format(max(res$ibm_long$date, na.rm = TRUE)), "\n")
  res
})
names(thresh_sims) <- thresh_labels
```

```{r}
#| label: thresh-outcomes

thresh_census   <- bind_rows(Map(oct1_census,   thresh_sims, thresh_labels)) |>
  mutate(label = factor(label, levels = thresh_labels))
thresh_sr       <- bind_rows(Map(stock_recruit, thresh_sims, thresh_labels)) |>
  mutate(label = factor(label, levels = thresh_labels))
thresh_thin     <- bind_rows(Map(thinning_rate, thresh_sims, thresh_labels)) |>
  mutate(label = factor(label, levels = thresh_labels))
thresh_waa      <- bind_rows(Map(wt_at_age,     thresh_sims, thresh_labels)) |>
  mutate(label = factor(label, levels = thresh_labels))
thresh_age0_dly <- bind_rows(Map(age0_daily_wt, thresh_sims, thresh_labels)) |>
  mutate(label = factor(label, levels = thresh_labels))
thresh_stable   <- stability_stats(thresh_census)
```

### Abundance and biomass

```{r}
#| label: fig-thresh-ts
#| fig-cap: "Oct 1 abundance and biomass by threshold window. Stricter thresholds create stronger early mortality and smaller but more stable populations."
#| fig-height: 5

p_N <- ggplot(thresh_census, aes(x = year, y = N, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = thresh_cols) +
  labs(x = NULL, y = "N (Oct 1)", color = NULL, title = "Annual abundance") +
  theme_sens() + theme(legend.position = "none")

p_B <- ggplot(thresh_census, aes(x = year, y = biomass_kg, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = thresh_cols) +
  labs(x = NULL, y = "Total biomass (kg)", color = NULL, title = "Total biomass") +
  theme_sens()

p_N | p_B
```

### Compensatory dynamics

```{r}
#| label: fig-thresh-comp
#| fig-cap: "Left: burn-in population map. Center: stock-recruitment. Right: critical-period survival rate vs cohort size."
#| fig-height: 5

thresh_map <- pop_map(thresh_census)

p_map <- ggplot(thresh_map, aes(x = N, y = N_next, color = label)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.75, linewidth = 0.9) +
  scale_color_manual(values = thresh_cols) +
  labs(x = "N (year t, Oct 1)", y = "N (year t+1, Oct 1)", color = NULL,
       title = "Population map (burn-in)",
       subtitle = "Does compensatory regulation persist across threshold values?") +
  theme_sens() + theme(legend.position = "none")

p_sr <- ggplot(thresh_sr |> filter(spawner_kg > 0),
               aes(x = spawner_kg, y = n_surv60, color = label)) +
  geom_point(alpha = 0.5, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_color_manual(values = thresh_cols) +
  labs(x = "Spawner biomass (kg)", y = "Fry surviving 60 days", color = NULL,
       title = "Stock-recruitment") +
  theme_sens()

p_thin <- ggplot(thresh_thin, aes(x = n_born, y = surv_rate, color = label)) +
  geom_point(alpha = 0.45, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = thresh_cols) +
  labs(x = "Fry born", y = "% surviving 60-day critical period",
       color = NULL, title = "Critical-period thinning",
       subtitle = "Stricter window (high lo/hi) = lower survival at any given cohort size") +
  theme_sens() + theme(legend.position = "none")

p_map | p_sr | p_thin
```

### Age structure

```{r}
#| label: fig-thresh-age0
#| fig-cap: "Left: mean weight at age (burn-in). Right: age-0 daily weight by day of year, averaged across burn-in cohorts."
#| fig-height: 5

p_waa <- ggplot(thresh_waa, aes(x = age_yrs, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_color_manual(values = thresh_cols) +
  scale_x_continuous(breaks = 0:4) +
  labs(x = "Age (years)", y = "Mean weight (g, Oct 1)",
       color = NULL, title = "Weight at age (burn-in mean)",
       subtitle = "Stricter window (high lo/hi) = stronger size-selection = larger older fish") +
  theme_sens()

p_aget <- ggplot(thresh_age0_dly |> filter(doy >= 118),
                 aes(x = doy, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = thresh_cols) +
  scale_x_continuous(breaks = month_doys, labels = month_labels) +
  labs(x = NULL, y = "Mean weight (g)", color = NULL,
       title = "Age-0 daily weight — burn-in cohort mean") +
  theme_sens()

p_waa | p_aget
```

### Stability summary {.unnumbered}

```{r}
#| label: fig-thresh-stable
#| fig-cap: "Mean Oct 1 abundance and CV by threshold window, split by simulation phase."
#| fig-height: 5

plot_stability_fig(thresh_stable, thresh_cols)
```

```{r}
#| label: tbl-thresh-stable
#| tbl-cap: "Stability statistics by threshold window."

thresh_stable |>
  knitr::kable(col.names = c("Threshold window", "Phase", "Mean N",
                              "CV (abundance)", "Mean biomass (kg)", "Years"))
```

---

## `crit_period_days`: critical-period duration

The Elliott (1989) critical period is approximately 60 days for *Salmo trutta*, but
the precise duration for the focal species is uncertain. A longer critical period
extends the window during which density-dependent thinning applies; a shorter window
concentrates mortality in the first month post-hatch. This section tests whether
the compensatory regulation is robust to the duration assumption.

All other parameters held at baseline (`K = 200`, `egg_surv = 0.1`, `dominance_beta = 1`,
`crit_pcmax_lo = 0.30`, `crit_pcmax_hi = 0.40`).

```{r}
#| label: sim-crit-days
#| cache: true

days_vals   <- c(0L, 30L, 60L, 90L)
days_labels <- c("0 days\n(disabled)", "30 days\n(narrow)",
                 "60 days\n(baseline)", "90 days\n(wide)")
# Grey for disabled (0 days), then three sequential colours
days_cols   <- c("#999999", "#0072B2", "#009E73", "#D55E00")

cat("Running crit_period_days sensitivity simulations...\n")
days_sims <- lapply(seq_along(days_vals), function(i) {
  cat(" crit_period_days =", days_vals[i], "... ")
  p   <- modifyList(baseline, list(crit_period_days = days_vals[i]))
  hdf <- build_habitat(p)
  res <- run_simulation(hdf, p, wt_growth = wt.growth)
  cat("ends", format(max(res$ibm_long$date, na.rm = TRUE)), "\n")
  res
})
names(days_sims) <- days_labels

# thinning_rate uses 60-day survival by default; create a version parameterised
# per scenario so the survival window matches the critical period being tested
thinning_rate_n <- function(sim, label, n_days) {
  ibm <- sim$ibm_long; hdf <- sim$habitat_df
  born <- ibm |>
    filter(!is.na(parent_pid)) |>
    mutate(birth_yr = year(hdf$date[birth_dayofsim])) |>
    filter(birth_yr < exp_start_yr) |>
    distinct(pid, birth_yr) |>
    count(birth_yr, name = "n_born")
  surv_n <- ibm |>
    filter(!is.na(parent_pid)) |>
    mutate(birth_yr = year(hdf$date[birth_dayofsim])) |>
    filter(birth_yr < exp_start_yr) |>
    group_by(pid, birth_yr) |>
    summarise(days_alive = n(), .groups = "drop") |>
    filter(days_alive >= n_days) |>
    count(birth_yr, name = "n_surv")
  left_join(born, surv_n, by = "birth_yr") |>
    mutate(surv_rate = replace_na(n_surv / n_born, 0), label = label)
}
```

```{r}
#| label: days-outcomes

days_census   <- bind_rows(Map(oct1_census,   days_sims, days_labels)) |>
  mutate(label = factor(label, levels = days_labels))
days_sr       <- bind_rows(Map(stock_recruit, days_sims, days_labels)) |>
  mutate(label = factor(label, levels = days_labels))
# Thinning rate: 0-day (disabled) scenario uses standard 60-day window to show
# natural cohort thinning without the function; others use their own period length.
days_thin <- bind_rows(
  thinning_rate(   days_sims[[1]], days_labels[1]),           # 0 days: no function
  thinning_rate_n( days_sims[[2]], days_labels[2], days_vals[2]),
  thinning_rate_n( days_sims[[3]], days_labels[3], days_vals[3]),
  thinning_rate_n( days_sims[[4]], days_labels[4], days_vals[4])
) |> mutate(label = factor(label, levels = days_labels))
days_waa      <- bind_rows(Map(wt_at_age,     days_sims, days_labels)) |>
  mutate(label = factor(label, levels = days_labels))
days_age0_dly <- bind_rows(Map(age0_daily_wt, days_sims, days_labels)) |>
  mutate(label = factor(label, levels = days_labels))
days_stable   <- stability_stats(days_census)
```

### Abundance and biomass

```{r}
#| label: fig-days-ts
#| fig-cap: "Oct 1 abundance and biomass by critical-period duration."
#| fig-height: 5

p_N <- ggplot(days_census, aes(x = year, y = N, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = days_cols) +
  labs(x = NULL, y = "N (Oct 1)", color = NULL, title = "Annual abundance") +
  theme_sens() + theme(legend.position = "none")

p_B <- ggplot(days_census, aes(x = year, y = biomass_kg, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = days_cols) +
  labs(x = NULL, y = "Total biomass (kg)", color = NULL, title = "Total biomass") +
  theme_sens()

p_N | p_B
```

### Compensatory dynamics

```{r}
#| label: fig-days-comp
#| fig-cap: "Left: burn-in population map. Center: stock-recruitment. Right: critical-period survival rate vs cohort size (each line uses its own period length as threshold)."
#| fig-height: 5

days_map <- pop_map(days_census)

p_map <- ggplot(days_map, aes(x = N, y = N_next, color = label)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.75, linewidth = 0.9) +
  scale_color_manual(values = days_cols) +
  labs(x = "N (year t, Oct 1)", y = "N (year t+1, Oct 1)", color = NULL,
       title = "Population map (burn-in)",
       subtitle = "Does the compensation attractor persist across period lengths?") +
  theme_sens() + theme(legend.position = "none")

p_sr <- ggplot(days_sr |> filter(spawner_kg > 0),
               aes(x = spawner_kg, y = n_surv60, color = label)) +
  geom_point(alpha = 0.5, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_color_manual(values = days_cols) +
  labs(x = "Spawner biomass (kg)", y = "Fry surviving 60 days", color = NULL,
       title = "Stock-recruitment") +
  theme_sens()

p_thin <- ggplot(days_thin, aes(x = n_born, y = surv_rate, color = label)) +
  geom_point(alpha = 0.45, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = days_cols) +
  labs(x = "Fry born", y = "% surviving critical period",
       color = NULL, title = "Critical-period thinning") +
  theme_sens() + theme(legend.position = "none")

p_map | p_sr | p_thin
```

### Age structure

```{r}
#| label: fig-days-age0
#| fig-cap: "Left: mean weight at age (burn-in). Right: age-0 daily weight by day of year, averaged across burn-in cohorts."
#| fig-height: 5

p_waa <- ggplot(days_waa, aes(x = age_yrs, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_color_manual(values = days_cols) +
  scale_x_continuous(breaks = 0:4) +
  labs(x = "Age (years)", y = "Mean weight (g, Oct 1)",
       color = NULL, title = "Weight at age (burn-in mean)",
       subtitle = "Longer critical period = more size-selection = larger adults") +
  theme_sens()

p_aget <- ggplot(days_age0_dly |> filter(doy >= 118),
                 aes(x = doy, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = days_cols) +
  scale_x_continuous(breaks = month_doys, labels = month_labels) +
  labs(x = NULL, y = "Mean weight (g)", color = NULL,
       title = "Age-0 daily weight — burn-in cohort mean") +
  theme_sens()

p_waa | p_aget
```

### Stability summary {.unnumbered}

```{r}
#| label: fig-days-stable
#| fig-cap: "Mean Oct 1 abundance and CV by critical-period duration, split by simulation phase."
#| fig-height: 5

plot_stability_fig(days_stable, days_cols)
```

```{r}
#| label: tbl-days-stable
#| tbl-cap: "Stability statistics by critical-period duration."

days_stable |>
  knitr::kable(col.names = c("Duration", "Phase", "Mean N",
                              "CV (abundance)", "Mean biomass (kg)", "Years"))
```

---

## `pcmax_cold` / `pcmax_warm`: maximum consumption

`pcmax` sets the ceiling on achievable daily consumption as a proportion of
physiological maximum Cmax. The effective ration is
`pcmax_adjusted = min(fT(temp), pcmax)`, so at peak summer cold-patch temperatures
(T ≈ 15°C, fT ≈ 0.91), `pcmax_cold = 0.5` is the binding constraint — fish are
food-limited by habitat quality, not by temperature. Raising pcmax relaxes this
ceiling and directly increases `pcmax_adjusted_dd` for all fish.

All other parameters held at baseline (`K = 200`, `egg_surv = 0.1`,
`dominance_beta = 1`, `crit_pcmax_lo = 0.30`, `crit_pcmax_hi = 0.40`).

::: {.callout-warning}
## Interaction with critical-period thresholds

The thresholds `crit_pcmax_lo = 0.30` and `crit_pcmax_hi = 0.40` were calibrated
when `pcmax_cold = 0.5`, where boom-year fry sit at pcmax_adjusted_dd ≈ 0.38 —
just below `crit_pcmax_hi`. At `pcmax = 0.75`, the same fry may achieve
pcmax_adjusted_dd ≈ 0.57, which is well above the threshold, so the
critical-period function barely fires and compensatory regulation weakens or
disappears. At `pcmax = 1.0` temperature becomes the binding constraint
(fT ≈ 0.91) and pcmax_adjusted_dd ≈ 0.82. **The thresholds should be
recalibrated proportionally if pcmax is changed from the baseline.**
Plots below will reveal whether compensation survives the shift.
:::

```{r}
#| label: sim-pcmax
#| cache: false

pcmax_vals   <- c(0.50, 0.75, 1.00)
pcmax_labels <- paste0("pcmax = ", pcmax_vals)
pcmax_cols   <- c("#0072B2", "#009E73", "#D55E00")

cache_file <- "cache/pcmax_sims.rds"
if (file.exists(cache_file)) {
  cat("Loading pcmax simulations from cache...\n")
  pcmax_sims <- readRDS(cache_file)
} else {
  cat("Running pcmax sensitivity simulations...\n")
  pcmax_sims <- lapply(seq_along(pcmax_vals), function(i) {
    pv <- pcmax_vals[i]
    cat(" pcmax =", pv, "... ")
    p   <- modifyList(baseline, list(pcmax_cold = pv, pcmax_warm = pv))
    hdf <- build_habitat(p)
    res <- run_simulation(hdf, p, wt_growth = wt.growth)
    cat("ends", format(max(res$ibm_long$date, na.rm = TRUE)), "\n")
    res
  })
  names(pcmax_sims) <- pcmax_labels
  dir.create("cache", showWarnings = FALSE)
  saveRDS(pcmax_sims, cache_file)
}
```

```{r}
#| label: pcmax-outcomes

pcmax_census   <- bind_rows(Map(oct1_census,   pcmax_sims, pcmax_labels)) |>
  mutate(label = factor(label, levels = pcmax_labels))
pcmax_sr       <- bind_rows(Map(stock_recruit, pcmax_sims, pcmax_labels)) |>
  mutate(label = factor(label, levels = pcmax_labels))
pcmax_thin     <- bind_rows(Map(thinning_rate, pcmax_sims, pcmax_labels)) |>
  mutate(label = factor(label, levels = pcmax_labels))
pcmax_waa      <- bind_rows(Map(wt_at_age,     pcmax_sims, pcmax_labels)) |>
  mutate(label = factor(label, levels = pcmax_labels))
pcmax_age0_dly <- bind_rows(Map(age0_daily_wt, pcmax_sims, pcmax_labels)) |>
  mutate(label = factor(label, levels = pcmax_labels))
pcmax_stable   <- stability_stats(pcmax_census)
```

### Abundance and biomass

```{r}
#| label: fig-pcmax-ts
#| fig-cap: "Oct 1 abundance and biomass by pcmax value."
#| fig-height: 5

p_N <- ggplot(pcmax_census, aes(x = year, y = N, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = pcmax_cols) +
  labs(x = NULL, y = "N (Oct 1)", color = NULL, title = "Annual abundance") +
  theme_sens() + theme(legend.position = "none")

p_B <- ggplot(pcmax_census, aes(x = year, y = biomass_kg, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = pcmax_cols) +
  labs(x = NULL, y = "Total biomass (kg)", color = NULL, title = "Total biomass") +
  theme_sens()

p_N | p_B
```

### Compensatory dynamics

```{r}
#| label: fig-pcmax-comp
#| fig-cap: "Left: burn-in population map. Center: stock-recruitment. Right: critical-period survival rate vs cohort size. Weakening compensation at higher pcmax indicates threshold recalibration is needed."
#| fig-height: 5

pcmax_map <- pop_map(pcmax_census)

p_map <- ggplot(pcmax_map, aes(x = N, y = N_next, color = label)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.75, linewidth = 0.9) +
  scale_color_manual(values = pcmax_cols) +
  labs(x = "N (year t, Oct 1)", y = "N (year t+1, Oct 1)", color = NULL,
       title = "Population map (burn-in)") +
  theme_sens() + theme(legend.position = "none")

p_sr <- ggplot(pcmax_sr |> filter(spawner_kg > 0),
               aes(x = spawner_kg, y = n_surv60, color = label)) +
  geom_point(alpha = 0.5, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_color_manual(values = pcmax_cols) +
  labs(x = "Spawner biomass (kg)", y = "Fry surviving 60 days", color = NULL,
       title = "Stock-recruitment") +
  theme_sens()

p_thin <- ggplot(pcmax_thin, aes(x = n_born, y = surv_rate, color = label)) +
  geom_point(alpha = 0.45, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = pcmax_cols) +
  labs(x = "Fry born", y = "% surviving 60-day critical period",
       color = NULL, title = "Critical-period thinning",
       subtitle = "Rising survival rate at higher pcmax = compensation breaking down") +
  theme_sens() + theme(legend.position = "none")

p_map | p_sr | p_thin
```

### Age structure

```{r}
#| label: fig-pcmax-age0
#| fig-cap: "Left: mean weight at age (burn-in). Right: age-0 daily weight by day of year, averaged across burn-in cohorts."
#| fig-height: 5

p_waa <- ggplot(pcmax_waa, aes(x = age_yrs, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_color_manual(values = pcmax_cols) +
  scale_x_continuous(breaks = 0:4) +
  labs(x = "Age (years)", y = "Mean weight (g, Oct 1)",
       color = NULL, title = "Weight at age (burn-in mean)",
       subtitle = "Higher pcmax = better ration = larger fish at all ages") +
  theme_sens()

p_aget <- ggplot(pcmax_age0_dly |> filter(doy >= 118),
                 aes(x = doy, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = pcmax_cols) +
  scale_x_continuous(breaks = month_doys, labels = month_labels) +
  labs(x = NULL, y = "Mean weight (g)", color = NULL,
       title = "Age-0 daily weight — burn-in cohort mean",
       subtitle = "Higher pcmax lifts the growth ceiling; peak weight and winter decline both affected") +
  theme_sens()

p_waa | p_aget
```

### Stability summary {.unnumbered}

```{r}
#| label: fig-pcmax-stable
#| fig-cap: "Stability by pcmax."
#| fig-height: 5

plot_stability_fig(pcmax_stable, pcmax_cols)
```

```{r}
#| label: tbl-pcmax-stable
#| tbl-cap: "Stability statistics by pcmax."

pcmax_stable |>
  knitr::kable(col.names = c("pcmax", "Phase", "Mean N",
                              "CV (abundance)", "Mean biomass (kg)", "Years"))
```

---

## `A_cold`: habitat area

`A_cold` sets the area of cold habitat available to fish (in arbitrary area units;
`eff_density = (n - 1) / A_cold`). For the null cold-only scenario, `A_warm = 0`
throughout, and cold habitat declines linearly from `A_cold` to 0 over the 60-year
experiment period regardless of starting area. Larger starting area means a larger
effective carrying capacity during the burn-in and a longer absolute decline before
the population is forced into critically small habitat.

All other parameters held at baseline (`K = 200`, `egg_surv = 0.1`,
`dominance_beta = 1`, `crit_pcmax_lo = 0.30`, `crit_pcmax_hi = 0.40`).

::: {.callout-warning}
## Interaction with K and critical-period thresholds

Increasing `A_cold` is mathematically equivalent to scaling K upward: both reduce
`eff_density` for a given N, shifting boom-year fry pcmax_adjusted_dd upward. At
`A_cold = 2`, eff_density halves for the same population size, pushing pcmax_dd
toward ~0.45–0.48 and above `crit_pcmax_hi = 0.40`. At `A_cold = 4`, the
critical-period function likely fires negligibly. This is the same interaction as
for K: **K and A_cold are partially substitutable**, and the critical-period
thresholds may need recalibration if habitat area deviates substantially from the
baseline. The population map and SR curve will reveal whether compensation degrades
under larger areas.
:::

```{r}
#| label: sim-area
#| cache: false

area_vals   <- c(1, 2, 4)
area_labels <- paste0("A_cold = ", area_vals)
area_cols   <- c("#0072B2", "#009E73", "#D55E00")

cache_file <- "cache/area_sims.rds"
if (file.exists(cache_file)) {
  cat("Loading habitat area simulations from cache...\n")
  area_sims <- readRDS(cache_file)
} else {
  cat("Running habitat area sensitivity simulations...\n")
  area_sims <- lapply(seq_along(area_vals), function(i) {
    av <- area_vals[i]
    cat(" A_cold =", av, "... ")
    # A_cold_target = 0 retained: cold habitat still declines to 0 over experiment.
    # A_warm = 0 retained: null cold-only scenario.
    p   <- modifyList(baseline, list(A_cold = av))
    hdf <- build_habitat(p)
    res <- run_simulation(hdf, p, wt_growth = wt.growth)
    cat("ends", format(max(res$ibm_long$date, na.rm = TRUE)), "\n")
    res
  })
  names(area_sims) <- area_labels
  dir.create("cache", showWarnings = FALSE)
  saveRDS(area_sims, cache_file)
}
```

```{r}
#| label: area-outcomes

area_census   <- bind_rows(Map(oct1_census,   area_sims, area_labels)) |>
  mutate(label = factor(label, levels = area_labels))
area_sr       <- bind_rows(Map(stock_recruit, area_sims, area_labels)) |>
  mutate(label = factor(label, levels = area_labels))
area_thin     <- bind_rows(Map(thinning_rate, area_sims, area_labels)) |>
  mutate(label = factor(label, levels = area_labels))
area_waa      <- bind_rows(Map(wt_at_age,     area_sims, area_labels)) |>
  mutate(label = factor(label, levels = area_labels))
area_age0_dly <- bind_rows(Map(age0_daily_wt, area_sims, area_labels)) |>
  mutate(label = factor(label, levels = area_labels))
area_stable   <- stability_stats(area_census)
```

### Abundance and biomass

```{r}
#| label: fig-area-ts
#| fig-cap: "Oct 1 abundance and biomass by cold habitat area. All scenarios decline to extinction as A_cold → 0; larger starting area supports a larger population and may delay collapse."
#| fig-height: 5

p_N <- ggplot(area_census, aes(x = year, y = N, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = area_cols) +
  labs(x = NULL, y = "N (Oct 1)", color = NULL, title = "Annual abundance") +
  theme_sens() + theme(legend.position = "none")

p_B <- ggplot(area_census, aes(x = year, y = biomass_kg, color = label)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.4) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = area_cols) +
  labs(x = NULL, y = "Total biomass (kg)", color = NULL, title = "Total biomass") +
  theme_sens()

p_N | p_B
```

### Compensatory dynamics

```{r}
#| label: fig-area-comp
#| fig-cap: "Left: burn-in population map. Center: stock-recruitment. Right: critical-period survival rate vs cohort size."
#| fig-height: 5

area_map <- pop_map(area_census)

p_map <- ggplot(area_map, aes(x = N, y = N_next, color = label)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.75, linewidth = 0.9) +
  scale_color_manual(values = area_cols) +
  labs(x = "N (year t, Oct 1)", y = "N (year t+1, Oct 1)", color = NULL,
       title = "Population map (burn-in)") +
  theme_sens() + theme(legend.position = "none")

p_sr <- ggplot(area_sr |> filter(spawner_kg > 0),
               aes(x = spawner_kg, y = n_surv60, color = label)) +
  geom_point(alpha = 0.5, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_color_manual(values = area_cols) +
  labs(x = "Spawner biomass (kg)", y = "Fry surviving 60 days", color = NULL,
       title = "Stock-recruitment") +
  theme_sens()

p_thin <- ggplot(area_thin, aes(x = n_born, y = surv_rate, color = label)) +
  geom_point(alpha = 0.45, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = area_cols) +
  labs(x = "Fry born", y = "% surviving 60-day critical period",
       color = NULL, title = "Critical-period thinning",
       subtitle = "Larger area = lower eff_density = weaker thinning (same interaction as high K)") +
  theme_sens() + theme(legend.position = "none")

p_map | p_sr | p_thin
```

### Age structure

```{r}
#| label: fig-area-age0
#| fig-cap: "Left: mean weight at age (burn-in). Right: age-0 daily weight by day of year, averaged across burn-in cohorts."
#| fig-height: 5

p_waa <- ggplot(area_waa, aes(x = age_yrs, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  scale_color_manual(values = area_cols) +
  scale_x_continuous(breaks = 0:4) +
  labs(x = "Age (years)", y = "Mean weight (g, Oct 1)",
       color = NULL, title = "Weight at age (burn-in mean)",
       subtitle = "Larger area = lower competition = better individual growth") +
  theme_sens()

p_aget <- ggplot(area_age0_dly |> filter(doy >= 118),
                 aes(x = doy, y = mean_wt, color = label)) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = area_cols) +
  scale_x_continuous(breaks = month_doys, labels = month_labels) +
  labs(x = NULL, y = "Mean weight (g)", color = NULL,
       title = "Age-0 daily weight — burn-in cohort mean") +
  theme_sens()

p_waa | p_aget
```

### Stability summary {.unnumbered}

```{r}
#| label: fig-area-stable
#| fig-cap: "Stability by cold habitat area."
#| fig-height: 5

plot_stability_fig(area_stable, area_cols)
```

```{r}
#| label: tbl-area-stable
#| tbl-cap: "Stability statistics by cold habitat area. Experiment years are comparable across scenarios only if populations survive long enough; earlier collapse at smaller area is expected."

area_stable |>
  knitr::kable(col.names = c("A_cold", "Phase", "Mean N",
                              "CV (abundance)", "Mean biomass (kg)", "Years"))
```

---

## Size and age structured competition (`age_structured_competition` × `dominance_beta`)

`age_structured_competition` controls whether fry (age-0) and older fish (age-1+)
compete as separate pools. When `TRUE`, fry compete only with other fry and adults
only with other adults, reflecting ontogenetic habitat segregation (channel margins
vs. pools and runs). When `FALSE`, all fish in a patch compete in one pool.

The effect of this toggle depends critically on `dominance_beta`:

- At **β = 1** adults are already near-immune to fry competition — a fry imposes a
  competitive cost of `(m_fry / m_adult)^1 ≈ 0` on a large adult regardless of
  age-class structuring. The toggle has little practical effect.
- At **β = 0.5** a fry imposes `(m_fry / m_adult)^0.5 ≈ 0.05` — non-trivial — so
  adults feel meaningful fry competition when structuring is off.

This section uses a **2×2 design** (age_structured × dominance_beta) to make the
interaction explicit. `sim_crit` (TRUE, β = 1) and `sim_b05` (TRUE, β = 0.5) are
aliased from `beta_sims`; two new simulations complete the grid.
All other parameters at baseline (`K = 200`, `egg_surv = 0.1`, `crit_period_days = 60`).

```{r}
#| label: sim-agestr
#| cache: false

# Pull the structured runs directly from beta_sims (baseline has
# age_structured_competition = TRUE, so these are identical to sim_crit / sim_b05)
sim_crit <- beta_sims[["β = 1"]]
sim_b05  <- beta_sims[["β = 0.5"]]

cache_file <- "cache/agestr_sims.rds"
if (file.exists(cache_file)) {
  cat("Loading age-structure simulations from cache...\n")
  agestr_cache     <- readRDS(cache_file)
  sim_agestr_F_b05 <- agestr_cache$sim_agestr_F_b05
  sim_agestr_F_b1  <- agestr_cache$sim_agestr_F_b1
} else {
  cat("Running age-class competition 2x2 simulations...\n")
  sim_agestr_F_b05 <- {
    cat(" age_structured=FALSE, beta=0.5 ... ")
    p   <- modifyList(baseline, list(age_structured_competition = FALSE,
                                     dominance_beta             = 0.5))
    res <- run_simulation(build_habitat(p), p, wt_growth = wt.growth)
    cat("ends", format(max(res$ibm_long$date, na.rm = TRUE)), "\n")
    res
  }
  sim_agestr_F_b1 <- {
    cat(" age_structured=FALSE, beta=1.0 ... ")
    p   <- modifyList(baseline, list(age_structured_competition = FALSE,
                                     dominance_beta             = 1))
    res <- run_simulation(build_habitat(p), p, wt_growth = wt.growth)
    cat("ends", format(max(res$ibm_long$date, na.rm = TRUE)), "\n")
    res
  }
  dir.create("cache", showWarnings = FALSE)
  saveRDS(list(sim_agestr_F_b05 = sim_agestr_F_b05,
               sim_agestr_F_b1  = sim_agestr_F_b1),
          cache_file)
}
```

```{r}
#| label: agestr-outcomes

agestr_sims <- list(
  "Structured, β = 1\n(baseline)" = sim_crit,
  "Structured, β = 0.5"           = sim_b05,
  "Unstructured, β = 1"           = sim_agestr_F_b1,
  "Unstructured, β = 0.5"         = sim_agestr_F_b05
)
agestr_labels <- names(agestr_sims)

# Colour = beta; linetype = competition structure
agestr_cols <- c("β = 1" = "#D55E00", "β = 0.5" = "#0072B2")
agestr_lty  <- c("Structured" = "solid", "Unstructured" = "dashed")

mk_agestr <- function(df)
  df |>
    mutate(beta_lab = if_else(grepl("0\\.5", label), "β = 0.5", "β = 1"),
           str_lab  = if_else(grepl("Unstr", label), "Unstructured", "Structured"))

agestr_census <- bind_rows(Map(oct1_census,   agestr_sims, agestr_labels)) |>
  mutate(label = factor(label, levels = agestr_labels)) |> mk_agestr()
agestr_thin   <- bind_rows(Map(thinning_rate, agestr_sims, agestr_labels)) |>
  mutate(label = factor(label, levels = agestr_labels)) |> mk_agestr()
agestr_waa    <- bind_rows(Map(wt_at_age,     agestr_sims, agestr_labels)) |>
  mutate(label = factor(label, levels = agestr_labels)) |> mk_agestr()
agestr_stable <- stability_stats(agestr_census)
```

### Abundance over time and population map

```{r}
#| label: fig-agestr-ts
#| fig-cap: "Oct 1 abundance and burn-in population map for the 2×2 age-structured × beta design. Colour = beta; linetype = competition structure."
#| fig-height: 9

p_ts <- ggplot(agestr_census,
               aes(x = year, y = N, color = beta_lab, linetype = str_lab)) +
  annotate("rect", xmin = burn_start, xmax = exp_start_yr,
           ymin = -Inf, ymax = Inf, fill = "grey92", alpha = 0.7) +
  geom_vline(xintercept = exp_start_yr, linetype = "dashed",
             color = "grey55", linewidth = 0.35) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = agestr_cols) +
  scale_linetype_manual(values = agestr_lty) +
  labs(x = NULL, y = "N (Oct 1)", color = "dominance_beta",
       linetype = "Competition structure",
       title = "A: Annual abundance",
       subtitle = "Divergence between solid and dashed of same colour shows effect of age-class structuring") +
  theme_sens()

agestr_map <- agestr_census |>
  filter(phase == "burn-in") |>
  group_by(label, beta_lab, str_lab) |> arrange(year) |>
  mutate(N_next = lead(N)) |> ungroup() |> filter(!is.na(N_next))

p_map <- ggplot(agestr_map,
                aes(x = N, y = N_next, color = beta_lab, linetype = str_lab)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.75, linewidth = 0.9) +
  scale_color_manual(values = agestr_cols) +
  scale_linetype_manual(values = agestr_lty) +
  labs(x = "N (year t, Oct 1)", y = "N (year t+1, Oct 1)",
       color = "dominance_beta", linetype = "Competition structure",
       title = "B: Population map (burn-in only)") +
  theme_sens()

p_ts / p_map
```

### Critical-period thinning and weight at age

```{r}
#| label: fig-agestr-thin
#| fig-cap: "Thinning rate and weight-at-age for the 2×2 design."
#| fig-height: 5

p_thin <- ggplot(agestr_thin,
                 aes(x = n_born, y = surv_rate,
                     color = beta_lab, linetype = str_lab)) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.7, linewidth = 0.9) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_color_manual(values = agestr_cols) +
  scale_linetype_manual(values = agestr_lty) +
  labs(x = "Fry born", y = "% surviving 60 days",
       color = "dominance_beta", linetype = "Competition structure",
       title = "Critical-period thinning") +
  theme_sens() + theme(legend.position = "none")

p_waa <- ggplot(agestr_waa,
                aes(x = age_yrs, y = mean_wt,
                    color = beta_lab, linetype = str_lab)) +
  geom_line(linewidth = 0.9) + geom_point(size = 2.5) +
  scale_color_manual(values = agestr_cols) +
  scale_linetype_manual(values = agestr_lty) +
  scale_x_continuous(breaks = 0:4) +
  labs(x = "Age (years)", y = "Mean weight (g, Oct 1)",
       color = "dominance_beta", linetype = "Competition structure",
       title = "Weight at age (burn-in mean)") +
  theme_sens()

p_thin | p_waa
```

### Stability summary {.unnumbered}

```{r}
#| label: tbl-agestr-stable
#| tbl-cap: "Stability statistics for the 2×2 age-class structured competition design."

agestr_stable |>
  knitr::kable(col.names = c("Scenario", "Phase", "Mean N",
                              "CV (abundance)", "Mean biomass (kg)", "Years"))
```

---

## Cross-parameter summary

```{r}
#| label: fig-cv-summary
#| fig-width: 13
#| fig-cap: "Mean Oct 1 abundance, CV of abundance, and years simulated during the burn-in phase, across all parameter values. Years simulated < 40 indicates the population collapsed before completing the burn-in."

param_levels <- c("K_cold", "egg_surv", "dominance_beta",
                  "pcmax threshold", "crit_period_days",
                  "pcmax_cold/warm", "A_cold")
param_labels <- c("K", "egg_surv", "dom.\nbeta",
                  "pcmax\nthreshold", "crit_period\ndays",
                  "pcmax\ncold/warm", "A_cold")

# Burn-in stability (mean N, CV) joined with total years across both phases
total_yrs_all <- bind_rows(
  total_sim_yrs(K_census)      |> mutate(parameter = "K_cold"),
  total_sim_yrs(egg_census)    |> mutate(parameter = "egg_surv"),
  total_sim_yrs(beta_census)   |> mutate(parameter = "dominance_beta"),
  total_sim_yrs(thresh_census) |> mutate(parameter = "pcmax threshold"),
  total_sim_yrs(days_census)   |> mutate(parameter = "crit_period_days"),
  total_sim_yrs(pcmax_census)  |> mutate(parameter = "pcmax_cold/warm"),
  total_sim_yrs(area_census)   |> mutate(parameter = "A_cold")
)

cv_all <- bind_rows(
  K_stable      |> filter(phase == "burn-in") |> mutate(parameter = "K_cold"),
  egg_stable    |> filter(phase == "burn-in") |> mutate(parameter = "egg_surv"),
  beta_stable   |> filter(phase == "burn-in") |> mutate(parameter = "dominance_beta"),
  thresh_stable |> filter(phase == "burn-in") |> mutate(parameter = "pcmax threshold"),
  days_stable   |> filter(phase == "burn-in") |> mutate(parameter = "crit_period_days"),
  pcmax_stable  |> filter(phase == "burn-in") |> mutate(parameter = "pcmax_cold/warm"),
  area_stable   |> filter(phase == "burn-in") |> mutate(parameter = "A_cold")
) |>
  left_join(total_yrs_all, by = c("label", "parameter")) |>
  mutate(parameter = factor(parameter, levels = param_levels,
                            labels = param_labels))

cv_all_long <- cv_all |>
  pivot_longer(c(mean_N, cv, total_yrs),
               names_to  = "metric",
               values_to = "value") |>
  mutate(metric = factor(metric,
                         levels = c("mean_N", "cv", "total_yrs"),
                         labels = c("Mean N (Oct 1)", "CV (abundance)",
                                    "Years simulated\n(max = 100)")))

param_fill <- c("#0072B2", "#009E73", "#D55E00", "#CC79A7", "#E69F00",
                "#56B4E9", "#999999")
names(param_fill) <- param_labels

ggplot(cv_all_long, aes(x = label, y = value, fill = parameter)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  # Year 40: end of burn-in (bars below = collapsed during burn-in)
  geom_hline(
    data = \(d) filter(d, metric == "Years simulated\n(max = 100)"),
    aes(yintercept = 40),
    linetype = "dotted", color = "grey35", linewidth = 0.6
  ) +
  # Year 100: end of full simulation (bars below = collapsed during experiment)
  geom_hline(
    data = \(d) filter(d, metric == "Years simulated\n(max = 100)"),
    aes(yintercept = 100),
    linetype = "dashed", color = "grey35", linewidth = 0.6
  ) +
  facet_grid(metric ~ parameter, scales = "free") +
  scale_fill_manual(values = param_fill) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(x = NULL, y = NULL,
       title = "Burn-in stability across all parameter sensitivity analyses",
       subtitle = paste0(
         "Row 1: burn-in mean N.  Row 2: burn-in CV (lower = more stable).  ",
         "Row 3: total years simulated (burn-in + experiment). ",
         "Dotted = yr 40 (burn-in end); dashed = yr 100 (full simulation). "
         )) +
  theme_bw() +
  theme(strip.background  = element_rect(fill = "grey88"),
        strip.text.x      = element_text(size = 8),
        strip.text.y      = element_text(size = 7.5, lineheight = 0.85),
        panel.grid.minor  = element_blank(),
        axis.text.x       = element_text(size = 6.5, lineheight = 0.8),
        plot.title        = element_text(size = 10, face = "bold"),
        plot.subtitle     = element_text(size = 7.5, color = "grey30"),
        panel.spacing     = unit(0.6, "lines"))
```

### Age-class structured competition — 2×2 stability {.unnumbered}

The age-class section uses a factorial design rather than a single-axis sweep, so
it is not directly comparable to the single-parameter bar charts above. The table
below summarises burn-in stability for all four combinations.

```{r}
#| label: tbl-agestr-summary
#| tbl-cap: "Burn-in stability for the 2×2 age-structured × dominance_beta design. Rows where years simulated < 100 indicate premature collapse."

agestr_stable |>
  left_join(
    bind_rows(Map(function(s, l)
      tibble(label = l, total_yrs = n_distinct(
        year(s$ibm_long$date[!is.na(s$ibm_long$date)]))),
      agestr_sims, agestr_labels)),
    by = "label"
  ) |>
  filter(phase == "burn-in") |>
  select(label, mean_N, cv, total_yrs) |>
  knitr::kable(col.names = c("Scenario", "Burn-in mean N",
                              "Burn-in CV", "Total years simulated"))
```


