---
title: "Simple IBM"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(ggpubr)
library(egg)
```

## Simulate habitat

### Temperature

Simulate thermal regimes as annual sine waves. Cold vs. Warm using simple multiplier.

```{r}
# Set seed for reproducibility
set.seed(123)

# 1. Create a date range (1 year)
dates <- seq(as.Date("2024-01-01"), as.Date("2024-12-31"), by="day")
n <- length(dates)
doy <- as.numeric(format(dates, "%j")) # Day of year

# 2. Simulate Stream Temperature (Sine wave + Noise)
# Parameters: Midpoint, Amplitude, Offset
base_temp <- 10        # Average annual temperature
amplitude <- 12         # Seasonal fluctuation amplitude
# Peak temperature occurs late in the year (July/Aug)
stream_temp <- base_temp + amplitude * sin(2 * pi * (doy - 100) / 365)

# 3. Add random noise (daily weather variation)
noise <- rnorm(n, mean = 0, sd = 0) # no variation for simplicity
simulated_temp <- stream_temp + noise

# 4. Force values near 0 to small postive, max = 25
simulated_temp[simulated_temp < 0.5] <- 0.5
simulated_temp[simulated_temp > 25] <- 25

# 4. Simulate cold regime using a multiplier
simulated_temp_cold <- simulated_temp*0.6

# 4. Create Data Frame
habitat_df <- tibble(date = dates, 
                     doy = 1:length(dates),#doy, 
                     temp_warm = round(simulated_temp, digits = 1), 
                     temp_cold = round(simulated_temp_cold, digits = 1))

# 5. Plot the simulation
habitat_df %>% ggplot() + 
  geom_line(aes(x = date, y = temp_warm), color = "red") +
  geom_line(aes(x = date, y = temp_cold), color = "blue") +
  theme_bw() + 
  xlab("Date") + 
  ylab("Simulated daily stream temperature (deg C)")


# function to retrieve water temperature in patch = patch and at doy/time = t
get_patchtemp <- function(t, patch) {
  if (patch == "warm") {
    habitat_df$temp_warm[habitat_df$doy == t]
  } else {
    habitat_df$temp_cold[habitat_df$doy == t]
  }
}
```

### Ration

```{r}
# define habitat specific ration size
ration_cold <- 0.10
ration_warm <- 0.10

# define random variation in ration size
ration_sd <- 0.00 # no variation for simplicity

# function to generate patch food availability at time t (stochastic)
patch_food <- function(patch) {
  if (patch == "warm") {
    max(0, min(1, rnorm(1, ration_warm, ration_sd)))
  } else {
    max(0, min(1, rnorm(1, ration_cold, ration_sd)))
  }
}

# define food in habitat tibble
habitat_df <- habitat_df %>% 
  mutate(ration_warm = replicate(n = dim(.)[1], patch_food("warm")),
         ration_cold = replicate(n = dim(.)[1], patch_food("cold")))

# plot ration size
habitat_df %>% ggplot() + 
  geom_line(aes(x = date, y = ration_warm), color = "red") +
  geom_line(aes(x = date, y = ration_cold), color = "blue") +
  theme_bw() + 
  xlab("Date") + 
  ylab("Simulated daily ration size")


# function to retrieve ration in patch = patch and at doy/time = t
get_patchration <- function(t, patch) {
  if (patch == "warm") {
    habitat_df$ration_warm[habitat_df$doy == t]
  } else {
    habitat_df$ration_cold[habitat_df$doy == t]
  }
}
```

View habitat data structure

```{r}
str(habitat_df)
head(habitat_df)
```

## Basic functions

Rescale a vector to a specific range...e.g., \[0-1\].

```{r}
# Rescale a vector to a specific
fncRescale <- function(x, to = c(0, 1), from = range(x, na.rm = TRUE, finite = TRUE)) {
    (x - from[1]) / diff(from) * diff(to) + to[1]
}
```

## BioEnergetics model

Build bioenergetics model, equations from Hanson et al. (1997), and general approach from A. Fullerton (<https://github.com/aimeefullerton/growth_regime_IBM>). Currently using *O. mykiss* parameters because the full suite of respiration estimates does not exist for bull trout, and hence no effect of ration.

Get BioEnergetics Parameters

```{r}
fncGetBioEParms <- function(spp, pred.en.dens, prey.en.dens, oxy, pff, wt.nearest, 
  startweights = rep(initial.mass, numFish), pvals = rep(0.5, numFish), ration = rep(0.1, numFish)){
  
	N.sites <- nrow(wt.nearest) #here, sites are fish in each reach
  N.steps <- 1 #one time step
  Species <- spp
  SimMethod <- 1 #method that predicts growth
  Pred <- pred.en.dens #predator energy density
  Oxygen <- oxy #oxygen consumed
	PFF <- pff #percent indigestible prey
  stab.factor <- 0.5 #stability factor for other simulation methods
  epsilon <- 0.5 #also for other simulation methods
	endweights <- startweights*5
	TotalConsumption <- rep(100, N.sites)
	pvalues <- t(matrix(round(pvals, 5)))
	sitenames <- t(matrix(wt.nearest[,"pid"]))
	temperature <- t(wt.nearest[,"WT"])
  prey.energy.density <- t(matrix(rep(prey.en.dens, N.sites))) 
  ration <- t(matrix(round(ration, 5)))
  
  return(list(	
    "Species" = Species,
    "SimMethod" = SimMethod,
    "Wstart" = startweights,
    "Endweights" = endweights,
    "TotalConsumption" = TotalConsumption,
    "pp "= pvalues,
    "Temps" = temperature,
    "N.sites" = N.sites,
    "N.steps" = N.steps,
    "sitenames" = sitenames,
    "Pred" = Pred,
    "prey.energy.density" = prey.energy.density,
    "Oxygen" = Oxygen,
    "stab.factor" = stab.factor,
    "PFF" = PFF,
    "epsilon" = epsilon,
    "ration" = ration)
  )
  
}
```

Constants, hard-coded for *O. mykiss*. Could alternatively do for bull trout (constants commented out) but we don't have parameter estimates for respiration equations that would allow growth to vary by ration size.

```{r}
# Constants, hard-coded - using RATION instead of p-values
fncReadConstants <- function() {
  
  # Get consumption constants
  # Cons <- data.frame(
  #   ConsEQ = 3,
  #   CA = 0.1317,
  #   CB = -0.1396,
  #   CQ = 3.0,
  #   CTO = 15.8,
  #   CTM = 17.5,
  #   CTL = 21.0,
  #   CK1 = 0.06,
  #   CK4 = 0.38)
  # 
  # # Get respiration constants
  # Resp <- data.frame(
  #   RespEQ = 1, 
  #   RA = 0.0009,
  #   RB = -0.1266,
  #   RQ = 0.0833,
  #   RTO = 0,
  #   RTM = 0,
  #   RTL = 0,
  #   RK1 = 1,
  #   RK4 = 0,
  #   ACT = 1,
  #   BACT = 0,
  #   SDA = 0.172)
  # 
  # # Get Excretion / Egestion Constants
  # Excr <- data.frame(
  #   ExcrEQ = 2, 
  #   FA = 0.212,  
  #   FB = -0.222, 
  #   FG = 0.631, 
  #   UA = 0.0314, 
  #   UB = 0.58, 
  #   UG = -0.299)
  
    Cons <- data.frame(
    ConsEQ = 4,
    CA = 0.628,
    CB = -0.3,
    CQ = 5,
    CTO = 20,
    CTM = 20,
    CTL = 24,
    CK1 = 0.33,
    CK4 = 0.2)
  
  # Get respiration constants
  Resp <- data.frame(
    RespEQ = 1, 
    RA = 0.00264,
    RB = -0.217,
    RQ = 0.06818,
    RTO = 0.0234,
    RTM = 0,
    RTL = 25,
    RK1 = 1,
    RK4 = 0.13,
    ACT = 9.7,
    BACT = 0.0405,
    SDA = 0.172)
  
  # Get Excretion / Egestion Constants
  Excr <- data.frame(
    ExcrEQ = 4, 
    FA = 0.212,  
    FB = -0.222, 
    FG = 0.631, 
    UA = 0.0314, 
    UB = 0.58, 
    UG = -0.299)
  
  # Return the Constants
  return(list("Consumption" = Cons, 
              "Respiration" = Resp, 
              "Excretion" = Excr))
}
```

Consumption, respiration, and excretion equations

```{r}
# Consumption Equation 1
ConsumptionEQ1 <- function(W, TEMP, PP, PREY, CA, CB, CQ) {

	CMAX <- CA * (W ** CB)			#max specific feeding rate (g_prey/g_pred/d)
	CONS <<- (CMAX * PP * exp(CQ * TEMP))	#specific consumption rate (g_prey/g_pred/d) - grams prey consumed per gram of predator mass per day
	CONSj <<- CONS * PREY          		 #specific consumption rate (J/g_pred/d) - Joules consumed for each gram of predator for each day
	return(list("CMAX" = CMAX, "CONS" = CONS, "CONSj" = CONSj))
}

# Consumption Equation 2
ConsumptionEQ2 <- function(W, TEMP, PP, PREY, CA, CB, CTM, CTO, CQ) {

	Y <- log(CQ) * (CTM - CTO + 2)
	Z <- log(CQ) * (CTM - CTO)
	X <- (Z ^ 2 * (1 + (1 + 40 / Y) ^ .5) ^ 2) / 400
	V <- (CTM - TEMP) / (CTM - CTO)
	CMAX <- CA * (W ** CB)		
	CONS <- CMAX * PP * (V ** X) * exp(X * (1 - V))
	CONSj <- CONS*PREY             

	return(list("CMAX" = CMAX, "CONS" = CONS, "CONSj" = CONSj))
}

# Consumption Equation #3: Temperature Dependence for cool-cold water species
ConsumptionEQ3 <- function(W, TEMP, PP, PREY, CA, CB, CK1, CTO, CQ, CK4, CTL, CTM) {
	G1 <- (1 / (CTO - CQ)) * (log((0.98 * (1 - CK1)) / (CK1 * 0.02)))
	L1 <- exp(G1 * (TEMP - CQ))
	KA <- (CK1 * L1) / (1 + CK1 * (L1 - 1))
	G2 <- (1 / (CTL - CTM)) * (log((0.98 * (1 - CK4)) / (CK4 * 0.02)))
	L2 <- exp(G2 * (CTL - TEMP))
	KB <- (CK4 * L2) / (1 + CK4 * (L2 - 1))
	CMAX <- CA * (W ** CB)		#max specific feeding rate (g_prey/g_pred/d)
	
	CONS <- CMAX * PP * KA * KB		#specific consumption rate (g_prey/g_pred/d) - grams prey consumed per gram of predator mass per day
	
	CONSj <- CONS * PREY             #specific consumption rate (J/g_pred/d) - Joules consumed for each gram of predator for each day

	return(list("CMAX" = CMAX, "CONS" = CONS, "CONSj" = CONSj))
	}

# Consumption Equation 3 that gives CONS based on RATION instead of PP
ConsumptionEQ4 <- function(W, TEMP, RATION, PREY, CA, CB, CK1, CTO, CQ, CK4, CTL, CTM) {
  G1 <- (1 / (CTO - CQ)) * (log((0.98 * (1 - CK1)) / (CK1 * 0.02)))
  L1 <- exp(G1 * (TEMP - CQ))
  KA <- (CK1 * L1) / (1 + CK1 * (L1 - 1))
  G2 <- (1 / (CTL - CTM)) * (log((0.98 * (1 - CK4)) / (CK4 * 0.02)))
  L2 <- exp(G2 * (CTL - TEMP))
  KB <- (CK4 * L2) / (1 + CK4 * (L2 - 1))
  CMAX <- CA * (W ** CB)		#max specific feeding rate (g_prey/g_pred/d)
  
  RATION[RATION >= CMAX] <- CMAX[RATION >= CMAX]
  CONS <- RATION * KA * KB  #specific consumption rate (g_prey/g_pred/d) - grams prey consumed per gram of predator mass per day
  
  CONSj <- CONS * PREY  #specific consumption rate (J/g_pred/d) - Joules consumed for each gram of predator for each day
  
  return(list("CMAX" = CMAX, "CONS" = CONS, "CONSj" = CONSj))
}

# Excretion Equation 1
ExcretionEQ1 <- function(CONS, CONSj, TEMP, PP, FA, UA) {

	EG <- FA * CONS				# egestion (fecal waste) in g_waste/g_pred/d
	U <- UA * (CONS - EG)	 			# excretion (nitrogenous waste) in g_waste/g_pred/d

	EGj <- FA * CONSj				# egestion in J/g/d
	Uj <- UA * (CONSj - EGj)			# excretion in J/g/d

	return(list("EG" = EG, "EGj" = EGj, "U" = U, "Uj" = Uj))
	}	

# Excretion Equation 2
ExcretionEQ2 <- function(CONS, CONSj, TEMP, PP, FA, UA, FB, FG, UB, UG) {

	EG <- FA * TEMP ^ FB * exp(FG * PP) * CONS			# egestion (fecal waste) in g_waste/g_pred/d
	U <- UA * TEMP ^ UB * exp(UG * PP) * (CONS - EG)			# excretion (nitrogenous waste) in g_waste/g_pred/d

	EGj <- EG * CONSj / CONS					# egestion in J/g/d
	Uj <- U * CONSj / CONS					# excretion in J/g/d

	return(list("EG" = EG, "EGj" = EGj, "U" = U, "Uj" = Uj))
	}

# Excretion Equation 3 (W/ correction for indigestible prey as per Stewart 1983)
ExcretionEQ3 <- function(CONS, CONSj, TEMP, PP, FA, UA, FB, FG, UB, UG, PFF) {

	#Note: In R, "F" means "FALSE", here we use EG as the variable name for egestion instead of F (as in the FishBioE 3.0 manual)
	#Note:  PFF = 0 assumes prey are entirely digestible, making this essentially the same as Equation 2

	PE <- FA * (TEMP ** FB) * exp(FG * PP)

	PF <- ((PE - 0.1) / 0.9) * (1 - PFF) + PFF

	EG <- PF * CONS					# egestion (fecal waste) in g_waste/g_pred/d
	U <- UA * (TEMP ** UB) * (exp(UG * PP)) * (CONS - EG)	# excretion (nitrogenous waste) in g_waste/g_pred/d

	EGj <- PF * CONSj				# egestion in J/g/d
	Uj <- UA * (TEMP ** UB) * (exp(UG * PP)) * (CONSj - EGj)	# excretion in J/g/d

	return(list("EG" = EG, "EGj" = EGj, "U" = U, "Uj" = Uj))
	}	

# Excretion Equation 3 but using RATION instead of PP
ExcretionEQ4 <- function(CONS, CONSj, TEMP, RATION, CMAX, FA, UA, FB, FG, UB, UG, PFF) {
  
  #Note: In R, "F" means "FALSE", here we use EG as the variable name for egestion instead of F (as in the FishBioE 3.0 manual)
  #Note:  PFF = 0 assumes prey are entirely digestible, making this essentially the same as Equation 2
  
  RATION[RATION >= CMAX] <- CMAX[RATION >= CMAX]
  PP <- RATION / CMAX #calculating p-value based on ration and CMax inputs
  
  PE <- FA * (TEMP ** FB) * exp(FG * PP)
  
  PF <- ((PE - 0.1) / 0.9) * (1 - PFF) + PFF
  
  EG <- PF * CONS					# egestion (fecal waste) in g_waste/g_pred/d
  U <- UA * (TEMP ** UB) * (exp(UG * PP)) * (CONS - EG)	# excretion (nitrogenous waste) in g_waste/g_pred/d
  
  EGj <- PF * CONSj				# egestion in J/g/d
  Uj <- UA * (TEMP ** UB) * (exp(UG * PP)) * (CONSj - EGj)	# excretion in J/g/d
  
  return(list("EG" = EG, "EGj" = EGj, "U" = U, "Uj" = Uj))
}	

# Respiration Equation 1
RespirationEQ1 <- function(W, TEMP, CONS, EG, PREY, OXYGEN, RA, RB, ACT, SDA, RQ, RTO, RK1, RK4, RTL, BACT){

	VEL <- (RK1 * W ^ RK4) * (TEMP > RTL) + ACT * W ^ RK4 * exp(BACT * TEMP) * (1 - 1 * (TEMP > RTL))
	ACTIVITY <- exp(RTO * VEL)
	S <- SDA * (CONS - EG)					# proportion of assimilated energy lost to SDA in g/g/d (SDA is unitless)
	Sj <- S * PREY						# proportion of assimilated energy lost to SDA in J/g/d - Joules lost to digestion per gram of predator mass per day
	R <- RA * (W ** RB) * ACTIVITY * exp(RQ * TEMP)       	# energy lost to respiration (metabolism) in g/g/d
	Rj <- R * OXYGEN           				# energy lost to respiration (metabolism) in J/g/d - Joules per gram of predator mass per day
	return(list("R" = R, "Rj" = Rj, "S" = S, "Sj" = Sj))
	}

# Respiration Equation 2 (Temp dependent w/ ACT multiplier)
RespirationEQ2 <- function(W, TEMP, CONS, EG, PREY, OXYGEN, RA, RB, ACT, SDA, RTM, RTO, RQ) {

	V <- (RTM - TEMP) / (RTM - RTO)
	V[V < 0] <- 0.001 #AHF Added to stop errors when Water temps exceed RTM!
	Z <- (log(RQ)) * (RTM - RTO)
	Y <- (log(RQ)) * (RTM - RTO + 2)
	X <- ((Z ** 2) * (1 + (1 + 40 / Y) ** 0.5) ** 2) / 400
	S <- SDA * (CONS - EG)					# proportion of assimilated energy lost to SDA in g/g/d (SDA is unitless)
	Sj <- S * PREY						# proportion of assimilated energy lost to SDA in J/g/d - Joules lost to digestion per gram of predator mass per day
	R <- RA * (W ** RB) * ACT * ((V ** X) * (exp(X * (1 - V) )))     	# energy lost to respiration (metabolism) in g/g/d
	Rj <- R * OXYGEN           				# energy lost to respiration (metabolism) in J/g/d - Joules per gram of predator mass per day
	return(list("R" = R, "Rj" = Rj, "S" = S, "Sj" = Sj))
	}
```

Calculate growth

```{r}
CalculateGrowth <- function(Constants, Input) {
  
  pred <- Input$Pred
  Oxygen <- Input$Oxygen
  
  # Initialize Fish Weights
  W <- array(rep(0, (Input$N.sites * (Input$N.steps + 1))), c(Input$N.steps + 1, Input$N.sites))
  W[1,] <- as.numeric(Input$Wstart[1:ncol(W)])
  Growth <- array(rep(0, Input$N.sites * Input$N.steps), c(Input$N.steps, Input$N.sites))
  Growth_j <- Consumpt <- Consumpt_j <- Excret <- Excret_j <- Egest <- Egest_j <- Respirat <- Respirat_j <- S.resp <- Sj.resp <- Gg_WinBioE <- Gg_ELR <- Growth
  TotalC <- rep(0, Input$N.sites)
  
  ##Start Looping Through Time - for Known Consumption, solving for Weight
  t = 1
  for (t in 1:(Input$N.steps)) {
    
    ### Consumption 
    if(Constants$Consumption$ConsEQ == 1) {
      Cons <- with(Constants$Consumption, ConsumptionEQ1(W[t,], Input$Temps[t,], Input$pp[t,], Input$prey.energy.density[t,], Constants$Consumption$CA, Constants$Consumption$CB, Constants$Consumption$CQ))
    } else if(Constants$Consumption$ConsEQ == 2){
      Cons <- with(Constants$Consumption, ConsumptionEQ2(W[t,], Input$Temps[t,], Input$pp[t,], Input$prey.energy.density[t,], Constants$Consumption$CA, Constants$Consumption$CB, Constants$Consumption$CTM, Constants$Consumption$CTO, Constants$Consumption$CQ))
    } else if(Constants$Consumption$ConsEQ == 3){
      Cons <- with(Constants$Consumption, ConsumptionEQ3(W[t,], Input$Temps[t,], Input$pp[t,], Input$prey.energy.density[t,], Constants$Consumption$CA, Constants$Consumption$CB, Constants$Consumption$CK1, Constants$Consumption$CTO, Constants$Consumption$CQ, Constants$Consumption$CK4, Constants$Consumption$CTL, Constants$Consumption$CTM))
    } else if(Constants$Consumption$ConsEQ == 4){
      Cons <- with(Constants$Consumption, ConsumptionEQ4(W[t,], Input$Temps[t,], Input$ration[t,], Input$prey.energy.density[t,], Constants$Consumption$CA, Constants$Consumption$CB, Constants$Consumption$CK1, Constants$Consumption$CTO, Constants$Consumption$CQ, Constants$Consumption$CK4, Constants$Consumption$CTL, Constants$Consumption$CTM))
    }
    
    TotalC <- TotalC + Cons$CONS * W[t,]
    
    # store daily consumption 
    Consumpt[t,] <- as.numeric(Cons$CONS)
    Consumpt_j[t,] <- as.numeric(Cons$CONSj)
    
    
    ### Excretion / Egestion
    if(Constants$Excretion$ExcrEQ == 1) {
      ExcEgest <- with(Constants$Excretion, ExcretionEQ1(Cons$CONS, Cons$CONSj, Input$Temps[t,], Input$pp[t,], Constants$Excretion$FA, Constants$Excretion$UA))
    } else if (Constants$Excretion$ExcrEQ == 2) {
      ExcEgest <- with(Constants$Excretion, ExcretionEQ2(Cons$CONS, Cons$CONSj, Input$Temps[t,], Input$pp[t,], Constants$Excretion$FA, Constants$Excretion$UA, Constants$Excretion$FB, Constants$Excretion$FG, Constants$Excretion$UB, Constants$Excretion$UG ))
    } else if (Constants$Excretion$ExcrEQ == 3) {
      ExcEgest <- with(Constants$Excretion, ExcretionEQ3(Cons$CONS, Cons$CONSj, Input$Temps[t,], Input$pp[t,], Constants$Excretion$FA, Constants$Excretion$UA, Constants$Excretion$FB, Constants$Excretion$FG, Constants$Excretion$UB, Constants$Excretion$UG, Input$PFF) )
    } else if (Constants$Excretion$ExcrEQ == 4) {
      ExcEgest <- with(Constants$Excretion, ExcretionEQ4(Cons$CONS, Cons$CONSj, Input$Temps[t,], Input$ration[t,], Cons$CMAX, Constants$Excretion$FA, Constants$Excretion$UA, Constants$Excretion$FB, Constants$Excretion$FG, Constants$Excretion$UB, Constants$Excretion$UG, Input$PFF) )
    }
    
    # store daily excretion and egestion
    Excret[t,] <- as.numeric(ExcEgest$U)
    Excret_j[t,] <- as.numeric(ExcEgest$Uj)
    Egest[t,] <- as.numeric(ExcEgest$EG)
    Egest_j[t,] <- as.numeric(ExcEgest$EGj)
    
    
    ### Respiration
    if(Constants$Respiration$RespEQ == 1) {
      Resp <- with(Constants$Respiration, RespirationEQ1(W[t,], Input$Temps[t,], Cons$CONS, ExcEgest$EG, Input$prey.energy.density[t,], Input$Oxygen, Constants$Respiration$RA, Constants$Respiration$RB, Constants$Respiration$ACT, Constants$Respiration$SDA, Constants$Respiration$RQ,Constants$Respiration$RTO, Constants$Respiration$RK1, Constants$Respiration$RK4, Constants$Respiration$RTL, Constants$Respiration$BACT))
    } else if (Constants$Respiration$RespEQ == 2) {
      Resp <- with(Constants$Respiration, RespirationEQ2(W[t,], Input$Temps[t,], Cons$CONS, ExcEgest$EG, Input$prey.energy.density[t,], Input$Oxygen, Constants$Respiration$RA, Constants$Respiration$RB, Constants$Respiration$ACT, Constants$Respiration$SDA, Constants$Respiration$RTM, Constants$Respiration$RTO, Constants$Respiration$RQ))
    }
    
    #store daily respiration results
    Respirat[t,] <- as.numeric(Resp$R)
    Respirat_j[t,] <- as.numeric(Resp$Rj)
    S.resp[t,] <- as.numeric(Resp$S)
    Sj.resp[t,] <- as.numeric(Resp$Sj)
    
    
    ### Now calculate Growth
    
    # growth in J/g/d - Joules allocated to growth for each gram of predator on each day
    Gj <- Cons$CONSj - Resp$Rj - ExcEgest$EGj - ExcEgest$Uj - Resp$Sj	
    G <- Cons$CONS - Resp$R - ExcEgest$EG - ExcEgest$U - Resp$S
    # growth in g/d - Grams of predator growth each day
    Growth[t,] <- as.numeric(Gj * W[t,]) / pred
    Growth_j[t,] <- as.numeric(Gj)
    
    # growth in g/g/d (DailyWeightIncrement divided by fishWeight)
    Gg_WinBioE[t,] <- as.numeric(Growth[t,] / W[t,])			
    
    # growth in g/g/d (DailyWeightIncrement divided by average of fish start end weights)
    Gg_ELR[t,] <- Growth[t,] / ((as.numeric(Input$Wstart[1:ncol(W)]) + W[t,]) / 2)		
    
    # Calculate absolute weight at time t+1
    W[t + 1,] <- W[t,] + Growth[t,]
    
    
  } # End of cycles through time
  
  # Return Results	
  return(list(
    "TotalC" = TotalC,
    "W" = W, 
    "Growth" = Growth, 
    "Gg_WinBioE" = Gg_WinBioE, 
    "Gg_ELR" = Gg_ELR,
    "Growth_j" = Growth_j,
    "Consumption" = Consumpt,
    "Consumption_j" = Consumpt_j,
    "Excretion" = Excret,
    "Excretion_j" = Excret_j,
    "Egestion" = Egest,
    "Egestion_j" = Egest_j,
    "Respiration" = Respirat, 
    "Respiration_j" = Respirat_j, 
    "S.resp" = S.resp,
    "Sj.resp" = Sj.resp
  ))
}
```

The main bioenergetics function that calls other functions

```{r}
BioE <- function (Input, Constants) {

# for simulation method =1 (we have p-vals, and want to solve for weights
if (Input$SimMethod == 1) {
  
	# Method 1: Calculate Growth from p-values and Temperatures
	Results <- CalculateGrowth(Constants, Input)
	W <- Results$W
	Growth <- Results$Growth
	
} else{

# We don't know p-values, but need to iteratively solve for them
# Method 2 or 3:  Calculate P-values from Total Growth or Consumption
# Need to assume p-values are constant with time
	
	# initialize first guess p-values of .5
	Input$pp <- array(rep(0.5, Input$N.sites * Input$N.step), 
		 c(Input$N.step, Input$N.sites))

	# set error at high value, iterate until it's small
	Error <- rep(99, Input$N.sites)
	iteration <- 0

### Interate until error is less than .1
	while(max(abs(Error)) > Input$epsilon) 
{
	iteration <- iteration + 1
		Results<- CalculateGrowth(Constants, Input)
		W <- Results$W
    TConsumption<- Results$TotalC

 # Find error (depending on which thing on which we're converging), and
 # and come up with new estimate for average p-value
		if (Input$SimMethod == 2)  {
		      Error <- (W[Input$N.step + 1,] - Input$Endweights)
  # Delta is the amount by which we'll change the p-value (prior to
  # scaling by the stability factor
		  	Delta <- (Input$Endweights) / (W[Input$N.step + 1,]) *
      	         Input$pp[1,] - Input$pp[1,]

		Pnew <- as.vector(Input$p[1,] + Input$stab.factor * Delta)
	# Guard against negative p-values
	 for (i in 1:length(Pnew)) {Pnew[i] = max(0, Pnew[i])}
		} else 
		{
	      Error <- (TConsumption - Input$TotalConsumption) / Input$TotalConsumption
	  	Delta <- Input$TotalConsumption/TConsumption * Input$p[1,] - Input$p[1,]
		Pnew <- as.vector(Input$p[1,] + Input$stab.factor * Delta)
	}
	# Update for user, show p-values and error, see if we're converging
		for (i in 1:Input$N.step) {Input$p[i,] <- as.numeric(Pnew)}
			print(paste("Pnew =",Pnew, "  Error =", Error))
	            print(paste("iteration = ", iteration))	
		} # end of while statement 
	} 

Results$pp = Input$pp
# We're done.  Let's return the results!
return(Results)
}
```

## Movement functions

### Current growth

Look up growth based on actual WT, ration, and weight of fish

```{r}
# PIDs: vector of fish IDs, or NA for all fish
# fish: data frame with columns pid, weight, ration, patch ("warm" or "cold")
# t: current day of year
fncGrowthFish <- function(PIDs = NA, fish, t) {
  
  # sequences must match the dimensions of the pre-calculated wt.growth array
  wt.seq <- seq(0.1, 25, 0.1)       # water temperature (250 values)
  ra.seq <- seq(0.02, 0.17, 0.001)  # ration (151 values)
  ma.seq <- seq(1, 1500, 1)         # fish mass (1500 values)
  
  if (any(is.na(PIDs))) {
    td <- fish[, c("pid", "weight", "ration", "patch")]  # all fish
  } else {
    td <- fish[fish$pid %in% PIDs, c("pid", "weight", "ration", "patch")]  # specific fish
  }
  td[, "weight"][td[, "weight"] < 1] <- 1  # minimum lookup weight is 1 g

  if (nrow(td) > 0) {
    n <- nrow(td)
    growth <- watemp <- vector(length = n)
    
    for (x in 1:n) {
      temp     <- get_patchtemp(t, td[x, "patch"])           # temperature at fish's current patch
      wt.idx   <- which.min(abs(wt.seq - temp))              # nearest water temperature index
      ra.idx   <- which.min(abs(ra.seq - td[x, "ration"]))   # nearest ration index
      ma.idx   <- which.min(abs(ma.seq - td[x, "weight"]))   # nearest mass index
      growth[x] <- wt.growth[wt.idx, ra.idx, ma.idx]
      watemp[x] <- wt.seq[wt.idx]
    }
    
    lookup <- data.frame(
      pid      = td[, "pid"],
      weight   = td[, "weight"],
      ration   = td[, "ration"],
      patch    = td[, "patch"],
      WT.actual = sapply(td[, "patch"], function(p) get_patchtemp(t, p)),
      WT       = watemp,
      growth   = growth
    )
    return(lookup)
  } else {
    message("No fish IDs provided")
    return(NA)
  }
  
}
```

### Best possible growth

Look up best possible patch for growth based on current temperatures across both habitat patches

```{r}
# fweight: fish mass in grams
# t: current day of year
# ration_warm, ration_cold: ration sizes available in each patch
fncGrowthPossible <- function(fweight, t, ration_warm, ration_cold) {
  
  # sequences must match the dimensions of the pre-calculated wt.growth array
  wt.seq <- seq(0.1, 25, 0.1)       # water temperature (250 values)
  ra.seq <- seq(0.02, 0.17, 0.001)  # ration (151 values)
  ma.seq <- seq(1, 1500, 1)         # fish mass (1500 values)
  
  fw     <- max(1, min(4500, round(fweight)))  # clamp weight to lookup range
  ma.idx <- which.min(abs(ma.seq - fw))

  patches <- c("warm", "cold")
  rations <- c(ration_warm, ration_cold)
  growth  <- watemp <- vector(length = 2)
  
  for (x in 1:2) {
    temp      <- get_patchtemp(t, patches[x])
    wt.idx    <- which.min(abs(wt.seq - temp))
    ra.idx    <- which.min(abs(ra.seq - rations[x]))
    growth[x] <- wt.growth[wt.idx, ra.idx, ma.idx]
    watemp[x] <- wt.seq[wt.idx]
  }
  
  lookup <- data.frame(
    patch    = patches,
    weight   = rep(fweight, 2),
    ration   = rations,
    WT.actual = c(get_patchtemp(t, "warm"), get_patchtemp(t, "cold")),
    WT       = watemp,
    growth   = growth
  )
  
  best_patch <- patches[which.max(growth)]
  
  return(list(lookup = lookup, best_patch = best_patch))
  
}
```

### Movement cost

As in Fullerton model, energetic cost to movement is applied as a percentage of possible growth potential.

```{r}
fncMoveCost <- function(gr = fish$growth) {
  return(abs(gr) * move_cost)
}
```

## Density dependence

Create functions to representing density-dependent food availability. I.e., reduce ration size in patches by patch density.

### Fullerton approach

Ration size decreases linearly with fish density, up to some threshold above which there is no effect of there is no effect of density of ration size. Fullerton uses 1 fish per m\^2 of as the upper threshold. Here, I use an arbitrary *MaxDensity4Growth*, but we can change this to a value that makes sense for our model. Fullerton's IBM calculates this scalar, then multiplies by densities after movement to compute experienced/realized ration.

```{r}
MaxDensity4Growth <- 50 # growth is not depressed further above this density (currently this is arbitary)
fdens_raw <- fdens <- seq(from = 1, to = 100, by = 1) # create sequence of fish density
fdens[fdens > MaxDensity4Growth] <- MaxDensity4Growth # high densities cap out at MaxDensity4Growth
density.effect <- fncRescale((1 - c(fdens, 0.01, MaxDensity4Growth)), c((0.5),1)) # calculate density effect scalar
density.effect <- density.effect[-c(length(density.effect), (length(density.effect) - 1))] #remove the last 2 temporary values

par(mar = c(4.5,4.5,1,1))
plot(density.effect ~ fdens_raw, type = "l", xlab = "Fish density", ylab = "Density effect on ration (multiplier)")
abline(v = MaxDensity4Growth, col = "red", lty = 2)
legend("topright", legend = c("Density effect", "Density threshold"), col = c("black", "red"), lty = c(1,2), bty = "n")
```

### Hyperbolic reduction

Hyperbolic (Beverton-Holt style) reduction in food availability, i.e., scramble competition. This function can allow the strength of competition to differ among habitats.

-   k = half-saturation density: fish density at which ration is halved
    -   At N = 0, ration = R (no reduction when alone)
    -   At N = K, ration = R/2
-   Separate k allowed per patch to reflect differences in habitat productivity
-   Asymptotes to 0 at very high density

**Choice of K should depend on range of possible densities... K is the sensitive parameter and will substantially affect results.**

```{r}
fncDensityRation <- function(fish_pop, R_warm, R_cold, k_warm = 50, k_cold = 50) {
  fish_pop %>%
    mutate(
      R_base = if_else(patch == "warm", R_warm, R_cold),
      k      = if_else(patch == "warm", k_warm, k_cold),
      ration = R_base * k / (k + density)
    ) %>%
    select(-R_base, -k)
}
```

Visualize how ration scales with density under different values of k to help calibrate the parameter. *Both patches have the same baseline ration, so curves are identical*

```{r}
# Use actual R_warm and R_cold from the session as the baseline
expand.grid(
  density = 1:200,
  k       = c(10, 25, 50, 100, 200),
  patch   = c("warm", "cold")
) %>%
  mutate(
    R_base = if_else(patch == "warm", R_warm, R_cold),
    ration = R_base * k / (k + density),
    k_lab  = factor(paste0("k = ", k), levels = paste0("k = ", sort(unique(k))))
  ) %>%
  ggplot(aes(x = density, y = ration, color = k_lab)) +
  geom_line() +
  geom_hline(aes(yintercept = R_base), linetype = "dashed", color = "grey50") +
  facet_wrap(~ patch, labeller = label_both) +
  labs(
    x     = "Fish density (N per patch)",
    y     = "Realized ration (proportion of Cmax)",
    color = "Half-saturation\ndensity (k)",
    title = "Density-dependent ration reduction",
    caption = "Dashed line = baseline ration (no density effect)"
  )
```

-   k = 10–25: Severe food limitation even at low densities (ration halved by \~10–25 fish). Probably too strong unless you're modeling highly food-limited systems.
-   k = 50–100: Moderate competition; ration is meaningfully reduced at realistic densities (50–150 fish/patch), which is likely the right range to start with given your n_fish_per_strat setup.
-   k = 200: Very mild — food limitation only kicks in at very high densities.

### Linear dilution

Here is a simpler formulation, where ration size is distributed evenly across all individuals (pure scramble competition) and food supply within a patch is fixed. I.e., ration size of 0.1 g is evenly distributed among all fish in a patch. If we use this, it would probably make sense to define sub-patches (feeding stations) within the larger patches.

```{r}
fncDensityRation_ld <- function(fish_pop, R_warm, R_cold) {
  fish_pop %>%
    mutate(food   = if_else(patch == "warm", R_warm, R_cold),
           ration = food / density
           ) %>%
    select(-food)
}


# set up objects as they would appear in the simulation loop
R_warm      <- get_patchration(1, "warm")
R_cold      <- get_patchration(1, "cold")
fish_pop <- bind_rows(
  tibble(strategy = "resident_warm", 
         patch = "warm",
         weight = pmax(1, rnorm(100, 0.1, 0)),
         density = c(1:100)),
  tibble(strategy = "resident_cold", 
         patch = "cold",
         weight = pmax(1, rnorm(100, 0.1, 0)),
         density = c(1:100))) %>%
  mutate(pid = row_number(),
         ration = if_else(patch == "warm", ration_warm, ration_cold))
# fish_pop

# for example, competition more intense in warm habitat
fish_pop <- fncDensityRation_ld(fish_pop, R_warm, R_cold)
fish_pop %>%
  ggplot(aes(x = density, y = ration, color = patch)) +
  geom_line() +
  scale_color_manual(values = c("warm" = 2, "cold" = 4)) +
  theme_bw() +
  xlab("Fish density") +
  ylab("Ration")
```

## Survival

Fish experience mortality based on their size (cumulative experience) and instantaneous growth rate (measure of current conditions). Smaller fish with lower growth rates more likely to die (sampled probabilistically).

**Notes:**

-   Fullerton's original code for the survival function also accounted for recent consumption (p-values: proportion of max consumption) in addition to size and recent growth. But the IBM simulation code has p-values fixed at 1, and the supporting papers state that survival is only based on fish weight and recent growth. Thus, here I remove any consumption effects to maintain consistency with published work
-   Note that effect of growth rate is scaled to the growth rates experienced by all fish, but only in that time step. So even if all fish are experiencing positive growth rates, survival probability will be less than s-min for fish growing the slowest. Likewise, even if all fish are experiencing negative growth, many will still live.

```{r}
fncSurvive <- function(df, minprob = 0.997){
  # df: data frame of fish table with only the survivors, e.g., fish[fish$survive == 1, c("weight", "growth")]
  # minprob: smallest probability any fish can have of dying in any time step
  
  # Weights
  w <- df$weight # weight of fish that are alive at this time step
  
  # Function: 
  b <- 1 # beta: b=1 has no effect, but could change the shape of the curve with different values if desired
  v <- minprob + (1 - minprob) * (1 - 1 / exp(b * w))
  
  # Growth during this time step that reflects recent conditions (i.e., a hungry/stressed fish may behave in ways that make it more vulnerable to predation, etc.)
  g <- df$growth 
  g <- fncRescale(g, to = c(-0.001, 0.001))
  
  # Amount of food eaten in previous time step (better survival if bellies full b/c hunkered down somewhere)
  # f <- df$pvals
  # f <- fncRescale(f, to = c(-0.001, 0.001))
  
  # Probability of survival
  prb.srv <- v + g #+ f
  prb.srv[prb.srv > 1] <- 1 # set upper bound at 1
  
  # Sample from binomial distribution with probabilities of prb.srv to determine which fish survive this time step
  survivors <- rbinom(n = nrow(df), size = 1, prob = prb.srv)
  
  return(list(prb.srv, survivors))
  #return(survivors)
}
```

As an example, plot daily survival probabilities based on fish size and instantaneous growth

```{r}
df <- expand_grid(
  weight = seq(from = 0, to = 100, by = 0.1),
  growth = seq(from = -0.03, to = 0.03, by = 0.005)#,
  #pvals  = seq(from = 0, to = 1, by = 0.1)
)

surv.list <- fncSurvive(df)
df <- df %>% mutate(prsurv = surv.list[[1]],
                    survivors = surv.list[[2]])

df %>% #filter(pvals == 0.1) %>%
  ggplot() +
  geom_line(aes(x = weight, y = prsurv, color = as_factor(growth))) +
  xlim(0,10) + theme_bw() +
  xlab("Fish weight (g)") + ylab("Survival probability") + labs(color = "Instantaneous\ngrowth (g/g/d)")
```

## Pre-calculate growth

Set parameters for *O. mykiss*

```{r}
spp <- "stealhead"
numFish <- 500 #number of fish in a network per scenario
Constants <- fncReadConstants()
st.wt <- 100 #starting weight
ration.lo <- 0.02
ration.hi <- ration.lo + 0.05
pred.en.dens <- 5900 #predator energy density
oxy <- 13560 #oxygen consumed
pff <- 0.1 #percent indigestible prey
prey.en.dens <- 3500 # see Railsback and Rose 1999, Beauchamp et al. 2007

  
# Set up arrays
waterTemps <- cbind("pid" = seq(1, length(seq(0.1, 25, 0.1))),"WT" = seq(0.1, 25, 0.1))
rations <- seq(0.02, 0.17, 0.001)
weights <- seq(1, 1500, 1)
wt.growth <- array(NA, dim = c(nrow(waterTemps), length(rations), length(weights))) #dim = c(250 WTs, 16 rations, 1500 weights)
p.values <- array(NA, dim = c(nrow(waterTemps), length(rations), length(weights)))
#respiration <- consumption <- wastes <- wt.growth

# Loop through arrays to record growth for each case
for(w in 1:dim(wt.growth)[3]){
  ww <- weights[w]
  for(r in 1:dim(wt.growth)[2]){
    rr <- rations[r]
    Input<- fncGetBioEParms("stealhead", pred.en.dens, prey.en.dens, oxy, pff,
                            waterTemps, 
                            startweights = rep(ww, nrow(waterTemps)), 
                            pvals = rep(1, nrow(waterTemps)),
                            ration = rep(rr, nrow(waterTemps)))
    
    
    Results <- BioE(Input, Constants) #run BioE code given input and constants
    wt.growth[,r,w] <- c(Results$Gg_WinBioE) # growth in g/g/d...growth standardized by fish weight
    p.values[,r,w] <- c(Results$pp)
    # consumption[,r,w] <- c(Results$Consumption)
    # wastes[,r,w] <- c(Results$Egestion + Results$Excretion + Results$S.resp)
    # respiration[,r,w] <- c(Results$Respiration)
    
    ## In Joules instead of grams:
    # wt.growth[,r,w] <- c(Results$Growth_j)
    # consumption[,r,w] <- c(Results$Consumption_j)
    # wastes[,r,w] <- c(Results$Egestion_j + Results$Excretion_j + Results$Sj.resp)
    # respiration[,r,w] <- c(Results$Respiration_j)
  }
  # print(w)
}

# Save outputs. Only 'wt.growth' is needed for simulations. The others are merely for plotting below.
# save("wt.growth", file = "data.in/wt.growth.array.RData")
# save("respiration", file = "data.in/respiration.array.RData")
# save("wastes", file = "data.in/wastes.array.RData")
# save("consumption", file = "data.in/consumption.array.RData")



# Load back in to plot
# load("data.in/wt.growth.array.RData")
# load("data.in/consumption.array.RData")
# load("data.in/wastes.array.RData")
# load("data.in/respiration.array.RData")
# fncGetBioEParms("stealhead", pred.en.dens, prey.en.dens, oxy, pff,
#                             waterTemps, 
#                             startweights = rep(ww, nrow(waterTemps)), 
#                             pvals = rep(1, nrow(waterTemps)),
#                             ration = rep(rr, nrow(waterTemps)))

```

### Visualize bioenergetics model output

Plot growth rate by temperature for fish of varying masses, variable ration sizes.

```{r fig.width=7, fig.height=6}
par(mfrow = c(2, 3), las = 1)
ylb <- "Growth rate (g/g/d)"
for(fish.mass in c(1, 10, 100, 250, 500, 1500)){
  plot(waterTemps[,"WT"], wt.growth[, 151, fish.mass], type='l', ylab = ylb, xlab = "Water temperature (C)", 
       col = 4, ylim = c(-0.04, 0.04), main = paste0("fish weight = ", fish.mass," g")) #max ration
  for(i in seq(2, length(rations)-1, by = 10)){ lines(waterTemps[,"WT"], wt.growth[,i, fish.mass], col = "gray70")}
  abline(h = 0, lty = 2)
  lines(waterTemps[,"WT"], wt.growth[, 1, fish.mass], col = 2, lwd = 1.5) # min ration
  lines(waterTemps[,"WT"], wt.growth[, 151, fish.mass], col = 4, lwd = 1.5) # max ration
}
```

Plot growth potential by weight under different temperatures (panels) and ration sizes (line color: red = high, blue = low).

```{r fig.width=7, fig.height=6}
par(mfrow = c(2, 3), las = 1)
# speccols <- hcl.colors(13, "Temps")
# ratiseq <- seq(1, length(rations), by = 30)
# tempseq <- seq(1, length(waterTemps[,"WT"]), by = 20)
# 
# for (r in 1:length(ratiseq)) {
#   
#   plot(wt.growth[1,ratiseq[r],] ~ weights, type = "n", ylim = c(-0.04, 0.04),
#        ylab = "Growth rate (g/g/d)", xlab = "Weight (g)", 
#        main = paste0("ration = ", rations[ratiseq[r]]," g"))
#   
#   for (i in 1:length(tempseq)) { 
#     lines(wt.growth[tempseq[i],ratiseq[r],] ~ weights, col = speccols[i])
#     # print(range(wt.growth[i,ratiseq[r],]))
#   }
# }


speccols <- hcl.colors(11, "Temps")
ratiseq <- seq(1, length(rations), by = 15)
tempseq <- seq(1, length(waterTemps[,"WT"]), by = 45)

for (i in 1:length(tempseq)) { 
  
  plot(wt.growth[tempseq[i],1,] ~ weights, type = "n", ylim = c(-0.04, 0.04),
       ylab = "Growth rate (g/g/d)", xlab = "Weight (g)", 
       main = paste0("temp = ", waterTemps[tempseq[i], "WT"]," C"))
  
  for (r in 1:length(ratiseq)) {
    lines(wt.growth[tempseq[i],ratiseq[r],] ~ weights, col = speccols[r])
    # print(range(wt.growth[i,ratiseq[r],]))
  }
  }
```

Plot optimal growth temperature and maximum growth rate by fish size: blue = maximum ration, grey = (nearly) minimum ration

```{r}
# dim(wt.growth)[3]
opttemptib <- tibble(weight = weights, 
                     opttemp_ratlo = NA, opttemp_rathi = NA, 
                     growth_ggd_ratlo = NA, growth_ggd_rathi = NA)

for (i in 1:dim(wt.growth)[3]) {
  opttemptib$opttemp_ratlo[i] <- waterTemps[which.max(wt.growth[,10,i]),2]
  opttemptib$opttemp_rathi[i] <- waterTemps[which.max(wt.growth[,151,i]),2]
  opttemptib$growth_ggd_ratlo[i] <- wt.growth[which.max(wt.growth[,10,i]),10,i]
  opttemptib$growth_ggd_rathi[i] <- wt.growth[which.max(wt.growth[,151,i]),151,i]
}

opttemptib %>% ggplot() +
  geom_line(aes(x = weight, y = opttemp_ratlo), color = 2) +
  geom_line(aes(x = weight, y = opttemp_rathi), color = 4) +
  theme_bw() + xlab("Fish weight (g)") + ylab("Optimal temperature for growth (C)")

opttemptib %>% ggplot() +
  geom_line(aes(x = weight, y = growth_ggd_ratlo), color = 2) +
  geom_line(aes(x = weight, y = growth_ggd_rathi), color = 4) +
  theme_bw() + xlab("Fish weight (g)") + ylab("Maximum growth rate (g/g/d)")
```

## Simple growth regime

Derive a generic growth regime for a 10 gram fish, sensu Armstrong et al. (2021). *Note: for simplicity, I did not simulate changing growth potential due to increase in fish mass over the course of the year, as in the NCC paper.*

In these \~simple scenarios, growth potential for the "optimal mover" is just calculated from the maximum of that across the two habitats. No need to simulate habitat selection/movement.

Get growth potential by temperature for a 10 gram fish, fed at \~max ration

```{r}
# get relative growth rates for a 10 g fish across range of water temps
grate_df <- tibble(temp = waterTemps[,2],
                   grate_ggd = wt.growth[,151,10]) %>%
  mutate(temp = trimws(temp))

# calculate growth potential (g/d/d) and cumulative growth (g) across warm and cold patches and for an "optimal mover"
growthregime_df <- habitat_df %>%
  mutate(temp_warm = trimws(temp_warm),
         temp_cold = trimws(temp_cold)) %>%
  left_join(grate_df, by = c("temp_warm" = "temp")) %>% rename(grpot_warm = grate_ggd) %>%
  left_join(grate_df, by = c("temp_cold" = "temp")) %>% rename(grpot_cold = grate_ggd) %>%
  mutate(grpot_track = pmax(grpot_warm, grpot_cold)) %>%
  mutate(cumul_growth_warm = cumsum(grpot_warm*10),
         cumul_growth_cold = cumsum(grpot_cold*10),
         cumul_growth_track = cumsum(grpot_track*10),
         habitat = case_when(
           grpot_cold > grpot_warm ~ "cold",
           grpot_warm > grpot_cold ~ "warm",
           TRUE ~ "tie"
         ))
  
head(growthregime_df)
```

Plot growth regimes (growth potential and g/g/d)

```{r}
p1 <- growthregime_df %>% ggplot() +
  geom_line(aes(x = date, y = grpot_warm), col = 2) +
  geom_line(aes(x = date, y = grpot_cold), col = 4) +
  theme_bw() +
  xlab("Date") +
  ylab("Growth potential (g/g/day)") +
  labs(title = "Warm vs. cold habitats")

p2 <- growthregime_df %>% ggplot() +
  geom_line(aes(x = date, y = grpot_track), col = 1, linetype = "dashed") +
  theme_bw() +
  xlab("Date") +
  ylab("Growth potential (g/g/day)") +
  labs(title = "Optimal mover")

ggarrange(p1, p2, ncol = 2)
```

Plot cumulative growth

```{r}
growthregime_df %>% ggplot() +
  geom_line(aes(x = date, y = cumul_growth_warm, color = "Warm resident")) +
  geom_line(aes(x = date, y = cumul_growth_cold, color = "Cold resident")) +
  geom_line(aes(x = date, y = cumul_growth_track, color = "Optimal mover"), linetype = "dashed") +
  scale_color_manual(values = c("Warm resident" = 2, "Cold resident" = 4, "Optimal mover" = 1)) +
  theme_bw() +
  xlab("Date") +
  ylab("Cumulative growth (g)")
```

Plot habitat use, i.e., temporal change in the habitat where growth potential is maximized.

```{r}
growthregime_df %>% 
  mutate(habitat_num = as.numeric(as.factor(habitat))) %>%
  ggplot() +
  geom_line(aes(x = date, y = habitat_num)) +
  geom_point(aes(x = date, y = habitat_num, color = habitat)) +
  theme_bw() +
  scale_color_manual(values = c("cold" = 4, "warm" = 2)) +
  xlab("Date") +
  ylab("Habitat with max. GP (cold = 1, warm = 2)")
```

Contributions of warm vs. cold patches

```{r}
# time
growthregime_df %>% 
  group_by(habitat) %>%
  summarize(ndays = n()) %>%
  mutate(habitat = factor(habitat)) %>%
  ggplot() +
  geom_bar(aes(x = habitat, y = ndays/dim(growthregime_df)[1], fill = habitat), stat = "identity") +
  theme_bw() + theme(legend.position = "none") +
  scale_fill_manual(values = c("cold" = 4, "warm" = 2)) +
  ylab("Proportion of time spent in cold vs. warm habitats")

# cumulative growth accrued
growthregime_df %>% 
  group_by(habitat) %>%
  summarize(cumul_grpot = sum(grpot_track*10)) %>%
  mutate(habitat = factor(habitat)) %>%
  ggplot() +
  geom_bar(aes(x = habitat, y = cumul_grpot, fill = habitat), stat = "identity") +
  theme_bw() + theme(legend.position = "none") +
  scale_fill_manual(values = c("cold" = 4, "warm" = 2)) +
  ylab("Proportion of growth accrued in cold vs. warm habitats")
```

## Individual-Based Model

***Configuration options***

-   **Stochasticity in movement** is accounted for by one of three alternative mechanisms:
    -   *No stochastic movement/entirely deterministic:* all movers do the same thing...they all can perfectly sense and compare growth opportunities
    -   *Probabilistic (softmax) rule:* the softmax function assigns probabilities to each patch based on the growth difference, with higher probabilities for patches offering greater growth benefits.
    -   *Individual-level movement threshold:* each fish has a unique threshold for movement, representing individual variation in boldness/dispersal propensity
-   **Density-dependent food availaility:** after fish have selected patches, calculate food availability based on patch density and baseline ration levels. Two alternative approaches:
    -   *Modified Fullerton approach:* ration size decreases linearly with fish density, up to some (*a priori* defined) threshold above which there is no effect of there is no effect of density of ration size.
    -   *Hyperbolic reduction:* describes a non-linear decrease in ration size with increasing fish density, where the rate of decrease slows as density increases.

***Senstive parameters:***

-   General
    -   Habitat variables (temperature regimes and food availability in each patch)
    -   Number of fish per strategy/total number of starting fish
-   Density-dependence
    -   Fullerton upper density threshold
    -   Hyperbolic half saturation density
-   Movement stochasticity
    -   Probabilistic/softmax tau (sensitivity to growth difference)
    -   Individual-level movement threshold, SD in individual boldness

***Notes:***

-   No penalty for movement. Could implement a simple movement cost penalty (subtract constant), or scaled by fish size. Fullerton scales by travel distance, but this is not relevant for spatially implicit movement.
-   Currently, we just look-up growth from the precalculated growth array, but this is going to be less accurate for the IBM b/c precalculated growth rates are calculated over 1g mass increments, whereas growth in the IBM is continuous. Probably fine for now, but perhaps should borrow code from above/Fullerton IBM to increase accuracy.

### Parameters and initialization

Three movement strategies are simulated: fish that remain year-round in the warm patch (`resident_warm`), fish that remain year-round in the cold patch (`resident_cold`), and fish that choose the better-growing patch each day (`optimal_mover`). Ration sizes match the unequal-ration scenario above.

**Notes:**

-   All fish start at same size, no variation
-   variation in boldness parameter probably needs tuning
-   If using individual-level movement thresholds (i.e., boldness), residents can be given *move_threshold = Inf* to formally unify all three strategies under one mechanism, rather than branching by strategy label — keeps the code cleaner if we expand to more strategies later.

```{r}
# Configuration options:
move_stochastic <- "prob" # individual movement stochasticity: options = "none", "prob", "indiv"
food_densdepen <- "hyperbolic" # density dependence: options = "fullerton", "hyperbolic

# misc. parameters
n_fish_per_strat <- 50     # fish per strategy
start_wt         <- 10     # initial weight (g)
MaxDensity4Growth <- 50    # growth is not depressed further above this density (currently this is arbitary)
move_cost        <- 0.005  # g/g/d subtracted from the alternative patch's growth rate before
                           # comparing — fish switch only when penalised gain > staying gain
sigma_bold       <- 0.002  # SD of boldness distribution, for dispersal propensity (g/g/d units)
tau              <- 0.001  # sensitivity to growth difference (for probabilistic/softmax variation in movement)
K_warm <- K_cold <- 500    # For hyperbolic reduction, strength of density dependence/half-saturation density: fish density at which ration is halved

# Fullerton has movement cost coded slightly differently, where movement cost is a function of the instantaneous growth rate, distance moved, and a constant...rather than simply a fixed cost as is the case here. But because we only have two patches that are spatially implicit, we can ignore this more complex parameterization...at least for now.

# ration_warm = 0.10, ration_cold = 0.07 (defined in "Unequal ration" section above)

# Lookup sequences — identical to the axes of the wt.growth array
wt_seq_ibm <- waterTemps[, "WT"]  # seq(0.1, 25, 0.1), 250 values       
ra_seq_ibm <- rations             # seq(0.02, 0.17, 0.001), 151 values

# Pre-compute ration indices (constant throughout simulation)
# ra_idx_warm <- which.min(abs(ra_seq_ibm - ration_warm))
# ra_idx_cold <- which.min(abs(ra_seq_ibm - ration_cold))

# set.seed(7843)
# initialize fish
fish_pop <- bind_rows(
  tibble(strategy = "resident_warm", patch = "warm",
         weight = pmax(1, rnorm(n_fish_per_strat, start_wt, 0))),
  tibble(strategy = "resident_cold", patch = "cold",
         weight = pmax(1, rnorm(n_fish_per_strat, start_wt, 0))),
  tibble(strategy = "optimal_mover", patch = "warm",
         weight = pmax(1, rnorm(n_fish_per_strat, start_wt, 0)))) %>%
  mutate(pid = row_number(),
         ration = if_else(patch == "warm", ration_warm, ration_cold),
         # Individual movement threshold — drawn once, fixed for life
         # Positive = reluctant to move (needs a clear advantage)
         # Negative = bold/prone to move (moves even when slightly disadvantaged)
         move_threshold = c(rep(Inf, n_fish_per_strat*2), # residents never move
                            rnorm(n_fish_per_strat, mean = 0, sd = sigma_bold)),
         prob_surv = 1, survive = 1) %>%
  group_by(patch) %>% mutate(density = n()) %>% ungroup()
fish_pop
```

### Parameter tuning

#### Individual movement threshold

Plot the distribution of move_threshold values against the distribution of daily g_move - g_stay values to check that sigma_bold creates meaningful variation rather than all fish behaving identically or randomly.

```{r fig.width=7, fig.height=6}
# ── 1. Daily g_move − g_stay across the season ─────────────────────────────────
# Computed at representative masses (start, mid, end-of-season)
rep_weights <- c(10, 50, 150)

gdiff_df <- map_dfr(rep_weights, function(w) {
  ma_idx <- pmax(1L, pmin(4500L, round(w)))
  map_dfr(seq_len(nrow(habitat_df)), function(d) {
    t           <- habitat_df$doy[d]
    T_warm      <- get_patchtemp(t, "warm")
    T_cold      <- get_patchtemp(t, "cold")
    R_warm      <- get_patchration(t, "warm")
    R_cold      <- get_patchration(t, "cold")
    wt_idx_warm <- which.min(abs(wt_seq_ibm - T_warm))
    wt_idx_cold <- which.min(abs(wt_seq_ibm - T_cold))
    ra_idx_warm <- which.min(abs(ra_seq_ibm - R_warm))
    ra_idx_cold <- which.min(abs(ra_seq_ibm - R_cold))
    g_warm <- wt.growth[wt_idx_warm, ra_idx_warm, ma_idx]
    g_cold <- wt.growth[wt_idx_cold, ra_idx_cold, ma_idx]
    tibble(date = habitat_df$date[d], g_diff = g_warm - g_cold, weight = w)
  })
}) |>
  mutate(weight_lab = paste0(weight, " g fish"))

# ── 2. move_threshold distribution for several sigma_bold candidates ────────────
sigma_candidates <- c(0.001, 0.003, 0.005, 0.01)
thresh_df <- map_dfr(sigma_candidates, function(s) {
  tibble(
    move_threshold = rnorm(5000, mean = 0, sd = s),
    sigma_lab = paste0("σ = ", s)
  )
}) |>
  mutate(sigma_lab = factor(sigma_lab, levels = paste0("σ = ", sigma_candidates)))

# ── 3. Plot ────────────────────────────────────────────────────────────────────
# Panel A: g_diff over time
p_time <- ggplot(gdiff_df, aes(x = date, y = g_diff, color = factor(weight))) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Date", y = "g_warm − g_cold (g/g/d)",
       color = "Fish weight (g)", title = "A: Daily growth advantage of warm over cold patch")

# Panel B: distribution of g_diff vs move_threshold candidates
p_dist <- ggplot() +
  # Shaded density of g_diff (pooled across weights — the "signal")
  geom_density(data = gdiff_df, aes(x = g_diff, fill = "Patch growth\nadvantage"),
               alpha = 0.3, color = NA) +
  # move_threshold distributions for each sigma
  geom_density(data = thresh_df, aes(x = move_threshold, color = sigma_lab),
               linewidth = 0.85) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = "steelblue") +
  scale_color_brewer(palette = "Reds", direction = 1) +
  labs(x = "g/g/d", y = "Density",
       fill = NULL, color = "Threshold SD (σ)",
       title = "B: Threshold distribution vs. actual growth advantage")

cowplot::plot_grid(p_time, p_dist, ncol = 1, rel_heights = c(1, 1.1))
```

Panel A shows the growth signal movers are responding to: g_warm - g_cold is small and near zero most of the year (\~0 to +0.005), with a large negative dip in summer (−0.025) when cold becomes strongly advantageous. Fish weight has minimal effect on the signal.

Panel B shows how the threshold distributions compare to that signal — here's what it means for behavior:

-   σ = 0.001: Threshold much narrower than the signal. Most fish behave nearly deterministically — modest individual variation, mainly near the switching crossover in spring/fall
-   σ = 0.003: Threshold width similar to the low-signal periods (winter/shoulder seasons). Good spread of individual responses when the patch advantage is marginal; still mostly consistent in summer
-   σ = 0.005: Threshold wider than the signal for most of the year — choices become largely random outside the summer peak
-   σ = 0.01: Nearly random throughout; boldness is essentially noise

*σ = 0.001–0.003 is the defensible range — it creates genuine individual variation in behavior during the biologically interesting transitional periods (when movers are near indifferent between patches), while still letting the signal dominate during the strong summer cold-patch advantage. A reasonable starting point is sigma_bold = 0.002*

#### Probabilistic (softmax)

Plot the probability of choosing the warm patch given different values of tau, across a range of between-patch differences in growth potential.

```{r}
# Softmax: P(warm) = exp(g_warm/tau) / (exp(g_warm/tau) + exp(g_cold/tau))
# Equivalently: sigmoid of (g_warm - g_cold) / tau
tau_vals <- c(0.001, 0.003, 0.005, 0.01, 0.025)

g_diff_seq <- seq(-0.03, 0.03, length.out = 400)

softmax_df <- map_dfr(tau_vals, function(tau) {
  tibble(
    g_diff = g_diff_seq,
    p_warm = 1 / (1 + exp(-g_diff / tau)),
    tau_lab = paste0("τ = ", tau)
  )
}) |>
  mutate(tau_lab = factor(tau_lab, levels = paste0("τ = ", tau_vals)))

# Observed g_diff range from the earlier gdiff_df (pooled across weights)
obs_range <- range(gdiff_df$g_diff)

ggplot(softmax_df, aes(x = g_diff, y = p_warm, color = tau_lab)) +
  # Shade the observed g_diff range from the simulation
  annotate("rect",
           xmin = obs_range[1], xmax = obs_range[2],
           ymin = -Inf, ymax = Inf,
           fill = "grey85", alpha = 0.6) +
  annotate("text", x = mean(obs_range), y = 0.9,
           label = "Observed\ng_diff range", size = 3, color = "grey40",
           hjust = 0.5) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = 0,   linetype = "dashed", color = "grey50") +
  geom_line(linewidth = 0.9) +
  scale_color_brewer(palette = "RdYlBu", direction = -1) +
  scale_y_continuous(labels = scales::percent_format(),
                     breaks = seq(0, 1, 0.25)) +
  scale_x_continuous(labels = scales::label_number(accuracy = 0.01)) +
  annotate("text", x =  0.028, y = 0.58, label = "warm\nbetter", size = 3,
           color = "grey40", hjust = 1) +
  annotate("text", x = -0.028, y = 0.58, label = "cold\nbetter", size = 3,
           color = "grey40", hjust = 0) +
  labs(
    x     = "g_warm − g_cold  (g/g/d)",
    y     = "P(choose warm patch)",
    color = "Sensitivity (τ)",
    title = "Softmax patch selection probability",
    subtitle = "Grey band = observed growth difference range in simulation"
  ) +
  theme_bw()
```

-   tau = 0.001: Near step-function — almost deterministic. Fish choose the better patch \>99% of the time when \|g_diff\| \> \~0.005
-   tau = 0.003: Mostly deterministic at the summer peak (g_diff ≈ −0.025), but probabilistic near the zero-crossing — similar in effect to the perceptual noise approach
-   tau = 0.005: Substantial uncertainty even at moderate growth differences
-   tau = 0.010-0.025: Gradual, shallow curves — even a strong cold-patch advantage (\~0.025) only shifts P(warm) to \~25–30%; choices are largely stochastic throughout

*For this simulation (given the range of observed growth differences), τ = 0.001–0.003 is the meaningful range — it preserves the deterministic signal in summer while allowing individual probabilistic variation near the switching crossover. τ much above 0.005 makes the patch signal mostly invisible to the decision rule, i.e., patch selection becomes increasingly random.*

### Simulation loop

For each day of year, optimal movers compare the bioenergetics growth rate at both patches — the vectorised equivalent of calling `fncGrowthPossible()` for each fish — and switch only when the movement-cost-penalised gain in the alternative patch exceeds staying (but this may differ among individuals depending on how movement stochasticity is configured). Warm and cold patch residents stay in their pre-defined patch. `fncGrowthFish()` then looks up the actual growth rate for every fish at their current patch, weight, and ration, and weights are updated accordingly.

```{r}
n_days  <- nrow(habitat_df) # number of days in the simulation
n_fish  <- nrow(fish_pop)   # number of fish in the simulation
fish_pop_init <- fish_pop   # snapshot of full population before any mortality

# Pre-allocate storage (rows = fish pid, columns = days)
# Rows are indexed by fish pid (1:n_fish), which is stable even as fish die.
# Dead fish slots remain NA for all days after death.
ggd_matrix     <- matrix(NA_real_,      nrow = n_fish, ncol = n_days)  # standardized growth rates experienced
gd_matrix      <- matrix(NA_real_,      nrow = n_fish, ncol = n_days)  # absolute growth rates
temp_matrix    <- matrix(NA_real_,      nrow = n_fish, ncol = n_days)  # water temps experienced 
patch_matrix   <- matrix(NA_character_, nrow = n_fish, ncol = n_days)  # patch occupied 
mass_matrix    <- matrix(NA_real_,      nrow = n_fish, ncol = n_days)  # weight/mass trajectories
surv_matrix    <- matrix(NA_integer_,   nrow = n_fish, ncol = n_days)  # survived (1) or died (0) each day
switches       <- integer(n_fish)                                       # count patch switches per fish (indexed by pid)


for (d in 1:n_days) {
  t <- habitat_df$doy[d] # day of year
      
  # 1. GET/FIND PATCH TEMPERATURE AND RATION -----------------------------------------------------
  
  # Temperature indices for this day — shared by all fish
  T_warm      <- get_patchtemp(t, "warm")             # get warm patch temp on day t
  T_cold      <- get_patchtemp(t, "cold")             # get cold patch temp on day t
  wt_idx_warm <- which.min(abs(wt_seq_ibm - T_warm))  # water temp index for warm patch on day t
  wt_idx_cold <- which.min(abs(wt_seq_ibm - T_cold))  # water temp index for cold patch on day t
  
  # Ration indices for this day - shared by all fish
  R_warm      <- get_patchration(t, "warm")
  R_cold      <- get_patchration(t, "cold")
  ra_idx_warm <- which.min(abs(ra_seq_ibm - R_warm))  # ration index for warm patch on day t
  ra_idx_cold <- which.min(abs(ra_seq_ibm - R_cold))  # ration index for cold patch on day t
  
  
  # 2. CHOOSE PATCH: OPTIMAL MOVERS ---------------------------------------------------------------

  # Mirrors fncGrowthPossible() logic, vectorized across all movers at once.
  # Patch choice senses baseline temperature and ration, but not density
  mover_rows <- which(fish_pop$strategy == "optimal_mover")  # get row indices for optimal movers
  
  if (length(mover_rows) > 0) {
    # Mass indices for each mover (clamped to array bounds)
    ma_idx_vec <- pmax(1L, pmin(4500L, round(fish_pop$weight[mover_rows]))) # get row index for current mass
    
    # Growth rate (g/g/d) at each patch for every mover (vectorized array lookup)
    g_warm_vec <- wt.growth[wt_idx_warm, ra_idx_warm, ma_idx_vec]  # growth in warm patch given current conditions
    g_cold_vec <- wt.growth[wt_idx_cold, ra_idx_cold, ma_idx_vec]  # growth in cold patch given current conditions
    
    # Calculate growth associated with staying in current patch vs. moving to alternate patch
    # If relevant, penalize the alternative patch by energetic cost of movement
    in_warm    <- fish_pop$patch[mover_rows] == "warm"  # is the fish in warm patch? T/F
    g_stay     <- ifelse(in_warm, g_warm_vec, g_cold_vec) # growth if staying in current patch
    g_move_net <- ifelse(in_warm, g_cold_vec, g_warm_vec) # growth if moving (unpenalized) 
    #g_move_net <- ifelse(in_warm, g_cold_vec - move_cost, g_warm_vec - move_cost) # growth if moving (penalized)
    prev_patch <- fish_pop$patch[mover_rows]
    
    # stochasticity in movement
    if (move_stochastic == "none") {
      ### 2a. Deterministic: All movers do the same thing
      # Update patch based on whether moving is beneficial. 
      # 1. If fish in is warm patch and growth potential of moving exceeds growth potential of staying, move to cold. 
      # 2. If fish in is cold patch and growth potential of moving exceeds growth potential of staying, move to warm. 
      fish_pop$patch[mover_rows] <- case_when(
        in_warm  & g_move_net > g_stay ~ "cold",
        !in_warm & g_move_net > g_stay ~ "warm", 
        TRUE                           ~ fish_pop$patch[mover_rows]
        )
    } else if (move_stochastic == "prob") {
      ### 2b. Probabilistic (softmax) rule: Instead of a hard threshold, fish switch with a probability that increases with the growth advantage. One parameter: τ (sensitivity; small = nearly deterministic, large = nearly random).
      p_warm <- exp(g_warm_vec / tau) / (exp(g_warm_vec / tau) + exp(g_cold_vec / tau)) # calculate probability of selecting warm habitat
      choose_warm <- runif(length(mover_rows)) < p_warm # choose warm if p_warm exceeds a random number between 0 and 1
      fish_pop$patch[mover_rows] <- if_else(choose_warm, "warm", "cold") # update patch
    } else if (move_stochastic == "indiv") {
      ### 2c. Individual-level movement threshold: each fish has a unique threshold for movement, representing individual variation in boldness/dispersal propensity
      fish_pop$patch[mover_rows] <- case_when(
      in_warm  & (g_move_net - g_stay) > fish_pop$move_threshold[mover_rows] ~ "cold",
      !in_warm & (g_move_net - g_stay) > fish_pop$move_threshold[mover_rows] ~ "warm",
      TRUE                                                                   ~ prev_patch) #fish_pop$patch[mover_rows])
    }
    
    # Tally switches (indexed by pid so counts remain correct as fish die)
    switched <- fish_pop$patch[mover_rows] != prev_patch
    switches[fish_pop$pid[mover_rows]] <- switches[fish_pop$pid[mover_rows]] + as.integer(switched)

  } # end patch choice
  
  # Update ration to reflect current patch
  fish_pop$ration <- if_else(fish_pop$patch == "warm", R_warm, R_cold)
  
  # 3. UPDATE HABITAT QUALITY (DENSITY-DEPENDENT RATION) -----------------------------------------

  # Update density after fish have moved
  fish_pop <- fish_pop %>% group_by(patch) %>% mutate(density = n()) %>% ungroup()
  
  if (food_densdepen == "fullerton") {
    ### 3a. Fullerton approach
    # get density effect on ration
    fdens <- fdens_raw <- fish_pop$density # get fish "density" per patch (here, simply abundance b/c space is not specified)
    fdens[fdens > MaxDensity4Growth] <- MaxDensity4Growth # high densities cap out at MaxDensity4Growth
    density.effect <- fncRescale((1 - c(fdens, 0.01, MaxDensity4Growth)), c((0.5),1)) # calculate density effect scalar
    density.effect <- density.effect[-c(length(density.effect), (length(density.effect) - 1))] # remove the last 2 temporary values
    # update ration in patches to account for density dependence in food availability
    fish_pop <- fish_pop %>% mutate(ration = ration * density.effect)
  } else if (food_densdepen == "hyperbolic") {
    ### 3b. Hyperbolic reduction approach
    fish_pop <- fncDensityRation(fish_pop = fish_pop, R_warm = R_warm, R_cold = R_cold, k_warm = K_warm, k_cold = K_cold)
  }
  
  
  # 4. GROW FISH (WISCONSIN BIOENERGETICS) -------------------------------------------------------
  
  # Growth lookup for all fish via fncGrowthFish. Growth in g/g/d
  growth_df <- fncGrowthFish(NA, as.data.frame(fish_pop), t)
  
  
  # 5. SURVIVAL BASED ON FISH WEIGHT AND GROWTH --------------------------------------------------
  
  surv.list <- fncSurvive(growth_df)
  growth_df <- growth_df %>% 
    mutate(prob_surv = surv.list[[1]], 
           survive   = surv.list[[2]])
  
  
  # 6. STORE RESULTS AND REMOVE NON-SURVIVORS ----------------------------------------------------
  
  # Store experienced rates before weight update.
  # All indexing uses fish pid (rows 1:n_fish) so slots remain stable as fish die.
  temp_matrix[fish_pop$pid, d]  <- growth_df$WT.actual
  ggd_matrix[fish_pop$pid, d]   <- growth_df$growth
  gd_matrix[fish_pop$pid, d]    <- growth_df$growth * fish_pop$weight
  
  # Update weight: W_{t+1} = W_t + (growth_rate * W_t)
  fish_pop$weight <- fish_pop$weight + (growth_df$growth * fish_pop$weight)
  
  # Store updated weight, patch, and survival outcome for this day
  mass_matrix[fish_pop$pid, d]  <- fish_pop$weight
  patch_matrix[fish_pop$pid, d] <- fish_pop$patch
  surv_matrix[fish_pop$pid, d]  <- growth_df$survive
  
  # Remove non-survivors — only living fish carry forward to next iteration
  fish_pop <- fish_pop[growth_df$survive == 1, ]
}
```

### General results

Summarize results

```{r}
# assign DOY column names
colnames(mass_matrix)    <- as.character(habitat_df$doy)
colnames(patch_matrix)   <- as.character(habitat_df$doy)
colnames(temp_matrix)    <- as.character(habitat_df$doy)
colnames(ggd_matrix)     <- as.character(habitat_df$doy)

# Convert weight matrix to long format, join with patch matrix
# Use fish_pop_init to label rows — fish_pop at this point only contains survivors
mass_long <- as_tibble(mass_matrix) %>%
  bind_cols(fish_pop_init %>% select(pid, strategy)) %>%
  pivot_longer(-c(pid, strategy), names_to = "doy", values_to = "weight") %>%
  mutate(doy = as.integer(doy)) %>%
  left_join(
    as_tibble(patch_matrix) %>%
      bind_cols(fish_pop_init %>% select(pid)) %>%
      pivot_longer(-pid, names_to = "doy", values_to = "patch") %>%
      mutate(doy = as.integer(doy)),
    by = c("pid", "doy")
  ) %>%
  left_join(habitat_df %>% select(doy, date), by = "doy")

# Convert temp matrix to long format
temp_long <- as_tibble(temp_matrix) %>%
  bind_cols(fish_pop_init %>% select(pid, strategy)) %>%
  pivot_longer(-c(pid, strategy), names_to = "doy", values_to = "temp") %>%
  mutate(doy = as.integer(doy))

# convert standardized growth rate matrix to long format
ggd_long <- as_tibble(ggd_matrix) %>%
  bind_cols(fish_pop_init %>% select(pid, strategy)) %>%
  pivot_longer(-c(pid, strategy), names_to = "doy", values_to = "ggd") %>%
  mutate(doy = as.integer(doy))

# convert survival matrix to long format (1 = survived, 0 = died, NA = already dead)
surv_long <- as_tibble(surv_matrix) %>%
  bind_cols(fish_pop_init %>% select(pid, strategy)) %>%
  pivot_longer(-c(pid, strategy), names_to = "doy", values_to = "survived") %>%
  mutate(doy = as.integer(doy))

# summarize by strategy
ibm_summary <- mass_long %>%
  left_join(temp_long, by = c("pid", "strategy", "doy")) %>%
  left_join(ggd_long,  by = c("pid", "strategy", "doy")) %>%
  left_join(surv_long, by = c("pid", "strategy", "doy")) %>%
  group_by(strategy, doy, date) %>%
  summarise(
    mean_weight = mean(weight, na.rm = TRUE),
    sd_weight   = sd(weight,   na.rm = TRUE),
    mean_temp   = mean(temp,   na.rm = TRUE),
    sd_temp     = sd(temp,     na.rm = TRUE),
    mean_ggd    = mean(ggd,    na.rm = TRUE),
    sd_ggd      = sd(ggd,      na.rm = TRUE),
    prop_warm   = mean(patch == "warm", na.rm = TRUE),
    n_alive     = sum(!is.na(weight)),   # fish with non-NA weight were alive this day
    .groups     = "drop"
  )

ibm_summary
```

#### Size trajectories

Weight trajectories by strategy (ribbon = ±1 SD across individuals).

*Uncertainty derives from variation in starting size (if allowed) and stochasticity in individual movement behavior (if configured).*

```{r}
ibm_summary %>%
  ggplot(aes(x = date, color = strategy, fill = strategy)) +
  geom_ribbon(aes(ymin = mean_weight - sd_weight, ymax = mean_weight + sd_weight),
              alpha = 0.15, color = NA) +
  geom_line(aes(y = mean_weight)) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  scale_fill_manual(values  = c(resident_warm = "red", 
                                resident_cold = "blue",
                                optimal_mover = "darkgreen")) +
  theme_bw() + #ylim(0,10000) +
  xlab("Date") + ylab("Mean weight ± SD (g)")
```

#### Survival

Plot daily counts of survivors and overall survival by strategy (plots are the same, survival is standardized by number of initial fish per strategy)

```{r}
p1 <- ibm_summary %>% 
  ggplot(aes(x = date, y = n_alive, color = strategy)) +
  geom_line() +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(
    x     = "Date",
    y     = "Number alive",
    color = "Strategy",
    title = "Population size by strategy"
  ) +
  theme_bw()

p2 <- ibm_summary %>% 
  ggplot(aes(x = date, y = n_alive / n_fish_per_strat, color = strategy)) +
  geom_line() +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(
    x     = "Date",
    y     = "Proportion alive",
    color = "Strategy",
    title = "Cumulative survival by strategy"
  ) +
  theme_bw()

ggarrange(p1, p2, common.legend = TRUE)
```

Summarize survival outcomes by strategy:

```{r}
ibm_summary %>% 
  filter(doy == n_days) %>%
  mutate(survive_rate = n_alive / n_fish_per_strat) %>%
  select(strategy, n_alive, survive_rate)
```

-   *resident_warm* has the lowest overall survival, consistent with it being the poorest-growing strategy.
    -   mid-year acceleration in *resident_warm* mortality coincides with slower/negative growth in mid-summer, which is exactly what the function should be responding to.
    -   similarly, late-year acceleration in mortality coincides with negative growth at this time
-   *resident_cold* has moderate overall survival (88%). Survival declines the early part of the year when growth is negative and fish are small. No mortality occurs during the summer, consistent with theory on the role of perennially cold habitat providing summer refugia.
-   *optimal_mover* has the highest overall survival (\~96%), reflecting its growth advantage keeping individual survival probabilities near 1.0 most of the year.

Check whether mortality is concentrated among the smallest or slowest-growing individuals within each strategy — e.g., plot final weight of survivors vs. non-survivors — to confirm size-selective mortality is operating as expected.

```{r}
n_days <- ncol(mass_matrix)

# Per-fish fate summary
fish_fate <- fish_pop_init %>%
  mutate(
    # Survived to end of year if final day weight is non-NA
    survived     = if_else(!is.na(mass_matrix[pid, n_days]), "Survived", "Died"),
    # Last recorded weight (weight at death for non-survivors, final weight for survivors)
    wt_last      = apply(mass_matrix, 1, function(x) tail(x[!is.na(x)], 1)),
    # Mean daily growth rate across all days alive
    mean_ggd     = rowMeans(ggd_matrix, na.rm = TRUE),
    # Days alive
    days_alive   = rowSums(!is.na(mass_matrix))
  )

fish_fate %>% count(strategy, survived)
```

```{r fig.height=7, fig.width=5}
strategy_order <- c("resident_warm", "resident_cold", "optimal_mover")

fd <- fish_fate %>% mutate(strategy = factor(strategy, levels = strategy_order))

p_wt <- ggplot(fd, aes(x = survived, y = wt_last)) +
  geom_boxplot(width = 0.5, outlier.shape = NA, fill = "grey") +
  geom_jitter(width = 0.15, size = 1.5, alpha = 0.6) +
  facet_wrap(~ strategy, nrow = 1) +
  #scale_fill_manual(values = c(Died = "tomato", Survived = "steelblue")) +
  labs(x = NULL, y = "Weight (g)",
       title = "A: Weight at last observation") +
  theme_bw() + theme(legend.position = "none")

p_ggd <- ggplot(fd, aes(x = survived, y = mean_ggd)) +
  geom_boxplot(width = 0.5, outlier.shape = NA, fill = "grey") +
  geom_jitter(width = 0.15, size = 1.5, alpha = 0.6) +
  facet_wrap(~ strategy, nrow = 1) +
  #scale_fill_manual(values = c(Died = "tomato", Survived = "steelblue")) +
  labs(x = NULL, y = "Mean growth rate (g/g/d)",
       title = "B: Mean daily growth rate across days alive") +
  theme_bw() + theme(legend.position = "none")

p_days <- ggplot(fd |> filter(survived == "Died"),
                 aes(x = strategy, y = days_alive, color = strategy)) +
  geom_jitter(width = 0.15, size = 2.5, alpha = 0.8) +
  scale_x_discrete(limits = strategy_order) +
  labs(x = NULL, y = "Day of death (day of year)",
       title = "C: When did non-survivors die?") +
  theme_bw() + theme(legend.position = "none") +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen"))

egg::ggarrange(p_wt, p_ggd, p_days, ncol = 1)
```

Yes, overall, the mortality function in behaving correctly: it's selecting against small, slow-growing fish in a biologically interpretable way.

-   Panel A — fish that died were consistently lighter at last observation across all strategies. The separation is sharpest for *optimal_mover* (tiny died fish vs. large survivors) and *resident_cold*, and more compressed for *resident_warm* where all fish are similarly small.
-   Panel B — The growth rate signal is the clearest confirmation. Non-survivors in *resident_cold* and *optimal_mover* had negative mean growth rates — they were actively losing mass before dying, which is exactly what you'd expect the survival function to respond to. *resident_warm* non-survivors overlap more with survivors, which makes sense: when the whole strategy is slow-growing, survival probabilities don't vary as much among individuals, so mortality there is more stochastic.
-   Panel C — Timing of deaths also makes biological sense:
    -   *resident_cold* deaths cluster early (\~day 55–70)...the smallest/slowest fish that can't sustain themselves even in the cold patch
    -   *resident_warm* deaths spread through summer/fall as the warm patch becomes suboptimal
    -   *optimal_mover* deaths also cluster early...a few unlucky fish

#### Patch occupancy

Patch occupancy for optimal movers — proportion in warm patch through the year.

-   If move_stochastic = "prob", movers only really start to pile into the warm habitat during the shoulder seasons when benefits of warm habitat far exceed the cold habitat.
-   If move_stochastic = "Indiv", ...

```{r}
ibm_summary %>%
  filter(strategy == "optimal_mover") %>%
  ggplot(aes(x = date, y = prop_warm)) +
  geom_line() +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray50") +
  theme_bw() +
  xlab("Date") + ylab("Proportion of optimal movers in warm patch")
```

#### Experienced temperature

Temperature experienced by optimal movers

```{r}
ibm_summary %>%
  filter(strategy == "optimal_mover") %>%
  ggplot(aes(x = date, color = strategy, fill = strategy)) +
  geom_ribbon(aes(ymin = mean_temp - sd_temp, ymax = mean_temp + sd_temp),
              alpha = 0.15, color = NA) +
  geom_line(aes(y = mean_temp)) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  scale_fill_manual(values  = c(resident_warm = "red", 
                                resident_cold = "blue",
                                optimal_mover = "darkgreen")) +
  theme_bw() + #ylim(0,10000) +
  xlab("Date") + ylab("Experienced temperature (deg. C)")
```

#### Experienced growth

Experienced standardized growth rates over time, by strategy.

```{r}
ibm_summary %>%
  # filter(strategy == "optimal_mover") %>%
  ggplot(aes(x = date, color = strategy, fill = strategy)) +
  geom_ribbon(aes(ymin = mean_ggd - sd_ggd, ymax = mean_ggd + sd_ggd),
              alpha = 0.15, color = NA) +
  geom_line(aes(y = mean_ggd)) +
  scale_color_manual(values = c(resident_warm = "red", 
                                resident_cold = "blue", 
                                optimal_mover = "darkgreen")) +
  scale_fill_manual(values  = c(resident_warm = "red", 
                                resident_cold = "blue",
                                optimal_mover = "darkgreen")) +
  theme_bw() + #ylim(0,10000) +
  xlab("Date") + ylab("Experienced growth rate (g/g/d)")
```

### Boldness results

Compile end of simulation results

```{r}
fish_results <- fish_pop %>%
  mutate(
    final_weight   = mass_matrix[, n_days],
    n_switches     = switches
    # boldness       = move_threshold   # positive = shy, negative = bold
  )

fish_results %>%
  group_by(strategy) %>%
  summarise(mean_wt = round(mean(final_weight), 1),
            mean_switches = round(mean(n_switches), 1))
```

Plot only if move_stochastic = "indiv"

```{r fig.width=5, fig.height=7}
if (move_stochastic == "indiv") {
  
movers <- fish_results %>% filter(strategy == "optimal_mover")

# Panel A: boldness vs final weight
p_wt <- ggplot(movers, aes(x = boldness, y = final_weight)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "steelblue", linewidth = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  annotate("text", x =  0.006, y = min(movers$final_weight) + 5,
           label = "shy →",  hjust = 0, size = 3.5, color = "grey40") +
  annotate("text", x = -0.006, y = min(movers$final_weight) + 5,
           label = "← bold", hjust = 1, size = 3.5, color = "grey40") +
  labs(x = "Boldness (move_threshold)", y = "Final weight (g)",
       title = "A: Boldness vs. end-of-season weight")

# Panel B: boldness vs number of patch switches
p_sw <- ggplot(movers, aes(x = boldness, y = n_switches)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "tomato", linewidth = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  labs(x = "Boldness (move_threshold)", y = "No. patch switches",
       title = "B: Boldness vs. switching frequency")

# Panel C: switching frequency vs final weight  
p_sw_wt <- ggplot(movers, aes(x = n_switches, y = final_weight)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "mediumpurple", linewidth = 0.8) +
  labs(x = "No. patch switches", y = "Final weight (g)",
       title = "C: Switching frequency vs. end-of-season weight")

cowplot::plot_grid(p_wt, p_sw, p_sw_wt, ncol = 1)

}

```

### Realistic growth regime

Here we simulate the growth regime where growth potential is an additional function of fish mass to generate more realistic size trajectories. This will also serve as the basis of the IBM.

*Note that growth trajectories are super sensitive to ration size. Simulating growth under the highest ration size (0.17) generates very unrealistic growth outputs. In Armstrong et al. (2021), Fullerton's IBM simulations allowed ration to vary between 0.02 and 0.07 across river networks.*

#### Equal ration

*Should edit this section to use the updated functions*

Ration size is equal between warm and cold habitat patches (0.10)

```{r}
# set up table
growthregime_df_2 <- habitat_df %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(1))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(2))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(3))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(4))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(5))) %>%
  mutate(grpot_warm = NA, grpot_cold = NA, weight_warm = NA, weight_cold = NA) %>%
  mutate(temp_warm = trimws(temp_warm), temp_cold = trimws(temp_cold)) #%>%
  # filter(date >= date("2024-06-01"))

# set starting weight
startweight <- 10

# habitat specific ration
ration_cold <- 0.10
ration_warm <- 0.10


for (i in 1:dim(growthregime_df_2)[1]) {
  if (i == 1) {
    grpot_warm <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_warm[i]),which(rations == ration_warm),startweight]
    grpot_cold <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_cold[i]),which(rations == ration_cold),startweight]
    weight_warm <- startweight + (startweight * grpot_warm)
    weight_cold <- startweight + (startweight * grpot_cold)
    
    growthregime_df_2$grpot_warm[i] <- grpot_warm
    growthregime_df_2$grpot_cold[i] <- grpot_cold
    growthregime_df_2$weight_warm[i] <- weight_warm
    growthregime_df_2$weight_cold[i] <- weight_cold
  } else {
    grpot_warm <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_warm[i]),which(rations == ration_warm),round(growthregime_df_2$weight_warm[i-1], digits = 0)]
    grpot_cold <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_cold[i]),which(rations == ration_cold),round(growthregime_df_2$weight_cold[i-1], digits = 0)]
    weight_warm <- growthregime_df_2$weight_warm[i-1] + (growthregime_df_2$weight_warm[i-1] * grpot_warm)
    weight_cold <- growthregime_df_2$weight_cold[i-1] + (growthregime_df_2$weight_cold[i-1] * grpot_cold)
    
    growthregime_df_2$grpot_warm[i] <- grpot_warm
    growthregime_df_2$grpot_cold[i] <- grpot_cold
    growthregime_df_2$weight_warm[i] <- weight_warm
    growthregime_df_2$weight_cold[i] <- weight_cold
  }
  # print(i)
}


p1 <- growthregime_df_2 %>% ggplot() +
  geom_line(aes(x = date, y = grpot_warm), col = "red") +
  geom_line(aes(x = date, y = grpot_cold), col = "blue") +
  theme_bw() +
  xlab("Date") +
  ylab("Growth potential (g/g/day)")

p2 <- growthregime_df_2 %>% ggplot() +
  geom_line(aes(x = date, y = weight_warm), col = "red") +
  geom_line(aes(x = date, y = weight_cold), col = "blue") +
  # geom_line(aes(x = date, y = cumul_grpot_track), col = "purple", linetype = "dashed") +
  theme_bw() +
  xlab("Date") +
  ylab("Cumulative growth potential (g)")

ggarrange(p1, p2)
```

#### Unequal ration

Ration size is 30% lower in the cold habitat

```{r}
# set up table
growthregime_df_2 <- habitat_df %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(1))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(2))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(3))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(4))) %>% 
  # bind_rows(temp_df %>% mutate(date = date + years(5))) %>%
  mutate(grpot_warm = NA, grpot_cold = NA, weight_warm = NA, weight_cold = NA) %>%
  mutate(temp_warm = trimws(temp_warm), temp_cold = trimws(temp_cold)) #%>%
  # filter(date >= date("2024-06-01"))

# set starting weight
startweight <- 10

# habitat specific ration
ration_cold <- 0.07
ration_warm <- 0.10


for (i in 1:dim(growthregime_df_2)[1]) {
  if (i == 1) {
    grpot_warm <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_warm[i]),which(rations == ration_warm),startweight]
    grpot_cold <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_cold[i]),which(rations == ration_cold),startweight]
    weight_warm <- startweight + (startweight * grpot_warm)
    weight_cold <- startweight + (startweight * grpot_cold)
    
    growthregime_df_2$grpot_warm[i] <- grpot_warm
    growthregime_df_2$grpot_cold[i] <- grpot_cold
    growthregime_df_2$weight_warm[i] <- weight_warm
    growthregime_df_2$weight_cold[i] <- weight_cold
  } else {
    grpot_warm <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_warm[i]),which(rations == ration_warm),round(growthregime_df_2$weight_warm[i-1], digits = 0)]
    grpot_cold <- wt.growth[which(waterTemps[,2] == growthregime_df_2$temp_cold[i]),which(rations == ration_cold),round(growthregime_df_2$weight_cold[i-1], digits = 0)]
    weight_warm <- growthregime_df_2$weight_warm[i-1] + (growthregime_df_2$weight_warm[i-1] * grpot_warm)
    weight_cold <- growthregime_df_2$weight_cold[i-1] + (growthregime_df_2$weight_cold[i-1] * grpot_cold)
    
    growthregime_df_2$grpot_warm[i] <- grpot_warm
    growthregime_df_2$grpot_cold[i] <- grpot_cold
    growthregime_df_2$weight_warm[i] <- weight_warm
    growthregime_df_2$weight_cold[i] <- weight_cold
  }
  # print(i)
}


p1 <- growthregime_df_2 %>% ggplot() +
  geom_line(aes(x = date, y = grpot_warm), col = "red") +
  geom_line(aes(x = date, y = grpot_cold), col = "blue") +
  theme_bw() +
  xlab("Date") +
  ylab("Growth potential (g/g/day)")

p2 <- growthregime_df_2 %>% ggplot() +
  geom_line(aes(x = date, y = weight_warm), col = "red") +
  geom_line(aes(x = date, y = weight_cold), col = "blue") +
  # geom_line(aes(x = date, y = cumul_grpot_track), col = "purple", linetype = "dashed") +
  theme_bw() +
  xlab("Date") +
  ylab("Cumulative growth potential (g)")

ggarrange(p1, p2)
```
