---
title: "Surplus Production"
---

```{r, include = FALSE, cache = FALSE}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(egg)
library(RColorBrewer)
library(ggridges)
library(patchwork)
library(scales)
library(ggrepel)

source("Functions.R")
source("PlotResults.R")
source("ScenarioDefs.R")
```

**Purpose:** Compare surplus production and the impacts of "harvest" across habitat scenarios.

## Setup

Load relevant .rds objects and bind directly into a list. Using a helper function ensures each `res` object is freed from memory as soon as `ibm_long` is extracted, avoiding duplicate references that would force R to copy the data frames.
```{r}
#| cache: false

# Extract params and habitat_df from a representative 50%-cold scenario
res            <- readRDS("results/TempOffset_ColdOnly_50percold.rds")
sim_params     <- res$params
habitat_df     <- res$habitat_df
exp_start_date <- sim_params$mindate + years(sim_params$nyears_burnin)
rm(res)

# Helper: read one scenario, extract ibm_long, and free res before returning
load_scen <- function(path, ht, pc, ac, aw, hr, pcmax_cold, pcmax_warm) {
  res <- readRDS(path)
  list(df = res$ibm_long, harvest = res$harvest_log, 
       ht = ht, pc = pc, ac = ac, aw = aw, hr = hr, pcmax_cold = pcmax_cold, pcmax_warm = pcmax_warm)
}

ibm_list_mech <- list(
  load_scen("results/TempOffset_ColdOnly_50percold.rds",  "Cold only",  0.50,  1.0,  0.0,  0.0,  0.5,  NA_real_),
  load_scen("results/Harvest_10_ColdOnly_50percold.rds",  "Cold only",  0.50,  1.0,  0.0,  0.1,  0.5,  NA_real_),
  load_scen("results/Harvest_20_ColdOnly_50percold.rds",  "Cold only",  0.50,  1.0,  0.0,  0.2,  0.5,  NA_real_),
  load_scen("results/Harvest_30_ColdOnly_50percold.rds",  "Cold only",  0.50,  1.0,  0.0,  0.3,  0.5,  NA_real_),
  load_scen("results/Harvest_40_ColdOnly_50percold.rds",  "Cold only",  0.50,  1.0,  0.0,  0.4,  0.5,  NA_real_),
  load_scen("results/Harvest_50_ColdOnly_50percold.rds",  "Cold only",  0.50,  1.0,  0.0,  0.5,  0.5,  NA_real_),
  load_scen("results/Harvest_70_ColdOnly_50percold.rds",  "Cold only",  0.50,  1.0,  0.0,  0.7,  0.5,  NA_real_),
  load_scen("results/Harvest_90_ColdOnly_50percold.rds",  "Cold only",  0.50,  1.0,  0.0,  0.9,  0.5,  NA_real_),
  
  load_scen("results/TempOffset_ColdWarm_50percold.rds",  "Cold + Warm",  0.50,  1.0,  1.0,  0.0,  0.5,  0.5),
  load_scen("results/Harvest_10_ColdWarm_50percold.rds",  "Cold + Warm",  0.50,  1.0,  1.0,  0.1,  0.5,  0.5),
  load_scen("results/Harvest_20_ColdWarm_50percold.rds",  "Cold + Warm",  0.50,  1.0,  1.0,  0.2,  0.5,  0.5),
  load_scen("results/Harvest_30_ColdWarm_50percold.rds",  "Cold + Warm",  0.50,  1.0,  1.0,  0.3,  0.5,  0.5),
  load_scen("results/Harvest_40_ColdWarm_50percold.rds",  "Cold + Warm",  0.50,  1.0,  1.0,  0.4,  0.5,  0.5),
  load_scen("results/Harvest_50_ColdWarm_50percold.rds",  "Cold + Warm",  0.50,  1.0,  1.0,  0.5,  0.5,  0.5),
  load_scen("results/Harvest_70_ColdWarm_50percold.rds",  "Cold + Warm",  0.50,  1.0,  1.0,  0.7,  0.5,  0.5),
  load_scen("results/Harvest_90_ColdWarm_50percold.rds",  "Cold + Warm",  0.50,  1.0,  1.0,  0.9,  0.5,  0.5)
)
```


Define misc. objects:
```{r}
#| cache: false

# ── Scenario metadata ─────────────────────────────────────────────────────────
scen_meta <- tibble(
  scen_key  = c("95", "75", "50", "25", "05"),
  label     = c("95% cold", "75% cold", "50% cold", "25% cold", "5% cold"),
  prop_cold = c(0.95, 0.75, 0.50, 0.25, 0.05)
)

# factor levels
pc_levels <- c(0.95, 0.75, 0.50, 0.25, 0.05)
pc_levels_mech <- c(0.95, 0.75, 0.50, 0.25, 0.05)
pc_labels_mech <- c("95% cold","75% cold","50% cold","25% cold","5% cold")

hr_levels <- c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.7, 0.9)
hr_levels_mech <- c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.7, 0.9)
hr_labels_mech <- c("0%","10%","20%","30%","40%", "50%", "70%", "90%")


# color palette for prop_cold
# scen_pal <- setNames(
#   hcl.colors(5, "Zissou 1"),
#   scen_meta$label
# )

# Linetype and shape palettes for habitat availability
hab_cols  <- c("Cold only" = "#6BAED6",  "Cold + Warm" = "#FC8D59",   "Cold + Warm (high P)" = "#B30000")
hab_lty   <- c("Cold only" = "solid" ,   "Cold + Warm" = "solid",     "Cold + Warm (high P)" = "solid")
hab_shape <- c("Cold only" = 16,         "Cold + Warm" = 16,          "Cold + Warm (high P)" = 16)

```


## View simulation output

### Cold-only

#### 10% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[2]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 20% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[3]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 30% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[4]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 40% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[5]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 50% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[6]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 70% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[7]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

#### 90% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[8]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```



### Cold + Warm

#### 10% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[10]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 20% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[11]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 30% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[12]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 40% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[13]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 50% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[14]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 70% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[15]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


#### 90% harvest

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[16]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```


## Recruitment

The number of fry produced is a function of adult size and abundance (fecundity), which are driven by warm and cold habitat availability. Importantly, when only cold habitat is available, the number of fry produced declines ~linearly with cold habitat availability. But when warm habitat is also available, considerably more fry are produced at a given proportion cold (except for 95% cold); i.e., 

```{r}
#| cache: false
#| fig-width: 7
#| fig-height: 6

# ── Annual age-0 recruitment time series ─────────────────────────────────────
# Zero-fill missing years so post-crash scenarios contribute zeros for the
# remainder of the experimental period rather than being silently excluded.
all_exp_years <- tibble(
  yr = seq(year(exp_start_date),
           year(exp_start_date) + sim_params$nyears_experiment - 1L)
)

get_annual_age0 <- function(scen) {
  scen$df |>
    filter(date >= exp_start_date,
           month(date) == 11, day(date) == 1,
           survived == 1, floor(age) == 0) |>
    count(yr = year(date), name = "n_age0") |>
    right_join(all_exp_years, by = "yr") |>
    replace_na(list(n_age0 = 0L)) |>
    mutate(hab_type = scen$ht,
           harvest_rate = factor(scen$hr,
                                 levels = hr_levels_mech,
                                 labels = hr_labels_mech))
}

age0_ts <- purrr::map_dfr(ibm_list_mech, get_annual_age0)


# ── End-vs-start recruitment ratio ───────────────────────────────────────────
# Collapses each trajectory to a single scalar: mean n_age0 in the last 10
# experimental years divided by the first 10. Stable harvest → ratio ≈ 1;
# slow decline → ratio < 1; crashed population → ratio = 0.
n_exp   <- nrow(all_exp_years)
first10 <- all_exp_years$yr[1:10]
last10  <- all_exp_years$yr[(n_exp - 9):n_exp]

ratio_df <- age0_ts |>
  mutate(period = case_when(
    yr %in% first10 ~ "start",
    yr %in% last10  ~ "end",
    TRUE            ~ NA_character_
  )) |>
  filter(!is.na(period)) |>
  group_by(hab_type, harvest_rate, period) |>
  summarise(mean_n = mean(n_age0), sd_n = sd(n_age0), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = c(mean_n, sd_n)) |>
  mutate(ratio = mean_n_end / mean_n_start) |>
  left_join(
    age0_ts |>
      group_by(hab_type, harvest_rate) |>
      summarise(mean_n_all = mean(n_age0), sd_n_all = sd(n_age0), .groups = "drop"),
    by = c("hab_type", "harvest_rate")
  )


# Palette keys must match levels(age0_ts$harvest_rate)
cols <- hcl.colors(8, "inferno", rev = TRUE)
hr_pal <- c("0%"  = "grey70", "10%" = cols[2], "20%" = cols[3], "30%" = cols[4],
            "40%" = cols[5], "50%" = cols[6], "70%" = cols[7], "90%" = cols[8])


#### COLD ONLY
p_ts <- ggplot(age0_ts %>% filter(hab_type == "Cold only"), aes(x = yr, y = n_age0, color = harvest_rate)) +
  geom_line(alpha = 0.3) +
  geom_smooth(linewidth = 0.8, se = FALSE) +
  scale_color_manual(values = hr_pal, guide = "none") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Age-0 abundance (Nov. 1)",
                     limits = c(0, NA),
                     labels = scales::label_comma()) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())


p_all <- ggplot(ratio_df %>% filter(hab_type == "Cold only"), aes(x = harvest_rate, y = mean_n_all, color = harvest_rate)) +
  geom_errorbar(aes(ymin = mean_n_all - sd_n_all, ymax = mean_n_all + sd_n_all), width = 0.025, linewidth = 0.6) +
  geom_point(size = 4) +
  scale_color_manual(values = hr_pal, guide = "none") +
  scale_y_continuous(name = "Mean recruitment (all years)",) +
  coord_cartesian(ylim = c(0,NA)) +
  labs(x = "Harvest rate") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x      = element_text(angle = 45, hjust = 1))

p_end <- ggplot(ratio_df %>% filter(hab_type == "Cold only"), aes(x = harvest_rate, y = mean_n_end, color = harvest_rate)) +
  geom_errorbar(aes(ymin = mean_n_end - sd_n_end, ymax = mean_n_end + sd_n_end), width = 0.025, linewidth = 0.6) +
  geom_point(size = 4) +
  scale_color_manual(values = hr_pal, guide = "none") +
  scale_y_continuous(name = "Mean recruitment (final 10 years)",
                     limits = c(0, NA)) +
  labs(x = "Harvest rate") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x      = element_text(angle = 45, hjust = 1))

p_ratio <- ggplot(ratio_df %>% filter(hab_type == "Cold only"), aes(x = harvest_rate, y = ratio, color = harvest_rate)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey40", linewidth = 0.6) +
  geom_segment(aes(xend = harvest_rate, yend = 1), linewidth = 0.8) +
  geom_point(size = 4) +
  scale_color_manual(values = hr_pal, name = "Harvest rate") +
  scale_y_continuous(name = "End / start recruitment ratio",
                     limits = c(0, NA)) +
  labs(x = "Harvest rate") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x      = element_text(angle = 45, hjust = 1))

(p_ts / (p_all + p_end + p_ratio)) + plot_layout(guides = "collect") +
  plot_annotation(title = "Cold only -- November recruitment over variable harvest rates")



#### COLD + WARM
p_ts <- ggplot(age0_ts %>% filter(hab_type == "Cold + Warm"), aes(x = yr, y = n_age0, color = harvest_rate)) +
  geom_line(alpha = 0.3) +
  geom_smooth(linewidth = 0.8, se = FALSE) +
  scale_color_manual(values = hr_pal, guide = "none") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Age-0 abundance (Nov. 1)",
                     limits = c(0, NA),
                     labels = scales::label_comma()) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())


p_all <- ggplot(ratio_df %>% filter(hab_type == "Cold + Warm"), aes(x = harvest_rate, y = mean_n_all, color = harvest_rate)) +
  geom_errorbar(aes(ymin = mean_n_all - sd_n_all, ymax = mean_n_all + sd_n_all), width = 0.025, linewidth = 0.6) +
  geom_point(size = 4) +
  scale_color_manual(values = hr_pal, guide = "none") +
  scale_y_continuous(name = "Mean recruitment (all years)",) +
  coord_cartesian(ylim = c(0,NA)) +
  labs(x = "Harvest rate") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x      = element_text(angle = 45, hjust = 1))

p_end <- ggplot(ratio_df %>% filter(hab_type == "Cold + Warm"), aes(x = harvest_rate, y = mean_n_end, color = harvest_rate)) +
  geom_errorbar(aes(ymin = mean_n_end - sd_n_end, ymax = mean_n_end + sd_n_end), width = 0.025, linewidth = 0.6) +
  geom_point(size = 4) +
  scale_color_manual(values = hr_pal, guide = "none") +
  scale_y_continuous(name = "Mean recruitment (final 10 years)",
                     limits = c(0, NA)) +
  labs(x = "Harvest rate") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x      = element_text(angle = 45, hjust = 1))

p_ratio <- ggplot(ratio_df %>% filter(hab_type == "Cold + Warm"), aes(x = harvest_rate, y = ratio, color = harvest_rate)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey40", linewidth = 0.6) +
  geom_segment(aes(xend = harvest_rate, yend = 1), linewidth = 0.8) +
  geom_point(size = 4) +
  scale_color_manual(values = hr_pal, name = "Harvest rate") +
  scale_y_continuous(name = "End / start recruitment ratio",
                     limits = c(0, NA)) +
  labs(x = "Harvest rate") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x      = element_text(angle = 45, hjust = 1))

(p_ts / (p_all + p_end + p_ratio)) + plot_layout(guides = "collect") +
  plot_annotation(title = "Cold + Warm -- November recruitment over variable harvest rates")

```


## Harvest

Summarize harvest data
```{r}
all_exp_years <- tibble(
  year = seq(year(exp_start_date),
           year(exp_start_date) + sim_params$nyears_experiment - 1L)
)

get_annual_harvest <- function(scen) {
  scen$harvest %>%
    group_by(year) %>%
    summarise(n_harvested = n(),
              bm_harvested = sum(weight)/1000,
              mean_wt = mean(weight), 
              sd_wt = sd(weight),
              mean_age = mean(age),
              sd_age = sd(age),
              mean_cond = mean(condition),
              sd_cond = sd(condition)) %>%
    ungroup() %>%
    right_join(all_exp_years, by = "year") %>%
    replace_na(list(n_harvested = 0L,
                    bm_harvested = 0L,
                    mean_wt = NA,
                    sd_wt = NA,
                    mean_age = NA,
                    sd_age = NA,
                    mean_cond = NA,
                    sd_cond = NA)) %>%
    mutate(hab_type = scen$ht,
           harvest_rate = factor(scen$hr,
                                 levels = hr_levels_mech,
                                 labels = hr_labels_mech))
}

get_harvest_traits <- function(scen) {
  scen$harvest %>%
    summarise(mean_wt = mean(weight), 
              sd_wt = sd(weight),
              mean_age = mean(age),
              sd_age = sd(age),
              mean_cond = mean(condition),
              sd_cond = sd(condition)) %>%
    ungroup() %>%
    mutate(hab_type = scen$ht,
           harvest_rate = factor(scen$hr,
                                 levels = hr_levels_mech,
                                 labels = hr_labels_mech))
}


harvest_ts <- purrr::map_dfr(ibm_list_mech[c(2:8,10:16)], get_annual_harvest) |>
  bind_rows(
    purrr::map_dfr(list(ibm_list_mech[[1]], ibm_list_mech[[9]]), function(scen) {
      tibble(
        year         = sort(unique(age0_ts$yr)),
        n_harvested  = 0L,
        bm_harvested = 0,
        mean_wt      = NA_real_,
        sd_wt        = NA_real_,
        mean_age     = NA_real_,
        sd_age       = NA_real_,
        mean_cond    = NA_real_,
        sd_cond      = NA_real_,
        hab_type     = scen$ht,
        harvest_rate = factor(scen$hr,
                              levels = hr_levels_mech,
                              labels = hr_labels_mech)
      )
    })
  ) %>%
  mutate(hab_type = factor(hab_type, levels = c("Cold only", "Cold + Warm")))

harvest_traits <- purrr::map_dfr(ibm_list_mech[c(2:8,10:16)], get_harvest_traits) |>
  bind_rows(
    purrr::map_dfr(list(ibm_list_mech[[1]], ibm_list_mech[[9]]), function(scen) {
      tibble(
        mean_wt      = NA_real_,
        sd_wt        = NA_real_,
        mean_age     = NA_real_,
        sd_age       = NA_real_,
        mean_cond    = NA_real_,
        sd_cond      = NA_real_,
        hab_type     = scen$ht,
        harvest_rate = factor(scen$hr,
                              levels = hr_levels_mech,
                              labels = hr_labels_mech)
      )
    })
  ) %>%
  mutate(hab_type = factor(hab_type, levels = c("Cold only", "Cold + Warm")))


harvest_summ <- harvest_ts %>% 
  group_by(hab_type, harvest_rate) %>%
  summarize(mean_n = mean(n_harvested),
            sd_n = sd(n_harvested),
            mean_bm = mean(bm_harvested), 
            sd_bm = sd(bm_harvested)) %>%
  ungroup() %>%
  left_join(harvest_traits)
```

Show time series of harvested abundance, by habitat configuration and harvest rate
```{r}
#| cache: false
#| fig-width: 8
#| fig-height: 8

p1 <- ggplot(harvest_ts, aes(x = year, y = n_harvested, color = harvest_rate)) +
  geom_line(alpha = 0.3) +
  geom_smooth(linewidth = 0.8, se = FALSE) +
  scale_color_manual(values = hr_pal) +
  coord_cartesian(ylim = c(0, 90)) +
  labs(x = "Year", y = "Harvested abundance (N)") +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  facet_wrap(~hab_type, ncol = 2)

p2 <- ggplot(harvest_ts, aes(x = year, y = bm_harvested, color = harvest_rate)) +
  geom_line(alpha = 0.3) +
  geom_smooth(linewidth = 0.8, se = FALSE) +
  scale_color_manual(values = hr_pal) +
  coord_cartesian(ylim = c(0, 50)) +
  labs(x = "Year", y = "Harvested biomass (kg)") +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  facet_wrap(~hab_type, ncol = 2)

(p1 / p2) + plot_layout(guides = "collect")
```

Time series of individual-level traits

::: {.panel-tabset}

### Weight

```{r}
#| cache: false
#| fig-width: 8
#| fig-height: 8

ggplot(harvest_ts, aes(x = year, y = mean_wt, color = harvest_rate, shape = hab_type)) +
  geom_errorbar(aes(ymin = mean_wt - sd_wt, ymax = mean_wt + sd_wt), width = 0.025, linewidth = 0.6) +
  geom_point(size = 2) +
  facet_grid(harvest_rate ~ hab_type) +
  scale_color_manual(values = hr_pal, guide = "none") +
  labs(x = "Year", y = "Mean individual weight of harvested fish") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```


### Age

```{r}
#| cache: false
#| fig-width: 8
#| fig-height: 8

ggplot(harvest_ts, aes(x = year, y = mean_age, color = harvest_rate)) +
  geom_errorbar(aes(ymin = mean_age - sd_age, ymax = mean_age + sd_age), width = 0.025, linewidth = 0.6) +
  geom_point(size = 2) +
  facet_grid(harvest_rate ~ hab_type) +
  scale_color_manual(values = hr_pal, guide = "none") +
  labs(x = "Year", y = "Mean individual age of harvested fish") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```


### Condition

```{r}
#| cache: false
#| fig-width: 8
#| fig-height: 8

ggplot(harvest_ts, aes(x = year, y = mean_cond, color = harvest_rate)) +
  geom_errorbar(aes(ymin = mean_cond - sd_cond, ymax = mean_cond + sd_cond), width = 0.025, linewidth = 0.6) +
  geom_point(size = 2) +
  facet_grid(harvest_rate ~ hab_type) +
  scale_color_manual(values = hr_pal, guide = "none") +
  labs(x = "Year", y = "Mean individual condition of harvested fish") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

:::



```{r}
#| cache: false
#| fig-width: 8
#| fig-height: 7

# mean harvested abundance
p_abun <- ggplot(harvest_summ, aes(x = harvest_rate, y = mean_n, color = hab_type, group = hab_type)) +
  geom_errorbar(aes(ymin = mean_n - sd_n, ymax = mean_n + sd_n), width = 0.025, linewidth = 0.6, position = position_dodge(width = 0.3)) +
  geom_point(size = 4, position = position_dodge(width = 0.3)) +
  scale_color_manual(values = hab_cols) +
  scale_y_continuous(name = "Mean annnual harvested abundance",
                     limits = c(0, NA)) +
  labs(x = "Harvest rate") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x      = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

# mean harvested biomass
p_biom <- ggplot(harvest_summ, aes(x = harvest_rate, y = mean_bm, color = hab_type, group = hab_type)) +
  geom_errorbar(aes(ymin = mean_bm - sd_bm, ymax = mean_bm + sd_bm), width = 0.025, linewidth = 0.6, position = position_dodge(width = 0.3)) +
  geom_point(size = 4, position = position_dodge(width = 0.3)) +
  scale_color_manual(values = hab_cols) +
  scale_y_continuous(name = "Mean annnual harvested biomass (kg)",
                     limits = c(0, NA)) +
  labs(x = "Harvest rate") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x      = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

# mean individual weight
p_wt <- ggplot(harvest_summ, aes(x = harvest_rate, y = mean_wt, color = hab_type, group = hab_type)) +
  geom_errorbar(aes(ymin = mean_wt - sd_wt, ymax = mean_wt + sd_wt), width = 0.025, linewidth = 0.6, position = position_dodge(width = 0.3)) +
  geom_point(size = 4, position = position_dodge(width = 0.3)) +
  scale_color_manual(values = hab_cols) +
  scale_y_continuous(name = "Mean individual weight (g)",
                     limits = c(0, NA)) +
  labs(x = "Harvest rate") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x      = element_text(angle = 45, hjust = 1))

# mean individual age
p_age <- ggplot(harvest_summ, aes(x = harvest_rate, y = mean_age, color = hab_type, group = hab_type)) +
  geom_errorbar(aes(ymin = mean_age - sd_age, ymax = mean_age + sd_age), width = 0.025, linewidth = 0.6, position = position_dodge(width = 0.3)) +
  geom_point(size = 4, position = position_dodge(width = 0.3)) +
  scale_color_manual(values = hab_cols) +
  scale_y_continuous(name = "Mean individual age (years)",
                     limits = c(0, NA)) +
  labs(x = "Harvest rate") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x      = element_text(angle = 45, hjust = 1))


(p_abun + p_biom) / (p_wt + p_age) + plot_layout(guides = "collect")

```

## Fry production vs. recruitment

Declining fry input with stable recruitment output demonstrates density-dependent compensation at the cold-habitat summer bottleneck. The gap between the two panels quantifies how much excess fry production the system can absorb before the bottleneck is breached.

```{r eval=FALSE}
#| cache: false
#| fig-width: 7
#| fig-height: 6

get_annual_fry <- function(scen) {
  scen$df |>
    filter(date >= exp_start_date,
           !is.na(parent_pid),
           dayofsim == birth_dayofsim) |>
    count(yr = year(date), name = "n_fry") |>
    right_join(all_exp_years, by = "yr") |>
    replace_na(list(n_fry = 0L)) |>
    mutate(harvest_rate = factor(scen$hr,
                                 levels = hr_levels_mech,
                                 labels = hr_labels_mech),
           hab_type = scen$ht)
}

fry_ts    <- purrr::map_dfr(ibm_list_mech, get_annual_fry) |>
  filter(hab_type == "Cold only")
age0_cold <- age0_ts |>
  filter(hab_type == "Cold only")

p_fry <- ggplot(fry_ts, aes(x = yr, y = n_fry, color = harvest_rate)) +
  geom_line(alpha = 0.25) +
  geom_smooth(linewidth = 0.8, se = FALSE) +
  scale_color_manual(values = hr_pal, guide = "none") +
  scale_x_continuous(name = NULL) +
  scale_y_continuous(name = "Annual fry production",
                     labels = scales::label_comma(),
                     limits = c(0, NA)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

p_rec <- ggplot(age0_cold, aes(x = yr, y = n_age0, color = harvest_rate)) +
  geom_line(alpha = 0.25) +
  geom_smooth(linewidth = 0.8, se = FALSE) +
  scale_color_manual(values = hr_pal, name = "Harvest rate") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Age-0 recruits (Nov. 1)",
                     labels = scales::label_comma(),
                     limits = c(0, NA)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

p_fry / p_rec + plot_layout(guides = "collect") & theme(legend.position = "right")
```

