---
title: "View Simulation Results"
---

```{r, include = FALSE, cache = FALSE}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(egg)
library(RColorBrewer)
library(ggridges)
library(patchwork)
library(scales)

source("Functions.R")
source("PlotResults.R")
```

**Purpose:** View raw simulation results of fixed habitat/equilibrium simulation runs. 

## Setup

Load each result .rds object and bind directly into a list. Using a helper function ensures each `res` object is freed from memory as soon as `ibm_long` is extracted, avoiding duplicate references that would force R to copy the data frames.
```{r}
#| cache: false

# Extract params from one file (small object)
res            <- readRDS("results/TempOffset_ColdWarm_05percold_highPwarm.rds")
sim_params     <- res$params
habitat_df     <- res$habitat_df
exp_start_date <- sim_params$mindate + years(sim_params$nyears_burnin)
rm(res)

# Helper: read one scenario, extract ibm_long, and free res before returning
load_scen <- function(path, pc, ht, pcmax_cold, pcmax_warm) {
  res <- readRDS(path)
  list(df = res$ibm_long, pc = pc, ht = ht, pcmax_cold = pcmax_cold, pcmax_warm = pcmax_warm)
}

ibm_list_mech <- list(
  load_scen("results/TempOffset_ColdOnly_95percold.rds",             0.95, "Cold only",            0.5, NA_real_),
  load_scen("results/TempOffset_ColdOnly_75percold.rds",             0.75, "Cold only",            0.5, NA_real_),
  load_scen("results/TempOffset_ColdOnly_50percold.rds",             0.50, "Cold only",            0.5, NA_real_),
  load_scen("results/TempOffset_ColdOnly_25percold.rds",             0.25, "Cold only",            0.5, NA_real_),
  load_scen("results/TempOffset_ColdOnly_05percold.rds",             0.05, "Cold only",            0.5, NA_real_),

  load_scen("results/TempOffset_ColdWarm_95percold.rds",             0.95, "Cold + Warm",          0.5, 0.5),
  load_scen("results/TempOffset_ColdWarm_75percold.rds",             0.75, "Cold + Warm",          0.5, 0.5),
  load_scen("results/TempOffset_ColdWarm_50percold.rds",             0.50, "Cold + Warm",          0.5, 0.5),
  load_scen("results/TempOffset_ColdWarm_25percold.rds",             0.25, "Cold + Warm",          0.5, 0.5),
  load_scen("results/TempOffset_ColdWarm_05percold.rds",             0.05, "Cold + Warm",          0.5, 0.5),
  
  load_scen("results/TempOffset_ColdWarm_95percold_highPwarm.rds",   0.95, "Cold + Warm (high P)", 0.5, 0.6),
  load_scen("results/TempOffset_ColdWarm_75percold_highPwarm.rds",   0.75, "Cold + Warm (high P)", 0.5, 0.6),
  load_scen("results/TempOffset_ColdWarm_50percold_highPwarm.rds",   0.50, "Cold + Warm (high P)", 0.5, 0.6),
  load_scen("results/TempOffset_ColdWarm_25percold_highPwarm.rds",   0.25, "Cold + Warm (high P)", 0.5, 0.6),
  load_scen("results/TempOffset_ColdWarm_05percold_highPwarm.rds",   0.05, "Cold + Warm (high P)", 0.5, 0.6)
)
```


## Null: cold only

### 95% cold

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[1]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 75% cold

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[2]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 50% cold

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[3]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 25% cold

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[4]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 5% cold

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[5]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

---

## Cold + Warm, same Pcmax

### 95% cold / 5% warm

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[6]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 75% cold / 25% warm

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[7]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 50% cold / 50% warm

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[8]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 25% cold / 75% warm

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[9]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 5% cold / 95% warm

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[10]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

---

## Cold + Warm, high Pcmax in warm (0.6)

### 95% cold / 5% warm

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[11]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 75% cold / 25% warm

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[12]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 50% cold / 50% warm

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[13]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 25% cold / 75% warm

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[14]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

### 5% cold / 95% warm

Daily and annual abundance/biomass, age structure, weight-at-age, body-size distributions, survival, lambda, and cohort survivorship.

```{r}
#| fig-width: 10
#| fig-height: 6
plot_population(
  ibm_long       = ibm_list_mech[[15]][[1]],
  habitat_df     = habitat_df,
  exp_start_date = exp_start_date
)
```

---
